---
title: Analyze CHOP Study Data
author: 
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceFile: "Analyze_Data.qmd"
  LogFile: "Analyze_Data.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
    cfsize: "scriptsize"
---

\FloatBarrier

# Purpose

This file provides the analysis results of data file 
`Vinc Cases 1.9.24.xlsx` provided by Alison and Caitlin, which records 
observations from dogs receiving CHOP chemotherapy at MSU VMC between 2016 and 
2022.

::: {.callout-note}
The data file we received was called `Vinc Cases 1.9.23.xlsx`, but we renamed it 
to `Vinc Cases 1.9.24.xlsx` before using it because that was the date we 
received it. 
:::

# Setup

This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Results.

## Define Global Options

Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages

Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(here)
library(kableExtra)
library(dplyr)
library(knitr)
library(gtsummary)
library(glmmTMB)
library(readxl)
library(quarto)
library(lme4)
library(emmeans)
library(ggplot2)
```

## Declare Path

```{r}
#| label: declare-path

i_am("Masyr_C1358_2024-05-03.qmd")
```

## Read Data {#sec-Read-Data}

```{r}
#| label: load-data

data_all <- read_xlsx(here("../Data/Vinc Cases 1.9.24.xlsx")) 
names(data_all)
```

# Results {#sec-Results}

## Total bands

### Summary of total bands

@fig-chop1_phase1_total_bands shows the histogram (left) and the boxplot (right) 
of **Total bands** in the original units (x10.e3/uL=1000/uL), while 
@fig-chop1_phase1_total_bands_rescaled shows the same data after rescaling to 
units of (x100/ul) to obtain integer values usable in count models. This makes 
it more convenient for us when fitting models. Both these figures show that 
there are lots of zero values, which makes it hard to see the shape of the 
distribution among the non-zero values. 

```{r}
#| label: fig-chop1_phase1_total_bands
#| fig-cap: Histogram (left) and boxplot (right) of the total bands (x1000/ul)
par(mfrow=c(1,2))
hist(data_analyse$Total_Bands, main=NULL,
     xlab="Total bands (x1000/ul)",
     ylab = "Number of patients")
boxplot(data_analyse$Total_Bands, ylab="Total bands (x1000/ul)")
par(mfrow=c(1,1))
```

```{r}
#| label: fig-chop1_phase1_total_bands_rescaled
#| fig-cap: Histogram (left) and boxplot (right) of the total bands (x100/ul)
par(mfrow=c(1,2))
hist(data_analyse$Total_Bands_Rescaled, main = NULL, 
     xlab="Total bands (x100/ul)",
     ylab = "Number of patients")
boxplot(data_analyse$Total_Bands_Rescaled, ylab="Total bands (x100/ul)")
par(mfrow=c(1,1))
```

@tbl-chop1_bands_phase1 gives the mean and median value of the **Total bands** 
(x100/ul). We can see that more than 75% outcomes are 0. Therefore, it's 
possible that this distribution is **zero-inflated** (i.e., there may be more 
values of zero than one would expect under a Poisson distribution).

```{r}
#| label: tbl-chop1_bands_phase1
#| tbl-cap: Summary of total bands (x100/ul)
A=summary(data_analyse$Total_Bands_Rescaled)
round(as.matrix(A),4) %>% 
  kbl() %>% 
  kable_styling()
```

The histogram (left) and boxplot (right) of **log(Total bands [x100/ul])** and 
**log(Total bands [x1000/ul])** are shown in 
@fig-chop1_phase1_total_bands_rescaled_log and 
@fig-chop1_phase1_total_bands_log, respectively. The difference between these 
two graphs are in the units of measurement prior to the log transformation. In 
both graphs, the log equals -Infinity when total bands are 0, so we recoded 
these values to improve the figures and keep them compact and readable. And we 
also present figures in 
@fig-chop1_phase1_total_bands_rescaled_log_part and 
@fig-chop1_phase1_total_bands_log_part that show only the data where total bands 
\> 0 to better look at the shape of the distribution.

```{r}
#| label: fig-chop1_phase1_total_bands_rescaled_log
#| fig-cap: Histogram (left) and boxplot (right) of the log(total bands x100/ul)

par(mfrow=c(1,2))
hist(data_analyse$log_Total_Bands_Rescaled,
     xlab="log(Total bands [x100/ul])",
     ylab = "Number of patients",
     main=NULL,
     breaks = 7, xlim = c(-3, 5), xaxt = "n", ylim = c(0, 500))
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(1, at = axis_positions, labels = axis_labels)

boxplot(data_analyse$log_Total_Bands_Rescaled,
       ylab="log(Total bands [x100/ul])",
       ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)

par(mfrow=c(1,1))
```

```{r}
#| label: fig-chop1_phase1_total_bands_log
#| fig-cap: Histogram (left) and boxplot (right) of the log(total bands x1000/ul)

par(mfrow=c(1,2))
hist(data_analyse$log_Total_Bands, 
     xlab="log(Total bands [x1000/ul])", 
     ylab = "Number of patients",
     main=NULL, 
     breaks = 6, xlim = c(-5, 2), xaxt = "n", ylim = c(0, 500))
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(1, at = axis_positions, labels = axis_labels)

boxplot(data_analyse$log_Total_Bands,
        ylab="log(Total bands [x1000/ul])",
        ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)

par(mfrow=c(1,1))
```

::: {.callout-note}
@fig-chop1_phase1_total_bands_log above becomes Figure 1 in the manuscript. 
:::

@tbl-Numbers_of_zero_and_proportion shows the frequency distribution for Total 
Bands, which highlights the large proportion of zero values. The table also 
shows the rescaled Total Bands values used in count models, and log transformed
values for both original and recscaled Total Bands values. 

```{r}
#| label: tbl-Numbers_of_zero_and_proportion
#| tbl-cap: "Frequency Distribution of Total Bands"

data_analyse %>% 
  count(Total_Bands) %>% 
  mutate(Total_Bands_Rescaled = Total_Bands*10,
         Log_Total_Bands = log(Total_Bands), 
         Log_Total_Bands_Rescaled = log(Total_Bands_Rescaled),
         Proportion = n/nrow(data_analyse)) %>% 
  select(Total_Bands, Total_Bands_Rescaled, Log_Total_Bands, 
         Log_Total_Bands_Rescaled, n, Proportion) %>% 
  kable(digits = 3, col.names = c("Total bands (x1000/ul)", 
                                  "Total bands (x100/ul)", 
                                  "log(Total bands [x1000/ul])", 
                                  "log(Total bands [x100/ul])", 
                                  "n", 
                                  "Proportion")) %>%
  kable_styling() %>% 
  add_header_above(header = c("Raw Values", "Transformations of Total Bands" = 3, 
                              "Frequency" = 2))
```

```{r}
#| label: fig-chop1_phase1_total_bands_rescaled_log_part
#| fig-cap: "Histogram (left) and boxplot (right) of the log(total bands [x100/ul]) with only total bands > 0"
par(mfrow=c(1,2))
hist(data_analyse$log_Total_Bands_Rescaled[data_analyse$log_Total_Bands_Rescaled>-3],
     xlab="log(Total bands [x100/ul])", xlim = c(0, 5),
     ylab = "Number of patients",
     main=NULL)
boxplot(data_analyse$log_Total_Bands_Rescaled[data_analyse$log_Total_Bands_Rescaled>-3],
        ylab="log(Total bands [x100/ul])", ylim = c(0,5))
par(mfrow=c(1,1))
```

```{r}
#| label: fig-chop1_phase1_total_bands_log_part
#| fig-cap: "Histogram (left) and boxplot (right) of the log(total bands [x1000/ul]) with only total bands > 0"
par(mfrow=c(1,2))
hist(data_analyse$log_Total_Bands[data_analyse$log_Total_Bands>-5],
     xlab="log(Total bands [x1000/ul])", xlim = c(-3, 2),
     ylab = "Number of patients",
     main=NULL)
boxplot(data_analyse$log_Total_Bands[data_analyse$log_Total_Bands>-3],
        ylab="log(Total bands [x1000/ul])", ylim = c(-3,2))
par(mfrow=c(1,1))
```

### Summary of each drug

@tbl-chop1_drug shows the number of times each drug is used. We can see that
Drug 1 is used far more than other drugs.

```{r}
#| label: tbl-chop1_drug
#| tbl-cap: Summary the number of times each drug is used

data_analyse %>% 
  mutate(Drug = factor(Drug, levels = 1:3, labels = paste("Drug", 1:3))) %>% 
  count(Drug) %>% 
  mutate(Proportion = n/nrow(data_analyse)) %>% 
  kbl(digits = 3) %>%
  kable_styling()
```

@fig-total_bands_rescaled_by_drugs shows the boxplot of 
**Total bands (x100/ul)**, categorized by drugs, while @fig-total_bands_by_drugs 
shows the boxplot of **Total bands (x1000/ul)**, categorized by drugs.

```{r}
#| label: fig-total_bands_rescaled_by_drugs
#| fig-cap: Boxplot of the total bands (x100/ul), categorized by drugs.

boxplot(data_analyse$Total_Bands_Rescaled ~ data_analyse$Drug,
        ylab="Total bands (x100/ul)",
        xlab = "Drug")
```

```{r}
#| label: fig-total_bands_by_drugs
#| fig-cap: Boxplot of the total bands (x1000/ul), categorized by drugs.

boxplot(data_analyse$Total_Bands ~ data_analyse$Drug,
        ylab="Total bands (x1000/ul)",
        xlab = "Drug")
```

@tbl-total_bands_by_drugs gives the mean and median value of the **Total bands**, 
categorized by drugs.

```{r}
#| label: tbl-total_bands_by_drugs
#| tbl-cap: Summary of total bands (x100/ul), categorized by drugs.

A=matrix(0,ncol=3,nrow=6)
for(i in 1:3){
  A[,i]=summary(data_analyse$Total_Bands_Rescaled[data_analyse$Drug==i])
}
colnames(A)=paste0("Drug ",1:3)
rownames(A)=c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max")
round(A,4)%>%   kbl() %>%   kable_styling()
```

@fig-total_bands_rescaled_by_drugs_log shows the boxplot of 
**log(Total bands) (x100/ul)**, categorized by drugs, where -Inf means total 
bands=0. Meanwhile, @fig-total_bands_by_drugs_log shows the boxplot of 
**log(Total bands) (x1000/ul)**, categorized by drugs, where -Inf means total 
bands=0. 

```{r}
#| label: fig-total_bands_rescaled_by_drugs_log
#| fig-cap: Boxplot of the log(total bands)(x100/ul), categorized by drugs.

boxplot(data_analyse$log_Total_Bands_Rescaled ~ data_analyse$Drug,
        xlab = "Drug",
        ylab="log(Total bands) (x100/ul)", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))

axis(2, at = axis_positions, labels = axis_labels)
```

```{r}
#| label: fig-total_bands_by_drugs_log
#| fig-cap: Boxplot of the log(total bands [x1000/ul]), categorized by drugs.

boxplot(data_analyse$log_Total_Bands ~ data_analyse$Drug,
        xlab = "Drug",
        ylab="log(Total bands [x1000/ul])", ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

::: {.callout-note}
@fig-total_bands_by_drugs_log above becomes Figure 2 in the manuscript. 
:::

@fig-total_bands_by_drugs_log_part shows the boxplot of 
**log(Total bands) (x100/ul)** with only Total bands \> 0, categorized by drugs.

```{r}
#| label: fig-total_bands_by_drugs_log_part
#| fig-cap: Boxplot of the log(total bands [x100/ul]), categorized by drugs.

boxplot(data_analyse$log_Total_Bands_Rescaled[data_analyse$log_Total_Bands_Rescaled>-3] ~ data_analyse$Drug[data_analyse$log_Total_Bands_Rescaled>-1],
        ylab="log(Total bands [x100/ul])",
        xlab = "Drug")

```

@tbl-total_bands_by_drugs_zero shows the number and proportion of instances 
where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_drugs_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each drug used.
#| 
data_analyse %>% 
  mutate(Drug = factor(Drug, levels = 1:3, labels = paste("Drug", 1:3)),
         Zero_Bands = Total_Bands == 0) %>% 
  group_by(Drug) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Drug", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

### Summary of total bands by visit time

#### Total

@tbl-total_bands_by_visit provides the mean and standard deviation of 
**Total bands** for each visit time. From this data, we can observe that the 
mean value of Total bands tends to decrease over time.

```{r}
#| label: tbl-total_bands_by_visit
#| tbl-cap: Summary of total bands (x100/ul) for each visit time.
Total_Bands_Rescaled_counts <- data_analyse %>%
  group_by(Visit_Number) %>%
   summarise(
    mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
    sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
    min = min(Total_Bands_Rescaled, na.rm = TRUE),
    max= max(Total_Bands_Rescaled, na.rm = TRUE)
  )
Total_Bands_Rescaled_counts%>%kable()%>%kable_styling() 
```

@fig-total_bands_by_visit shows the boxplot of total bands (x100/ul) for each 
visit time.

```{r}
#| label: fig-total_bands_by_visit
#| fig-cap: Boxplot of total bands (x100/ul) for each visit time.
boxplot(Total_Bands_Rescaled~Visit_Number,data=data_analyse,xlab="Visit number",ylab="Total bands (x100/ul)")
```

@fig-total_bands_rescaled_by_visit_log shows the boxplot of 
**log(total bands x100/ul)** for each visit time, where -Inf means total bands = 
0. @fig-total_bands_by_visit_log shows the boxplot of 
**log(total bands x1000/ul)** for each visit time, where -Inf means total bands = 
0.

```{r}
#| label: fig-total_bands_rescaled_by_visit_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time.
boxplot(log_Total_Bands_Rescaled ~ Visit_Number, data=data_analyse,
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])",ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

```{r}
#| label: fig-total_bands_by_visit_log
#| fig-cap: Boxplot of log(total bands [x1000/ul]) for each visit time.
boxplot(log_Total_Bands ~ Visit_Number, data=data_analyse,
        xlab="Visit number",
        ylab="log(Total bands [x1000/ul])",ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

::: {.callout-note}
@fig-total_bands_by_visit_log above becomes Figure 3a in the manuscript. 
:::


@fig-total_bands_by_visit_log_part shows the boxplot of log(total bands x100/ul)
for each visit time and only include those with total bands \> 0.

```{r}
#| label: fig-total_bands_by_visit_log_part
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time (only include those with total bands \> 0).
boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled > 0],
        data=data_analyse,
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_zero shows the number and proportion of instances 
where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time.

data_analyse %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

#### Drug 1

@tbl-total_bands_by_visit_drug1 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 1 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug1
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 1 at each visit time.

data_analyse %>% 
  filter(Drug==1) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_total_bands is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_total_bands
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 1"
boxplot(Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==1,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug1_log shows the boxplot of 
log(total bands x100/ul) for individuals using Drug 1 at each visit time, where 
-Inf means total bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1."

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug1_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 1 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log_part
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1 (only include those with total bands > 0)."

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=data_analyse[data_analyse$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug1_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug1_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 1.

data_analyse %>% 
  filter(Drug == 1) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

```{r}
#| label: drug
drug1 <- data_analyse %>% filter(Drug==1)
drug2 <- data_analyse %>% filter(Drug==2)
drug3 <- data_analyse %>% filter(Drug==3)
```

@fig-visit_times_of_drug_1 shows the times of using drug 1 at each visit number.

```{r}
#| label: fig-visit_times_of_drug_1
#| fig-cap: Times of using drug 1 at each visit number

drug_counts <- drug1 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 2  

@tbl-total_bands_by_visit_drug2 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 2 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug2
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 2 at each visit time.

data_analyse %>% 
  filter(Drug==2) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_drug_2 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_2
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 2"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==2,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug2_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug2_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log_part
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2 (only include those with total bands \> 0).

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=data_analyse[data_analyse$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug2_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug2_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 2.

data_analyse %>% 
  filter(Drug == 2) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

@fig-visit_times_of_drug_2 shows the times of using drug 2 at each visit number

```{r}
#| label: fig-visit_times_of_drug_2
#| fig-cap: Times of using drug 2 at each visit number

drug_counts <- drug2 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 3


@tbl-total_bands_by_visit_drug3 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 3 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug3
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 3 at each visit time.

data_analyse %>% 
  filter(Drug==3) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

@fig-boxplot_drug_3 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_3
#| #| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 3"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==3,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug3_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 3.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=data_analyse[data_analyse$Drug==3,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug3_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log_part
#| fig-cap: Boxplot of log(total bands x100/ul) for each visit time for those using drug 3.
boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled>0]~Visit_Number[Total_Bands_Rescaled>0],data=data_analyse[data_analyse$Drug==3,],xlab="Visit number",ylab="log(Total bands) (x100/ul)")
```

@tbl-total_bands_by_visit_drug3_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug3_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 3.

data_analyse %>% 
  filter(Drug == 3) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

@fig-visit_times_of_drug_3 shows the times of using drug 3 at each visit number

```{r}
#| label: fig-visit_times_of_drug_3
#| fig-cap: Times of using drug 3 at each visit number

drug_counts <- drug3 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

### Additional values needed  
#### Total white blood cells (WBC) 

@tbl-chop1_WBC_phase1 gives the mean and median value of the 
**Total WBC (x1000/uL)**.  

```{r}
#| label: tbl-chop1_WBC_phase1
#| tbl-cap: Summary of total WBC (x1000/uL)
A=summary(data_analyse$Total_WBC)
round(as.matrix(A),4) %>%
  kbl() %>%
  kable_styling()
```

@tbl-total_WBC_by_visit provides the mean and standard deviation of 
**Total WBC** for each visit time. From this data, we can observe that the mean 
value of Total bands tends to decrease over time

```{r}
#| label: tbl-total_WBC_by_visit
#| tbl-cap: Summary of total white blood cells (x1000/uL) for each visit time.
data_analyse %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_WBC, na.rm = TRUE),4),
            sd= round(sd(Total_WBC, na.rm = TRUE),4),
            min = min(Total_WBC, na.rm = TRUE),
            max= max(Total_WBC, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```


#### neutrophils (neuts)
@tbl-chop1_neutrophils_phase1 gives the mean and median value of the 
**Total neutrophils (x10e3/uL)**. 

```{r}
#| label: tbl-chop1_neutrophils_phase1
#| tbl-cap: Summary of total neutrophils 

A=summary(data_analyse$Total_Neutrophil)
round(as.matrix(A),4)%>%
  kbl() %>%
  kable_styling()
```

@tbl-total_neutrophil_by_visit provides the mean and standard deviation of 
**Total neutrophil** for each visit time. 

```{r}
#| label: tbl-total_neutrophil_by_visit
#| tbl-cap: Summary of total neutrophil for each visit time.

data_analyse %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Neutrophil, na.rm = TRUE),4),
            sd= round(sd(Total_Neutrophil, na.rm = TRUE),4),
            min = min(Total_Neutrophil, na.rm = TRUE),
            max= max(Total_Neutrophil, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```

### Bandemia grade  

```{r}
A=matrix(0,nrow = 6,ncol=4)
colnames(A)=c("Bandemia grade","Value according to MSUVDL reference range (k/uL)","Numbers","Proportion")
A[,1]=c(0:5)
A[,2]=c("0.0-0.10","0.11-0.29","0.30-0.59","0.60-1.00",">1.00","death")
A[,3]=c(sum(data_analyse$Total_Bands_Rescaled<=1),
        sum(data_analyse$Total_Bands_Rescaled>1&data_analyse$Total_Bands_Rescaled<3),
        sum(data_analyse$Total_Bands_Rescaled>=3&data_analyse$Total_Bands_Rescaled<6),
        sum(data_analyse$Total_Bands_Rescaled>=6&data_analyse$Total_Bands_Rescaled<=10),
        sum(data_analyse$Total_Bands_Rescaled>10),0)
A=as.data.frame(A)
A$Numbers=as.numeric(A$Numbers)
A[,4]=A[,3]/609
A %>% 
  kable(digits = 3) %>%
  kable_styling()
```


### Model that use total bands as the outcome

#### "Generalized Linear Mixed Model" (GLMM) with a Negative Binomial link function

**Poisson regression** is a statistical model designed for analyzing count data,
specifically when dealing with non-negative integer counts that are expected to
follow a Poisson distribution. It's most effectively employed in scenarios
involving event frequencies, such as the number of occurrences of a particular
event. The primary objective of Poisson regression is to establish the
relationship between a dependent variable, which represents these counts, and
one or more independent variables. These independent variables can be either
categorical or continuous, allowing for a wide range of applications in various
fields.

So we considered to use Poisson regression here.

A key characteristic of Poisson regression is that it assumes the mean and
variance of the distribution are equal, a property known as equidispersion.

First let's exam it's mean and variance.

```{r}
#| label: mean_var
print(round(c("mean"=mean(data_analyse$Total_Bands_Rescaled),"var"=var(data_analyse$Total_Bands_Rescaled)),4))
```

Given that the sample data exhibit a mean of $1.823$, which is substantially
lower than the observed variance of $44.995$, the use of a Negative Binomial
link function should be justified in this context. This significant discrepancy
suggests overdispersion, a scenario where the Negative Binomial distribution is
particularly suitable due to its flexibility in handling extra-Poisson
variation.

Additionally, considering the use of a zero-inflated model might be appropriate,
as more than 75% of the values are zeros.

Next we check the correlation between the surface area and the weight.

```{r}
#| label: correlation
data_Info=data_analyse[1,]
for(i in 2:length(data_analyse$Patient_Number)){
  if(data_analyse$Patient_Number[i]!=data_analyse$Patient_Number[i-1]){
      data_Info=rbind(data_Info,data_analyse[i,])
  }
}
print(round(cor(data_Info$Surface_Area_m2,data_Info$Weight_kg),4))
```

The correlation is 0.99. Therefore, we include only one of them in the model to
avoid collinearity.

@fig-chop1_m2 shows the distribution of patients' surface area. It's nearly 
normal distribution.

```{r}
#| label: fig-chop1_m2
#| fig-cap: Histogram of surface area (in CHOP 1)

ggplot(data_Info, aes(x=Surface_Area_m2)) + 
  geom_histogram(binwidth = 0.1, colour="black", fill="white") + 
  ggtitle("Histogram of m^2") + 
  xlab("m^2") + 
  ylab("Frequency") + 
  xlim(0, 2)
```

@tbl-chop1_m2 provides the mean and median of the surface area

```{r}
#| label: tbl-chop1_m2
#| tbl-cap: Summary of patient's surface area (in CHOP 1)

A=summary(data_Info$Surface_Area_m2)
round(as.matrix(A),4)%>%   kbl() %>%   kable_styling() 
```

```{r}
#| label: model_regression
#colnames(data_analyse)[c(1,3,10,13,14)]=c("Patient_number","Visit_number","Total_Bands_Rescaled","Weight","Surface_area")
data_analyse$Drug=as.factor(data_analyse$Drug)
data_analyse$Surface_area=data_analyse$Surface_Area_m2-mean(data_Info$Surface_Area_m2)
data_analyse$Visit_number=data_analyse$Visit_Number-1
model_nb <- glmmTMB(Total_Bands_Rescaled ~ Drug +Surface_area+ Visit_number+(1 | Patient_Number),
                      data=data_analyse,
                      ziformula=~1,
                      family=nbinom2)
```

#### Results

Below shows the regression results of the model.

```{r}
#| label: summary_model
summary(model_nb)
```

Note: In this analysis, we employ the NB2 parameterization, which assumes that 
$\sigma^2 = \mu + \alpha\mu^2$, where $\mu$ is the mean, and $\sigma^2$ is the 
variance. The model results yield an estimated value of $\hat{\alpha} = 0.376$.

@fig-EMMs_with_CI_log provides the estimated marginal means of 
log(Total bands)(x100/ul).

```{r}
#| label: fig-EMMs_with_CI_log
#| fig-cap: Estimated Marginal Means of log(Total bands)(x100/ul)
emm <- emmeans(model_nb, ~ Drug)
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = emmean)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Means: log(Total bands)(x100/ul)") +
  ggtitle("EMMs with Confidence Intervals")


```

The estimated mean for each drug is calculated assuming the Surface_area equals 
the mean of Surface_area ($0.9470$), and the visit number is 1.

```{r}
#| label: tbl-log_total_bands
#| tbl-cap: Estimated Means of log(Total bands)(x100/ul) for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)
emm_df%>%kable()%>%kable_styling()
```

```{r}
#| label: EMMs_with_CI
#| fig-cap: Estimated Marginal Means of Total bands (x100/ul)
emm <- emmeans(model_nb, ~ Drug,type="response")
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = response)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Means: Total bands (x100/ul)") +
  ggtitle("EMMs with Confidence Intervals")
```

```{r}
#| label: tbl-total_bands
#| tbl-cap: Estimated Means of Total bands (x100/ul) for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)
emm_df%>%kable()%>%kable_styling()
```

@fig-EMMs_with_CI_line shows the estimated results from the model, with the 
red line representing the threshold $Total\_bands = 200/uL$ for the incidence of 
Bandemia

```{r}
#| label: fig-EMMs_with_CI_line
#| fig-cap: Estimated Marginal Means of Total bands (x100/ul)
emm <- emmeans(model_nb, ~ Drug+Visit_number,type="response")
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = response)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "red") + 
  xlab("Drug") +
  ylab("Estimated Marginal Means: Total bands (x100/ul)") +
  ggtitle("EMMs with Confidence Intervals")
```

In @tbl-rate_ratios, the rate ratios are presented. These are measures used to
compare the rates of occurrence of an event across different groups or under
different conditions. Also included are the corresponding confidence intervals
for these comparisons between the drugs.

```{r}
#| label: tbl-rate_ratios
#| tbl-cap: Rate ratios between the drugs
emm <- emmeans(model_nb, ~ Drug)

pairwise <- pairs(emm, simple = "each", by = NULL)
contrast <- contrast(emm, method = "revpairwise")

rr <- summary(contrast, infer = c(TRUE, TRUE), type = "response")
rr=rr[,-4]
rr=rr[,-6]
colnames(rr)=c("contrast","rate ratio","SE","lower CI","upper CI","z.ratio","p.value")
rr[,c(2:7)]=round(rr[,c(2:7)],4)

rr%>%kable()%>%kable_styling()
```

### Model that use total bands as the outcome (considering the interaction between surface area and drug type)

```{r}
#| label: model_regression_interaction
model_nb_interact <- glmmTMB(Total_Bands_Rescaled ~ Drug +Surface_area+ Visit_number+Drug*Surface_area+(1 | Patient_Number),
                      data=data_analyse,
                      ziformula=~1,
                      family=nbinom2)
```

#### Results

Below shows the regression results of the model.

```{r}
#| label: summary_model_2
summary(model_nb_interact)
```

Note: In this analysis, we employ the NB2 parameterization, which assumes that 
$\sigma^2 = \mu + \alpha\mu^2$, where $\mu$ is the mean, and $\sigma^2$ is the 
variance. The model results yield an estimated value of $\hat{\alpha} = 0.499$.

After including the interaction between surface area and drug type in the model,
we found that these interactions do not significantly impact the outcome.

We further use a nested model comparison to test the overall interaction effect.

```{r}
#| label: compare_anova
anova(model_nb, model_nb_interact)
```

Based on the testing results, including the interaction effect does not
significantly improve the model's fit to the data compared to the model without
the interaction term. This suggests that the interaction effect is not
statistically significant (pvalue=0.5332). Therefore, we can focus on
interpreting the model without the interaction effect.

## Bandemia

@tbl-chop1_Bandemia displays the proportion of Bandemia incidents among all the
analyzed samples. Bandemia is defined as a condition where the Total bands
(x10.e3/uL) are equal to or higher than 0.2.

```{r}
#| label: tbl-chop1_Bandemia
#| tbl-cap: Incidence of Bandemia
#data_analyse$Visit_number=data_analyse$Visit_number+1
#data_analyse$Total_bands=data_analyse$Total_bands/10

A=summary(as.factor(data_analyse$Bandemia))
A=as.matrix(A)
A_drop=apply(A, MARGIN = 2, function(x) {
  total <- sum(x)
  prop <- x / total * 100
  return(paste(x, " (", sprintf("%.1f%%", prop), ")", sep=""))
})
colnames(A_drop)=c("Times (%)")
rownames(A_drop)=c("No","Yes")
A_drop %>% 
  kbl() %>%
  kable_styling()
```

@tbl-chop1_Bandemia_each_drug displays the proportion of Bandemia incidents
among all the analyzed samples, categorized by drugs.

```{r}
#| label: tbl-chop1_Bandemia_each_drug
#| tbl-cap: Results of each drug
A=table(data_analyse$Drug, data_analyse$Bandemia)
A=as.matrix(A)
colnames(A)=c("No","Yes")
rownames(A)=c(paste0("Drug ",1:3))
A_prop <- apply(A, MARGIN = 1, function(x) {
  total <- sum(x)
  prop <- x / total * 100
  return(paste(x, " (", sprintf("%.1f%%", prop), ")", sep=""))
})
df_A_prop <- as.data.frame(t(A_prop))
colnames(df_A_prop)=c("No","Yes")
rownames(df_A_prop)=c(paste0("Drug ",1:3))
df_A_prop %>% 
  kbl() %>% 
  kable_styling()
```


@tbl-total_bandamia_by_visit provides the proportion of **bandemia** for each
visit time. From this, we can observe that the mean value of **bandemia** tends
to decrease over time.

```{r}
#| label: tbl-total_bandamia_by_visit
#| tbl-cap: Summary of total bands for each visit time.

total_bandemia_counts <- data_analyse %>%
  group_by(Visit_Number) %>%
  summarise(Proportion = round(mean(Bandemia, na.rm = TRUE),4)) 

total_bandemia_counts %>% 
  kable() %>%
  kable_styling()
```

```{r}
#| label: fig-Proportion_Visit_number
#| fig-cap: Proportion of Bandemia at each visit number
barplot(Proportion~Visit_Number, data=total_bandemia_counts,
        xlab="Visit number",
        ylim=c(0,0.5))
```

::: {.callout-note}
@fig-Proportion_Visit_number above becomes Figure 3b in the manuscript. 
:::

@tbl-total_bandamia_by_visit_drug1 provides the proportion of **bandemia** for
individuals using Drug 1 at each visit time. From this, we can observe that the
mean value of **bandemia** tends to decrease over time.

```{r}
#| label: tbl-total_bandamia_by_visit_drug1
#| tbl-cap: Summary of total bands for those using drug 1 at each visit time.

total_bandemia_counts_drug1 <- data_analyse %>% 
  filter(Drug==1) %>%
  group_by(Visit_Number) %>%
  summarise(Proportion = round(mean(Bandemia, na.rm = TRUE),4))

total_bandemia_counts_drug1 %>%
  kable() %>% 
  kable_styling()
```

```{r}
#| label: fig-Proportion_Visit_number_drug1
#| fig-cap: Proportion of Bandemia at each visit number for those using drug 1
barplot(Proportion~Visit_Number, data=total_bandemia_counts_drug1,
        xlab="Visit number",
        ylim=c(0,0.5))
```

@tbl-total_bandemia_by_visit_drug2 provides the proportion of **bandemia** for 
individuals using Drug 2 at each visit time

```{r}
#| label: tbl-total_bandemia_by_visit_drug2
#| tbl-cap: Summary of total bands for those using drug 2 at each visit time.

total_bandemia_counts_drug2 <- data_analyse %>% 
  filter(Drug==2) %>%
  group_by(Visit_Number) %>%
  summarise(Proportion = round(mean(Bandemia, na.rm = TRUE),4))

total_bandemia_counts_drug2 %>% 
  kable() %>% 
  kable_styling()
```


```{r}
#| label: Proportion_Visit_number_drug2
#| fig-cap: Proportion of Bandemia at each visit number for those using drug 2

barplot(Proportion~Visit_Number, data=total_bandemia_counts_drug2,
        xlab="Visit number",
        ylim=c(0,0.5))
```

@tbl-total_bandamia_by_visit_drug3 provides the proportion of **bandemia** for 
individuals using Drug 3 at each visit time

```{r}
#| label: tbl-total_bandamia_by_visit_drug3
#| tbl-cap: Summary of total bands for those using drug 3 at each visit time.

total_bandemia_counts_drug3 <- data_analyse %>% 
 filter(Drug==3) %>%
  group_by(Visit_Number) %>%
  summarise(Proportion = round(mean(Bandemia, na.rm = TRUE),4))

total_bandemia_counts_drug3 %>% 
  kable() %>% 
  kable_styling()
```

```{r}
#| label: Proportion_Visit_number_drug3
#| fig-cap: Proportion of Bandemia at each visit number for those using drug 3

barplot(Proportion~Visit_Number, data=total_bandemia_counts_drug3,
        xlab="Visit number",
        ylim=c(0,0.5))
```

### Model that use incidence of Bendemia as the outcome

#### Logistic regression with random effect

Due to the high correlation between weight and surface area, I have included
only **Surface area** in the model. Additionally, as the outcome variable is
binary, logistic regression was applied.

**Logistic regression** is a widely used statistical method for modeling the 
probability of a binary outcome based on one or more predictor variables. It's
particularly applicable when the response variable is categorical and binary in
nature, meaning it has two possible outcomes such as 'yes' or 'no', 'success' or
'failure', 'presence' or 'absence', etc.

Logistic regression is suitable in situations where you need to model the impact
of several independent variables (which can be either categorical or continuous)
on a binary dependent variable. Common applications include medical fields for
disease diagnosis, in finance for credit scoring, in marketing for predicting
customer churn, and in many other fields where binary outcomes are to be
predicted.

```{r}
#|label: incidence_m2

model <- glmer(Bandemia ~ Drug + Surface_area + Visit_number + (1 | Patient_Number), 
               data = data_analyse, 
               family = binomial(link = "logit"))
```

#### Results

```{r}
#|label: summary_model_logistic
summary(model)
```

In the model, Drug 1 is set as the baseline. Drug 2 is found to significantly
decrease the probability of causing Bandemia compared to the baseline (Drug 1).

@fig-EMMs_with_CI_logit shows the estimated means of logit(Bandemia) for each
drug. And it is calculated assuming the Surface_area equals the mean of
Surface_area ($0.9470$), and the visit number is 1.

```{r}
#| label: fig-EMMs_with_CI_logit
#| fig-cap: Estimated Marginal Means of logit(Bandemia) for each drug

emm <- emmeans(model, ~ Drug)
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = emmean)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Means:logit(Bandemia)") +
  ggtitle("EMMs with Confidence Intervals")
```

@tbl-logit_bandemia lists the values used to generate the figure above.

```{r}
#| label: tbl-logit_bandemia
#| tbl-cap: Estimated Means of logit(bandemia) for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)

emm_df%>%kable()%>%kable_styling()
```

@fig-Estimated_Marginal_Proportions shows the estimated means of proportion of
Bandemia for each drug. And it is calculated assuming the Surface_area equals
the mean of Surface_area ($0.9470$), and the visit number is 1.

```{r}
#| label: fig-Estimated_Marginal_Proportions
#| fig-cap: Estimated Marginal Proportions of Bandemia
emm <- emmeans(model, ~ Drug,type="response")
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = prob)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Proportions: Bandemia") +
  ggtitle("Estimated values with Confidence Intervals")
```

@tbl-bandemia lists the values used to generate the figure above.

```{r}
#| label: tbl-bandemia
#| tbl-cap: Estimated Proportions of Bandemia for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)
emm_df%>%kable()%>%kable_styling()
```

@tbl-odds_ratio presents the odds ratios comparing between the drugs. The odds
ratio of 3.2610 indicates that Drug 1 is significant (pvalue=0.003<0.5)
associated with a higher likelihood of the outcome compared to Drug 2, while
Drug 2 has an odds ratio of 0.2987, suggesting a significant (pvalue=0.018<0.5)
lower likelihood of the outcome as compared to Drug 3.

```{r}
#| label: tbl-odds_ratio
#| tbl-cap: odds ratios comparing between the drugs
emm <- emmeans(model, ~ Drug)

pairwise <- pairs(emm)
contrast <- contrast(emm, method = "revpairwise")

rr <- summary(pairwise, infer = c(TRUE, TRUE), type = "response")
rr=rr[,-4]
rr=rr[,-6]
colnames(rr)=c("contrast","rate ratio","SE","lower CI","upper CI","z.ratio","p.value")
rr[,c(2:7)]=round(rr[,c(2:7)],4)

rr%>%kable()%>%kable_styling()
```

We get similar results from these two models. In summary, the likelihood of
Bendemia occurrence with drug 2 is significantly lower compared to its
occurrence with drugs 1 and 3. Additionally, a negative association was found
between surface area and the total bands value, with this association being
statistically significant.

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
