---
title: Descriptive Analysis of CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
bibliography: references.bib
csl: apa.csl
params:
  SourceFile: "Descriptive_Analysis.qmd"
  LogFile: "Descriptive_Analysis.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    output-file: "Descriptive_Analysis_Draft.html"
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
---

# Purpose
This file is part of a research compendium [@Pierce-RN8714]. It provides some
descriptive analyses of data provided by Alison Masyr and Caitlin Eliason that
contains observations of dogs receiving CHOP chemotherapy at MSU Veterinary
Medicine Clinic (VMC) between 2016 and 2022. Some results here are used in a 
manuscript [@Eliason-RN8603].

# Setup
This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Results.

## Define Global Options
Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages
Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(devtools)     # for session_info()
library(here)       # for here(), i_am()
library(dplyr)      # for filter(), group_by(), mutate(), rename(), summarise()
library(forcats)    # for fct_recode()
library(tibble)     # for rowid_to_column()
library(tidyr)      # for pivot_longer()
library(ggplot2)    # for ggplot()
library(ggpmisc)    # for stat_correlation()
library(patchwork)  # for plot_layout()
library(knitr)      # for kable()
library(kableExtra) # for kable_styling 
library(readxl)     # for read_xlsx()
library(piercer)    # for git_report()
library(quarto)     # for quarto_version()
library(CHOPStudy)  # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Descriptive_Analysis.qmd")
```

## Load Data {#sec-Load-Data}

::: {.callout-caution}
Render the `scripts/Import_Data.qmd` file to create the data file loaded below. 
:::

@tbl-imported-data-file shows meta-data about the data file we will load.

```{r}
#| label: tbl-imported-data-file
#| tbl-cap: "Meta-Data About the Imported Data File Loaded"

DataFile <- here("data/Imported_CHOP_Data.RData")

file_details(DataFile) %>% 
  kable(, format = "html", 
        col.names = c("File Name", "Size", "Last Modified")) %>% 
  kable_styling() 
```

```{r}
#| label: load-data

load(file=here("data/Imported_CHOP_Data.RData"))
```

# Results {#sec-Results}

## Patient-Level Summaries
@tbl-patient-cvv-missing and @tbl-patient-var-missing show summaries of missing 
data in the patient-level dataset. 

```{r}
#| label: tbl-patient-cvv-missing
#| tbl-cap: "Missingness Overview for Patient-Level Dataset (`Final_Patients`): 
#|          Cases, Variables, and Values"

Final_Patients %>% 
  cvv_missingness() %>% 
  kable(format = "html", digits = 1,
          col.names = c("Subset", rep(c("n", "%"), times = 3))) %>%
    kable_styling() %>%
    add_header_above(header = c(" " = 1, "Cases" = 2, "Variables" = 2, "Values" = 2))
```

```{r}
#| label: tbl-patient-var-missing
#| tbl-cap: Misssingness Assessment for Variables in Patient-Level Dataset 
#|          (`Final_Patients`)

Final_Patients %>% 
var_missingness(.) %>% 
  kable(format = "html", digits = 1) %>%
  kable_styling() %>%
  add_header_above(header = c(" " = 3, "Observed Values" = 2, 
                              "Missing Values" = 2))
```

@tbl-patient-describe shows patient-level descriptive statistics for weight 
surface area, and number of visits, while @tbl-patient-Leukemic gives the 
univariate frequency distribution for leukemic status, which was a 
time-invariant patient-level characteristic within the CHOP cycle.

```{r}
#| label: tbl-patient-describe
#| tbl-cap: !expr paste0("Descriptive Statistics for Patient-Level Variables ", 
#|                       "(N = ", nrow(Final_Patients), " Patients)")
#| message: false

VNames <- c("Weight_kg", "Surface_Area_m2", "N_Visits")

CNames <- c("Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", "Q75", "Max")

Final_Patients %>% 
  select(Weight_kg, Surface_Area_m2, N_Visits) %>% 
  pivot_longer(cols = c(Weight_kg, Surface_Area_m2, N_Visits), 
               names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  kable(., format = "html", digits = 2, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T)
```

```{r}
#| label: weight

N_Weight_LT10 <- Final_Patients %>% filter(Weight_kg < 10) %>% nrow()
N_Weight_1015 <- Final_Patients %>% filter(between(Weight_kg, left = 10, right = 15)) %>% nrow()

Pct_Weight_LT10 <- 100*N_Weight_LT10/nrow(Final_Patients)
Pct_Weight_1015 <- 100*N_Weight_1015/nrow(Final_Patients)
```

:::{.callout-tip}
Below are specific statistics about numbers and percentage of patients at the 
very low end of weight distribution that we use in the manuscript.

* `r N_Weight_LT10` (`r round(Pct_Weight_LT10, digits = 0)`%) weighed less than 
  10 kg.
* `r N_Weight_1015` (`r round(Pct_Weight_1015, digits = 0)`%) weighed 10-15 kg.

:::

```{r}
#| label: tbl-patient-Leukemic
#| tbl-cap: !expr paste0("Frequency Distribution for Patient-Level Leukemic ", 
#|                       " Status (N = ", nrow(Final_Patients), " Patients)")
#| message: false

Final_Patients %>% 
  xtabs(~Leukemic, addNA = TRUE, data = .) %>% 
  as.data.frame() %>% 
  mutate(Percent = 100*Freq/nrow(Final_Patients)) %>% 
  kable(., format = "html", digits = 0) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T)
```

## Visit-Level Summaries
@tbl-visit-cvv-missing and @tbl-visit-var-missing show summaries of missing 
data in the visit-level dataset. 

```{r}
#| label: tbl-visit-cvv-missing
#| tbl-cap: "Missingness Overview for Visit-Level Dataset (`Final_visits`): 
#|          Cases, Variables, and Values"

Final_Visits %>% 
  cvv_missingness() %>% 
  kable(format = "html", digits = 1,
          col.names = c("Subset", rep(c("n", "%"), times = 3))) %>%
    kable_styling() %>%
    add_header_above(header = c(" " = 1, "Cases" = 2, "Variables" = 2, "Values" = 2))
```

```{r}
#| label: tbl-visit-var-missing
#| tbl-cap: Misssingness Assessment for Variables in Visit-Level Dataset 
#|          (`Final_Visits`)

Final_Visits %>% 
var_missingness(.) %>% 
  #arrange(desc(Pct_Missing)) %>%   # Sort by % missing
  kable(format = "html", digits = 1) %>%
  kable_styling() %>%
  add_header_above(header = c(" " = 3, "Observed Values" = 2, 
                              "Missing Values" = 2))
```

@tbl-visit-describe-cat shows univariate frequency distributions for binary 
cateogrical biomarker variables.

```{r}
#| label: tbl-visit-describe-cat
#| tbl-cap: !expr paste0("Descriptive Statistics for Categorical Biomarker ", 
#|                       "Variables (N = ", nrow(Final_Visits), " Visits)")
#| message: false

VNames <- c("Bands_Present", "Bandemia", "High_BN_Ratio", "Toxic_Change")

Final_Visits %>% 
  select(Bands_Present, Bandemia, High_BN_Ratio, Toxic_Change) %>% 
  pivot_longer(cols = c(Bands_Present, Bandemia, High_BN_Ratio, Toxic_Change), 
               names_to = "Variable", values_to = "Value") %>% 
  group_by(Variable, Value, .drop = FALSE) %>% 
  summarise(N = n()) %>% 
  mutate(Variable = factor(Variable, levels = VNames),
         Value = factor(Value, levels=0:1, labels = c("No", "Yes")),
         Percent = 100*(N/nrow(Final_Visits))) %>% 
  arrange(Variable, desc(Value)) %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  kable(., format = "html", digits = 1, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  collapse_rows(., columns = 1, valign = "top", 
                row_group_label_position = "stack") 
```

@tbl-visit-band-category shows the frequency distribution for 3-level ordinal 
band category variable. 

```{r}
#| label: tbl-visit-band-category
#| tbl-cap: !expr paste0("Frequency Distribution for Band Category ", 
#|                       "(N = ", nrow(Final_Visits), " Visits)")
#| message: false

Final_Visits %>% 
  count(Band_Category) %>% 
  rename(N_Visits = n) %>% 
  mutate(Percent = 100*(N_Visits/nrow(Final_Visits))) %>% 
  kbl(digits = 1) %>%
  kable_styling()
```

@tbl-visit-bands-toxic shows the crosstabulation of whether bands were present 
and whether toxic change was observed in the CBC sample in the visit-level data. 
This was initially used to investigate missing data issues that we have 
subsequently fixed. 

```{r}
#| label: tbl-visit-bands-toxic
#| tbl-cap: !expr paste0("Crosstab of Bands Present by Toxic Change ", 
#|                       "(N = ", nrow(Final_Visits), " Visits)")

Final_Visits %>% 
  mutate(Bands_Present = factor(Bands_Present, levels = 0:1, 
                                labels = c("No", "Yes")),
         Toxic_Change = factor(Toxic_Change, levels = 0:1, 
                                labels = c("No", "Yes"))) %>% 
  xtabs(~Toxic_Change + Bands_Present , addNA = TRUE, data = .) %>% 
  addmargins() %>% 
  as.data.frame() %>% 
  pivot_wider(names_from = Bands_Present, values_from = Freq) %>% 
  kable(., format = "html", digits = 1, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  row_spec(row = 3, italic = TRUE) %>% 
  column_spec(column = 4, italic = TRUE) %>% 
  add_header_above(header = c(" ", "Bands_Present" = 2, " "))
```

:::{.callout-tip}
@tbl-visit-bands-toxic shows that all of the instances of toxic change are for 
CBC samples where bands were present. Per Dr. Masyr, toxic change is a 
description given to mature neutrophils, not bands (though both can be 
simultaneously present). 
:::

@tbl-visit-describe-cont shows visit-level descriptive statistics for continuous 
biomarker variables, while @tbl-visit-describe-cont-by-Bandemia repeats the same 
summary broken down by whether banedmia was absent versus present at the time of 
the CBC sample. 

```{r}
#| label: tbl-visit-describe-cont
#| tbl-cap: !expr paste0("Descriptive Statistics for Continuous Biomarker ", 
#|                       "Variables (N = ", nrow(Final_Visits), " Visits)")
#| message: false

VNames <- c("Total_WBC", "Total_Neutrophil", "Total_Bands", "BN_Ratio")
VNamesUnits <- c("Total_WBC (k/ul)", 
                 "Total_Neutrophil (k/ul)", 
                 "Total_Bands (k/ul)", 
                 "BN_Ratio")

CNames <- c("Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", "Q75", "Max")

Final_Visits %>% 
  select(Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio) %>% 
  pivot_longer(cols = c(Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio), 
               names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Variable = factor(Variable, levels = VNames, labels = VNamesUnits)) %>% 
  kable(., format = "html", digits = 1, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) 
```

```{r}
#| label: fig-visit-bands-neuts
#| fig-cap: !expr paste0("Scatterplot of Total Bands and Total Neutrophils ", 
#|                       "With Linear and Loess Smoothing Curves ",
#|                       "(N = ", nrow(Final_Visits), " Visits)")
#| fig-width: 4
#| fig-height: 2.64
#| warning: false

Fig1 <- Final_Visits %>% 
  ggplot(., aes(x = Total_Neutrophil, y = Total_Bands)) + 
  geom_point() +
  stat_correlation(method = "pearson", small.r = TRUE, r.conf.level = 0.95, 
                   use_label("R", "R.CI", "n")) +
  geom_smooth(method="loess", se = TRUE, fullrange = FALSE, 
              level = 0.95, linetype = "dashed", col = "red") +
  geom_smooth(method="lm", se = TRUE, fullrange = FALSE, level = 0.95, 
              linetype = "solid", col = "red") +
  theme_bw() +
  coord_cartesian(xlim = c(0, 50), ylim = c(0, 7), ) +
  scale_x_continuous(breaks = seq(from = 0, to = 50, by = 10)) +
  scale_y_continuous(breaks = seq(from = 0, to = 7, by = 1)) +
  xlab("Total neutrophils (k/ul)") +
  ylab("Total bands (k/ul)") +
  annotate(geom = "text", x =45, y = 4.6, label = "Loess", col = "red") +
  annotate(geom = "text", x =45, y = 2.5, label = "Linear", col = "red") 

Fig1

# Export figure for use in journal submission
ggsave(filename = here(paste0("scripts/output/Figure_1.png")), plot = Fig1, 
       device = "png", width = 4, height = 2.64, units = "in", dpi = 600)
```

::: {.callout-tip}
@fig-visit-bands-neuts will be Figure 1 in the manuscript we are preparing
[@Eliason-RN8603]. Note that the Pearson correlation is specifically a measure 
of linear association. 
:::

```{r}
#| label: tbl-visit-describe-cont-by-Bandemia
#| tbl-cap: !expr paste0("Descriptive Statistics for Continuous Biomarker ", 
#|                       "Variables by Bandemia (N = ", nrow(Final_Visits), 
#|                       " Visits)")
#| message: false

VNames <- c("Total_WBC", "Total_Neutrophil", "Total_Bands", "BN_Ratio")
VNamesUnits <- c("Total_WBC (k/ul)", 
                 "Total_Neutrophil (k/ul)", 
                 "Total_Bands (k/ul)", 
                 "BN_Ratio")

CNames <- c("Bandemia", "Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", 
            "Q75", "Max")

Final_Visits %>% 
  select(Bandemia, Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio) %>% 
  pivot_longer(cols = c(Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio), 
               names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Bandemia, Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Bandemia = factor(Bandemia, levels = 0:1, labels = c("No", "Yes")),
         Variable = factor(Variable, levels = VNames, labels = VNamesUnits)) %>% 
  kable(., format = "html", digits = 1, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  collapse_rows(., columns = 1, valign = "top", row_group_label_position = "stack") 
```

@tbl-drug-freq shows the univariate frequency distribution for drug (i.e., the 
number of times each drug was used). 

```{r}
#| label: tbl-drug-freq
#| tbl-cap: !expr paste0("Frequency Distribution for Drug Used (N = ", 
#|                       nrow(Final_Visits), " Visits)")

DrugNames <- c("Vincristine", "Cyclophosphamide", "Doxorubicin")

Final_Visits %>% 
  count(Drug) %>% 
  rename(N_Visits = n) %>% 
  mutate(Name = factor(Drug, levels = 1:3, labels = DrugNames), 
         Percent = 100*(N_Visits/nrow(Final_Visits))) %>% 
  select(Drug, Name, N_Visits, Percent) %>% 
  kbl(digits = 1) %>%
  kable_styling()
```

@tbl-visit-freq shows the univariate frequency distribution for visit number. 

```{r}
#| label: tbl-visit-freq
#| tbl-cap: !expr paste0("Frequency Distribution for Visit Number (N = ", 
#|                       nrow(Final_Visits), " Visits)")

Final_Visits %>% 
  count(Visit_Number) %>% 
  rename(N_Obs = n) %>% 
  mutate(Percent = 100*(N_Obs/nrow(Final_Visits))) %>% 
  select(Visit_Number, N_Obs, Percent) %>% 
  kbl(digits = 1) %>%
  kable_styling()
```

## Total Bands
@tbl-freq-bands shows the frequency distribution for total bands in both its 
original units (x10.e3/uL=k/uL) and log-transformed units. It highlights the 
large proportion of zero values. 

```{r}
#| label: tbl-freq-bands
#| tbl-cap: "Frequency Distribution of Total Bands for All Visits"

Final_Visits %>% 
  count(Total_Bands) %>% 
  mutate(Log_Total_Bands = log(Total_Bands), 
         Proportion = n/nrow(Final_Visits)) %>% 
  select(Total_Bands, Log_Total_Bands, n, Proportion) %>% 
  kable(digits = 3, col.names = c("Total Bands (k/ul)", 
                                  "log(Total Bands [k/ul])", 
                                  "n", 
                                  "Proportion")) %>%
  kable_styling() %>% 
  add_header_above(header = c("Raw Values", "Transformed Values" = 1, 
                              "Frequency" = 2))
```

@fig-bands shows the distribution total bands in the original units via a
histogram. The heavy concentration of zero values makes it hard to see the shape 
of the distribution among the non-zero values.

```{r}
#| label: fig-bands
#| fig-cap: Distribution of Total Bands (k/ul) for All Visits.
#| fig-width: 6
#| fig-height: 3
#| message: false

ggplot(Final_Visits, aes(x = Total_Bands)) + 
  geom_histogram(aes(y=after_stat(count)), colour="black", fill="grey70") +
  theme_bw() +
  scale_x_continuous(limits = c(-0.2,7), n.breaks = 8, oob = function(x, limits) x) +
  scale_y_continuous(limits=c(-0.2, 500)) +
  xlab("Total Bands (k/ul)") +
  ylab("No. of Patients")
```

@fig-log-bands shows the distribution of log-transformed total bands via both a
histogram and a boxplot with jittered data points. It includes data from all 
visits. In both graphs, the log equals -Infinity when total bands are 0, so we
recoded these values to improve the figures and keep them compact and readable.

```{r}
#| label: fig-log-bands
#| fig-cap: Distribution of log(Total Bands [k/ul]) for All Visits.
#| fig-width: 6
#| fig-height: 4
#| message: false

p1 <- Final_Visits %>% 
  ggplot(., aes(x = log_Total_Bands)) + 
  geom_histogram(aes(y=after_stat(count)), colour="black", fill="grey70") +
  theme_bw() +
  scale_x_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
                     labels = c("-Inf", as.character(-3:2)),
                     oob = function(x, limits) x) +
  scale_y_continuous(limits=c(-0.1, 500)) +
  xlab("log(Total Bands [k/ul])") +
  ylab("No. of Patients") +
  ggtitle("Histogram")

p2 <- Final_Visits %>% 
  ggplot(., aes(x = log_Total_Bands, y = "")) +
  geom_jitter(alpha = .3, height = .25, width = .05) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(xlim = c(-5, 2)) +
  scale_y_discrete( ) +
  scale_x_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
                     labels = c("-Inf", as.character(-3:2))) +
  xlab("log(Total Bands [k/ul])")  + 
  ylab(" ")  +
  ggtitle("Boxplot")

p1 + p2 + 
  plot_layout(nrow = 2, heights = c(3,1)) + 
  plot_annotation(tag_levels = "A")
```

Meanwhile @fig-log-bands-gt0 also shows log-transformed values of total bands,
but it only uses data from visits with total bands \> 0 to get a better look at
the shape of the distribution.

```{r}
#| label: fig-log-bands-gt0
#| fig-cap: Distribution of log(Total Bands [k/ul]) for Visits with Total Bands > 0.
#| fig-width: 6
#| fig-height: 4
#| message: false

p1 <- Final_Visits %>% 
  filter(Total_Bands > 0) %>% 
  ggplot(., aes(x = log_Total_Bands)) + 
  geom_histogram(aes(y=after_stat(count)), colour="black", fill="grey70") +
  theme_bw() +
  scale_x_continuous(limits = c(-3, 2), breaks = c(-3:2), 
                     oob = function(x, limits) x) +
  scale_y_continuous(limits=c(-0.1, 40)) +
  xlab("log(Total Bands [k/ul])") +
  ylab("No. of Patients") +
  ggtitle("Histogram")

p2 <- Final_Visits %>% 
  filter(Total_Bands > 0) %>% 
  ggplot(., aes(x = log_Total_Bands, y = "")) +
  geom_jitter(alpha = .3, height = .25, width = .05) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 2)) +
  scale_y_discrete( ) +
  scale_x_continuous(limits = c(-3, 2), breaks = c(-3:2)) +
  xlab("log(Total Bands [k/ul])")  + 
  ylab(" ")  +
  ggtitle("Boxplot")

p1 + p2 + 
  plot_layout(nrow = 2, heights = c(3,1)) + 
  plot_annotation(tag_levels = "A")
```

### Total Bands by Drug

@fig-bands-drug shows the boxplot of **Total Bands (k/ul)**, categorized by drugs.

```{r}
#| label: fig-bands-drug
#| fig-cap: Boxplot of Total Bands [k/ul] by Drug for All Visits
#| fig-width: 4
#| fig-height: 2

Final_Visits %>% 
  ggplot(., aes(x = Name, y = Total_Bands)) +
  geom_jitter(alpha = .3, width = .25, height = 0) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 7)) +
  scale_y_continuous(limits = c(0, 7), breaks = c(0:7)) +
  xlab("Drug")  +
  ylab("Total Bands (k/ul)")
```

@tbl-visit-describe-cont-by-Drug gives descriptive statistics for continuous 
biomarkers, broken down by drug name.

```{r}
#| label: tbl-visit-describe-cont-by-Drug
#| tbl-cap: !expr paste0("Descriptive Statistics for Continuous Biomarker ", 
#|                       "Variables by Drug (N = ", nrow(Final_Visits), 
#|                       " Visits)")
#| message: false

VNames <- c("Total_WBC", "Total_Neutrophil", "Total_Bands", "BN_Ratio")
VNamesUnits <- c("Total_WBC (k/ul)", 
                 "Total_Neutrophil (k/ul)", 
                 "Total_Bands (k/ul)", 
                 "BN_Ratio")

CNames <- c("Drug", "Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", 
            "Q75", "Max")

Final_Visits %>% 
  select(Name, Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio) %>% 
  pivot_longer(cols = c(Total_WBC, Total_Neutrophil, Total_Bands, BN_Ratio), 
               names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Name, Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Variable = factor(Variable, levels = VNames, labels = VNamesUnits)) %>% 
  kable(., format = "html", digits = 1, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  collapse_rows(., columns = 1, valign = "top", row_group_label_position = "stack") 
```

@fig-log-total-bands-drug shows the boxplot of **log(Total bands) (k/ul)**, 
categorized by drugs, where -Inf means total bands = 0. 

```{r}
#| label: fig-log-total-bands-drug
#| fig-cap: Boxplots of log(Total Bands [k/ul]) by drug for (A) all visits and 
#|          (B) visits with Total Bands > 0 k/ul.
#| fig-width: 4
#| fig-height: 5

p1 <- Final_Visits %>% 
  ggplot(., aes(x = Name, y = log_Total_Bands)) +
  geom_jitter(alpha = .3, width = .25, height = 0) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(ylim = c(-5, 2)) +
  scale_y_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
                     labels = c("-Inf", as.character(-3:2))) +
  xlab("Drug")  +
  ylab("log(Total Bands [k/ul])")
  

p2 <- Final_Visits %>% 
  filter(Total_Bands > 0) %>% 
  ggplot(., aes(x = Name, y = log_Total_Bands)) +
  geom_jitter(alpha = .3, width = .25, height = 0) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  # coord_cartesian(ylim = c(-5, 2)) +
  # scale_y_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
  #                    labels = c("-Inf", as.character(-3:2))) +
  coord_cartesian(ylim = c(-3, 2)) +
  scale_y_continuous(limits = c(-3, 2), breaks = c(-3:2)) +
  xlab("Drug")  +
  ylab("log(Total Bands [k/ul])")

p1 + p2 +
 plot_layout(nrow = 2) + 
 plot_annotation(tag_levels = "A")
```

@tbl-drug-Band-Category and @fig-drug-Band-Category show the number and percent 
of CBCs by band neutrophil category for each drug used. 

```{r}
#| label: tbl-drug-Band-Category
#| tbl-cap: Band Category Frequency by Prior Drug Administered

Drug1_N <- Final_Visits %>% filter(Drug == 1) %>% nrow()
Drug2_N <- Final_Visits %>% filter(Drug == 2) %>% nrow()
Drug3_N <- Final_Visits %>% filter(Drug == 3) %>% nrow()

Drug_Band_Category <- Final_Visits %>% 
  xtabs(~Band_Category + Drug, data = .) %>% 
  addmargins(margin = 1) %>% 
  as.data.frame() %>% 
  rename(N = Freq) %>% 
  mutate(Name = factor(Drug, labels = DrugNames), 
         Percent = case_when(Drug == 1 ~ 100*N/Drug1_N,
                             Drug == 2 ~ 100*N/Drug2_N,
                             Drug == 3 ~ 100*N/Drug3_N)) %>% 
  relocate(Drug, Name, Band_Category) 

Drug_Band_Category %>% 
  kable(digits = 1, 
        col.names = c("Drug", "Name", "Band Category", "N", "Percent")) %>% 
  kable_styling() 
```

```{r}
#| label: fig-drug-Band-Category
#| fig-cap: Band Category Distribution by Prior Drug Administered
#| fig-width: 7
#| fig-height: 4
#| message: false

Drug_Band_Category %>% 
  filter(Band_Category != "Sum") %>% 
ggplot(., aes(x = Band_Category, y = Percent, fill = Band_Category)) + 
  geom_bar(stat = "identity", col = "black") +
  scale_fill_brewer(name = "Band Neutrophil Category", palette = "Greys") +
  xlab("Band Neutrophil Category") +
  scale_x_discrete(labels=c("Absent (0.0 k/ul)" = "Absent", 
                            "Present (<= 0.1 k/ul)" = "Present",
                            "Bandemia (> 0.1 k/ul)" = "Bandemia")) +
  ylab("Percent of CBCs") +
  scale_y_continuous(limits=c(0, 100)) +
  facet_wrap(~ Name) + 
  theme_bw()+
  theme(legend.position = "bottom")
```

### Total Bands by Visit Number
@tbl-bands-visit provides the mean and standard deviation of 
**Total Bands** for each visit. From this data, we can observe that the 
mean value of Total bands tends to decrease over time.

```{r}
#| label: tbl-bands-visit
#| tbl-cap: Summary of Total Bands (k/ul) By VIsit Number

Total_Bands_counts <- Final_Visits %>%
  group_by(Visit_Number) %>%
   summarise(N = sum(!is.na(Total_Bands)),
    Mean = round(mean(Total_Bands, na.rm = TRUE),4),
    SD= round(sd(Total_Bands, na.rm = TRUE),4),
    Min = min(Total_Bands, na.rm = TRUE),
    Q25 = quantile(Total_Bands, probs = .25, na.rm = TRUE),
    Q50 = quantile(Total_Bands, probs = .50, na.rm = TRUE),
    Q75 = quantile(Total_Bands, probs = .75, na.rm = TRUE),
    Max= max(Total_Bands, na.rm = TRUE))

Total_Bands_counts %>%
  kable(digits = 1) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) 
```

@fig-bands-visit shows the boxplot of total bands (k/ul) for each visit, 
@fig-log-bands-visit shows the corresponding boxplots for log-transformed 
values both for all visits and the subset of visits with total bands > 0. In the 
log scale, -Inf corresponds to total bands = 0 k/ul.

```{r}
#| label: fig-bands-visit
#| fig-cap: Boxplot of Total Bands (k/ul) by Visit Number
#| fig-width: 6
#| fig-height: 2.5

Final_Visits %>% 
  ggplot(., aes(x = factor(Visit_Number), y = Total_Bands, 
                group = Visit_Number)) +
  geom_jitter(alpha = .3, width = .25, height = .05) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 7)) +
  scale_y_continuous(limits = c(-.1, 7), breaks = c(0:7)) +
  xlab("Visit Number") +
  ylab("Total Bands (k/ul)")
```

```{r}
#| label: fig-log-bands-visit
#| fig-cap: Boxplot of Total Bands (k/ul) by Visit Number for (A) All Visits and 
#|          (B) Visits with Total Bands > 0 k/ul. 
#| fig-width: 6
#| fig-height: 5

p1 <- Final_Visits %>% 
  ggplot(., aes(x = factor(Visit_Number), y = log_Total_Bands, 
                group = Visit_Number)) +
  geom_jitter(alpha = .3, width = .25, height = .05) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  coord_cartesian(ylim = c(-5, 2)) +
  scale_y_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
                     labels = c("-Inf", as.character(-3:2))) +
  xlab("Visit Number") +
  ylab("log(Total Bands [k/ul])")

p2 <- Final_Visits %>% 
  filter(Total_Bands > 0) %>% 
  ggplot(., aes(x = factor(Visit_Number), y = log_Total_Bands, 
                group = Visit_Number)) +
  geom_jitter(alpha = .3, width = .25, height = .05) +
  geom_boxplot(outliers = FALSE, alpha = .5) +
  theme_bw() +
  # coord_cartesian(ylim = c(-5, 2)) +
  # scale_y_continuous(limits = c(-5.1, 2), breaks = c(-5, -3:2), 
  #                    labels = c("-Inf", as.character(-3:2))) +
  coord_cartesian(ylim = c(-3, 2)) +
  scale_y_continuous(limits = c(-3, 2), breaks = c(-3:2)) +
  xlab("Visit Number") +
  ylab("log(Total Bands [k/ul])")

p1 + p2 +
 plot_layout(nrow = 2) + 
 plot_annotation(tag_levels = "A")
```

::: {.callout-note}
@fig-log-bands-visit (or either of its panels alone) may become a figure 
in the manuscript. We used a variation of panel A as Figure 3 in the prior 
manuscript, which we are now revising for submission to another journal. 
:::

@tbl-bands-eq0-visit shows the number and proportion of instances where total
bands equal 0 for each drug used.

```{r}
#| label: tbl-bands-eq0-visit
#| tbl-cap: Number and Proportion of Visits With Total Bands = 0 k/ul by Visit.

Final_Visits %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "N", "Proportion")) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "Total Bands = 0" = 2))
```

#### Drug 1

@tbl-total_bands_by_visit_drug1 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 1 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug1
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 1 at each visit time.

Final_Visits %>% 
  filter(Drug==1) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_total_bands is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_total_bands
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 1"
boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug1_log shows the boxplot of 
log(total bands x100/ul) for individuals using Drug 1 at each visit time, where 
-Inf means total bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1."

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug1_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 1 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log_part
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1 (only include those with total bands > 0)."

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug1_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug1_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 1.

Final_Visits %>% 
  filter(Drug == 1) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "N", "Proportion")) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "Total Bands = 0" = 2))
```

```{r}
#| label: drug

drug1 <- Final_Visits %>% filter(Drug==1)
drug2 <- Final_Visits %>% filter(Drug==2)
drug3 <- Final_Visits %>% filter(Drug==3)
```

@fig-visit_times_of_drug_1 shows the times of using drug 1 at each visit number.

```{r}
#| label: fig-visit_times_of_drug_1
#| fig-cap: Times of using drug 1 at each visit number

drug_counts <- drug1 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 2  

@tbl-total_bands_by_visit_drug2 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 2 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug2
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 2 at each visit time.

Final_Visits %>% 
  filter(Drug==2) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_drug_2 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_2
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 2"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug2_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug2_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log_part
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2 (only include those with total bands \> 0).

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug2_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug2_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 2.

Final_Visits %>% 
  filter(Drug == 2) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "N", "Proportion")) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "Total Bands = 0" = 2))
```

@fig-visit_times_of_drug_2 shows the times of using drug 2 at each visit number

```{r}
#| label: fig-visit_times_of_drug_2
#| fig-cap: Times of using drug 2 at each visit number

drug_counts <- drug2 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 3


@tbl-total_bands_by_visit_drug3 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 3 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug3
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 3 at each visit time.

Final_Visits %>% 
  filter(Drug==3) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

@fig-boxplot_drug_3 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_3
#| #| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 3"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==3,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug3_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 3.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==3,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug3_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log_part
#| fig-cap: Boxplot of log(total bands x100/ul) for each visit time for those using drug 3.
boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled>0]~Visit_Number[Total_Bands_Rescaled>0],data=Final_Visits[Final_Visits$Drug==3,],xlab="Visit number",ylab="log(Total bands) (x100/ul)")
```

@tbl-total_bands_by_visit_drug3_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug3_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 3.

Final_Visits %>% 
  filter(Drug == 3) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "N", "Proportion")) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "Total Bands = 0" = 2))
```

@fig-visit_times_of_drug_3 shows the times of using drug 3 at each visit number

```{r}
#| label: fig-visit_times_of_drug_3
#| fig-cap: Times of using drug 3 at each visit number

drug_counts <- drug3 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

## Total White Blood Cells (WBC) 
@tbl-WBC-visit provides the mean and standard deviation of total WBC for each 
visit number. 

```{r}
#| label: tbl-WBC-visit
#| tbl-cap: Summary of total white blood cells (x1000/uL) for each visit time.
Final_Visits %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_WBC, na.rm = TRUE),4),
            sd= round(sd(Total_WBC, na.rm = TRUE),4),
            min = min(Total_WBC, na.rm = TRUE),
            max= max(Total_WBC, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```


## Neutrophils (neuts)
@tbl-total_neutrophil_by_visit provides the mean and standard deviation of 
**Total neutrophil** for each visit time. 

```{r}
#| label: tbl-total_neutrophil_by_visit
#| tbl-cap: Summary of total neutrophil for each visit time.

Final_Visits %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Neutrophil, na.rm = TRUE),4),
            sd= round(sd(Total_Neutrophil, na.rm = TRUE),4),
            min = min(Total_Neutrophil, na.rm = TRUE),
            max= max(Total_Neutrophil, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```

## Bandemia Grade  

```{r}
#| label: tbl-Bandemia-Grade
#| tbl-cap: Frequency Distribution of Bandemia Grade
#| warning: false

Final_Visits %>% 
  mutate(Bandemia_Grade = factor(Bandemia_Grade, levels = 0:5)) %>% 
  group_by(Bandemia_Grade, .drop = FALSE) %>% 
  summarize(Min = min(Total_Bands),
            Max = max(Total_Bands),
            N = n(),
            Proportion = N/nrow(Final_Visits)) %>% 
  ungroup() %>% 
  mutate(MSUVDL_Range = fct_recode(Bandemia_Grade, "0.00-0.10" = "0", 
                                   "0.11-0.29" = "1", "0.30-0.59" = "2", 
                                   "0.60-0.99" = "3", ">= 1.00" = "4", 
                                   "death" = "5" )) %>% 
  select(Bandemia_Grade, MSUVDL_Range, Min, Max, N, Proportion) %>% 
  kable(digits = 3, col.names = c("Bandemia Grade", 
                                  "MSUVDL Reference Range (k/ul)", "Min", "Max",
                                  "N", "Proportion")) %>%
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "Observed Total_Bands (k/ul)" = 2, " " = 2))
```

## Drug Sequences
@tbl-visit-patterns shows the frequency of occurrence for each unique sequence 
of drug use observed. Each row represents a group of dogs who got the same
sequence of drugs across the 10 visits of interest. Ideally, the most common
sequence will match the intended CHOP protocol sequence for using the drugs. The
table is sorted in descending order by sequence frequency (`N_Dogs`) then by
`N_Visits`, which is the number of visits in the pattern that were actually
observed. Columns `V1` to `V10` show the drug number used at a specific visit
number in the given sequence. Because only data about drugs 1-3 were included,
it is possible that blanks in the middle of a sequence were visits where some
other drug was administered.

```{r}
#| label: tbl-visit-patterns
#| tbl-cap: !expr paste0("Frequency Distribution of Drug Sequence ", 
#|                       "(N = ", length(unique(Final_Visits$Patient)), 
#|                       " Patients)")
#| warning: false

options(knitr.kable.NA = '-')

Final_Visits %>% 
  select(Patient, Visit_Number, Drug) %>% 
  rename(VN = Visit_Number) %>% 
  pivot_wider(id_cols = "Patient", names_from = "VN", names_prefix = "V", 
              values_from = "Drug") %>% 
  group_by(V1, V2, V3, V4, V5, V6, V7, V8, V9, V10) %>% 
  summarize(N_Dogs = n()) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(N_Visits = sum(!is.na(c_across(V1:V10)))) %>% 
  ungroup() %>% 
  arrange(desc(N_Dogs), desc(N_Visits)) %>% 
  rowid_to_column(var = "SequenceID") %>% 
  relocate(SequenceID, N_Dogs, N_Visits) %>% 
  kable() %>%
  kable_styling() %>% 
  add_header_above(header = c(" " = 3, "Drug Sequence (Drug Used by Visit Number)" = 10))
```

@tbl-visit-patterns-subset shows the subset of drug sequences that deviate from 
the CHOP protocol by using cyclosphosphamide at visit 1. 

```{r}
#| label: tbl-visit-patterns-subset
#| tbl-cap: !expr paste0("Subset of Drug Sequences That Used Cyclosphosphamide ", 
#|                       "at Visit 1 ",
#|                       "(N = ", length(unique(Final_Visits$Patient)), 
#|                       " Patients)")
#| warning: false

options(knitr.kable.NA = '-')

Final_Visits %>% 
  select(Patient, Visit_Number, Drug) %>% 
  rename(VN = Visit_Number) %>% 
  pivot_wider(id_cols = "Patient", names_from = "VN", names_prefix = "V", 
              values_from = "Drug") %>% 
  group_by(V1, V2, V3, V4, V5, V6, V7, V8, V9, V10) %>% 
  summarize(N_Dogs = n()) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(N_Visits = sum(!is.na(c_across(V1:V10)))) %>% 
  ungroup() %>% 
  arrange(desc(N_Dogs), desc(N_Visits)) %>% 
  rowid_to_column(var = "SequenceID") %>% 
  relocate(SequenceID, N_Dogs, N_Visits) %>% 
  filter(V1 == 2) %>% 
  kable() %>%
  kable_styling() %>% 
  add_header_above(header = c(" " = 3, "Drug Sequence (Drug Used by Visit Number)" = 10))
```


# References
::: {#refs}
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
