---
title: Descriptive Analysis of CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceFile: "Descriptive_Analysis.qmd"
  LogFile: "Descriptive_Analysis.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
---

\FloatBarrier

# Purpose

This file provides some descriptive analyses of data provided by Alison Masyr 
and Caitlin Eliason that contains observations of dogs receiving CHOP 
chemotherapy at MSU Veterinary Medicine Clinic (VMC) between 2016 and 2022. 

# Setup

This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Results.

## Define Global Options

Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages

Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(here)       # for here(), i_am()
library(dplyr)      # for filter(), group_by(), mutate(), rename(), summarise()
library(tidyr)      # for pivot_longer()
library(ggplot2)    # for ggplot()
library(knitr)      # for kable()
library(kableExtra) # for kable_styling 
library(gtsummary)  # for 
library(readxl)     # for read_xlsx()
library(quarto)     # for quarto_version()
library(CHOPStudy)  # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Descriptive_Analysis.qmd")
```

## Load Data {#sec-Load-Data}

::: {.callout-caution}
Render the `scripts/Import_Data.qmd` file to create the data file loaded below. 
:::

```{r}
#| label: load-data

load(file=here("data/Imported_CHOP_Data.RData"))
```

# Results {#sec-Results}
@tbl-patient-describe shows patient-level descriptive statistics for weight 
surface area, and number of visits, while @tbl-visit-describe-cont and 
@tbl-visit-describe-cat respectively show visit-level descriptive statistics 
for continuous and catgegorical biomarker variables. 

```{r}
#| label: tbl-patient-describe
#| tbl-cap: !expr paste0("Patient-Level Descriptive Statistics by Variable ", 
#|                       "(N = ", nrow(Final_Patients), " Patients)")
#| message: false
#| results: asis

VNames <- c("Weight_kg", "Surface_Area_m2", "N_Visits")

CNames <- c("Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", "Q75", "Max")

Final_Patients %>% 
  select(Weight_kg, Surface_Area_m2, N_Visits) %>% 
  pivot_longer(cols = c(Weight_kg, Surface_Area_m2), names_to = "Variable", 
               values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  kable(., format = "html", digits = 2, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  cat()
```


```{r}
#| label: tbl-visit-describe-cont
#| tbl-cap: !expr paste0("Visit-Level Descriptive Statistics for Continuous ", 
#|                       "Biomarker Variables (N = ", nrow(Final_Visits), " Visits)")
#| message: false
#| results: asis

VNames <- c("Total_WBC", "Total_Neutrophil", "Total_Bands", 
            "Total_Bands_Rescaled", "BN_Ratio", "log_Total_Bands",
            "log_Total_Bands_Rescaled")

CNames <- c("Variable", "N", "Mean", "SD", "Min", "Q25", "Q50", "Q75", "Max")

Final_Visits %>% 
  select(Total_WBC, Total_Neutrophil, Total_Bands, Total_Bands_Rescaled, 
  BN_Ratio, log_Total_Bands, log_Total_Bands_Rescaled) %>% 
  pivot_longer(cols = c(Total_WBC, Total_Neutrophil, Total_Bands, 
                        Total_Bands_Rescaled, BN_Ratio, log_Total_Bands,
                        log_Total_Bands_Rescaled), 
               names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = factor(Variable, levels = VNames)) %>% 
  # Aggregate to one row per variable.
  group_by(Variable) %>% 
  summarise(N = sum(!is.na(Value)), 
            Mean = mean(Value, na.rm = TRUE),
            SD = sd(Value, na.rm = TRUE),
            Min = min(Value, na.rm = TRUE),
            Q25 = quantile(Value, probs = .25, na.rm = TRUE),
            Q50 = quantile(Value, probs = .50, na.rm = TRUE),
            Q75 = quantile(Value, probs = .75, na.rm = TRUE),
            Max = max(Value, na.rm = TRUE)) %>% 
  kable(., format = "html", digits = 1, 
        col.names = CNames, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T) %>% 
  cat()
```

## Total bands

### Summary of total bands

@fig-total-bands shows the histogram (left) and the boxplot (right) 
of **Total bands** in the original units (x10.e3/uL=1000/uL), while 
@fig-total-bands-rescaled shows the same data after rescaling to 
units of (x100/ul) to obtain integer values usable in count models. This makes 
it more convenient for us when fitting models. Both these figures show that 
there are lots of zero values, which makes it hard to see the shape of the 
distribution among the non-zero values. 

```{r}
#| label: fig-total-bands
#| fig-cap: Distribution of Total Bands (x1000/ul)
#| fig-width: 6
#| fig-height: 4
#| message: false

ggplot(Final_Visits, aes(x = Total_Bands)) + 
  geom_histogram(aes(y=..count..), colour="black", fill="grey70") +
  scale_x_continuous(limits = c(-0.2,7), n.breaks = 8, oob = function(x, limits) x) +
  scale_y_continuous(limits=c(-0.2, 500)) +
  xlab("Total Bands (x1000/ul)") +
  ylab("No. of Patients")
```

```{r}
#| label: fig-total-bands-rescaled
#| fig-cap: Distribution of Rescaled Total Bands (x100/ul)
#| fig-width: 6
#| fig-height: 4
#| message: false

ggplot(Final_Visits, aes(x = Total_Bands_Rescaled)) + 
  geom_histogram(aes(y=..count..), colour="black", fill="grey70") +
  scale_x_continuous(limits = c(-2,70), n.breaks = 8, oob = function(x, limits) x) +
  scale_y_continuous(limits=c(-0.2, 500)) +
  xlab("Rescaled Total Bands (x100/ul)") +
  ylab("No. of Patients")
```


The histogram (left) and boxplot (right) of **log(Total bands [x100/ul])** and 
**log(Total bands [x1000/ul])** are shown in 
@fig-chop1_phase1_total_bands_rescaled_log and 
@fig-chop1_phase1_total_bands_log, respectively. The difference between these 
two graphs are in the units of measurement prior to the log transformation. In 
both graphs, the log equals -Infinity when total bands are 0, so we recoded 
these values to improve the figures and keep them compact and readable. And we 
also present figures in 
@fig-chop1_phase1_total_bands_rescaled_log_part and 
@fig-chop1_phase1_total_bands_log_part that show only the data where total bands 
\> 0 to better look at the shape of the distribution.

```{r}
#| label: fig-chop1_phase1_total_bands_rescaled_log
#| fig-cap: Histogram (left) and boxplot (right) of the log(total bands x100/ul)

par(mfrow=c(1,2))
hist(Final_Visits$log_Total_Bands_Rescaled,
     xlab="log(Total bands [x100/ul])",
     ylab = "Number of patients",
     main=NULL,
     breaks = 7, xlim = c(-3, 5), xaxt = "n", ylim = c(0, 500))
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(1, at = axis_positions, labels = axis_labels)

boxplot(Final_Visits$log_Total_Bands_Rescaled,
       ylab="log(Total bands [x100/ul])",
       ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)

par(mfrow=c(1,1))
```

```{r}
#| label: fig-chop1_phase1_total_bands_log
#| fig-cap: Histogram (left) and boxplot (right) of the log(total bands x1000/ul)

par(mfrow=c(1,2))
hist(Final_Visits$log_Total_Bands, 
     xlab="log(Total bands [x1000/ul])", 
     ylab = "Number of patients",
     main=NULL, 
     breaks = 6, xlim = c(-5, 2), xaxt = "n", ylim = c(0, 500))
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(1, at = axis_positions, labels = axis_labels)

boxplot(Final_Visits$log_Total_Bands,
        ylab="log(Total bands [x1000/ul])",
        ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)

par(mfrow=c(1,1))
```

::: {.callout-note}
@fig-chop1_phase1_total_bands_log above becomes Figure 1 in the manuscript. 
:::

@tbl-Numbers_of_zero_and_proportion shows the frequency distribution for Total 
Bands, which highlights the large proportion of zero values. The table also 
shows the rescaled Total Bands values used in count models, and log transformed
values for both original and recscaled Total Bands values. 

```{r}
#| label: tbl-Numbers_of_zero_and_proportion
#| tbl-cap: "Frequency Distribution of Total Bands"

Final_Visits %>% 
  count(Total_Bands) %>% 
  mutate(Total_Bands_Rescaled = Total_Bands*10,
         Log_Total_Bands = log(Total_Bands), 
         Log_Total_Bands_Rescaled = log(Total_Bands_Rescaled),
         Proportion = n/nrow(Final_Visits)) %>% 
  select(Total_Bands, Total_Bands_Rescaled, Log_Total_Bands, 
         Log_Total_Bands_Rescaled, n, Proportion) %>% 
  kable(digits = 3, col.names = c("Total bands (x1000/ul)", 
                                  "Total bands (x100/ul)", 
                                  "log(Total bands [x1000/ul])", 
                                  "log(Total bands [x100/ul])", 
                                  "n", 
                                  "Proportion")) %>%
  kable_styling() %>% 
  add_header_above(header = c("Raw Values", "Transformations of Total Bands" = 3, 
                              "Frequency" = 2))
```

```{r}
#| label: fig-chop1_phase1_total_bands_rescaled_log_part
#| fig-cap: "Histogram (left) and boxplot (right) of the log(total bands [x100/ul]) with only total bands > 0"
par(mfrow=c(1,2))
hist(Final_Visits$log_Total_Bands_Rescaled[Final_Visits$log_Total_Bands_Rescaled>-3],
     xlab="log(Total bands [x100/ul])", xlim = c(0, 5),
     ylab = "Number of patients",
     main=NULL)
boxplot(Final_Visits$log_Total_Bands_Rescaled[Final_Visits$log_Total_Bands_Rescaled>-3],
        ylab="log(Total bands [x100/ul])", ylim = c(0,5))
par(mfrow=c(1,1))
```

```{r}
#| label: fig-chop1_phase1_total_bands_log_part
#| fig-cap: "Histogram (left) and boxplot (right) of the log(total bands [x1000/ul]) with only total bands > 0"
par(mfrow=c(1,2))
hist(Final_Visits$log_Total_Bands[Final_Visits$log_Total_Bands>-5],
     xlab="log(Total bands [x1000/ul])", xlim = c(-3, 2),
     ylab = "Number of patients",
     main=NULL)
boxplot(Final_Visits$log_Total_Bands[Final_Visits$log_Total_Bands>-3],
        ylab="log(Total bands [x1000/ul])", ylim = c(-3,2))
par(mfrow=c(1,1))
```

### Summary of each drug

@tbl-chop1_drug shows the number of times each drug is used. We can see that
Drug 1 is used far more than other drugs.

```{r}
#| label: tbl-chop1_drug
#| tbl-cap: Summary the number of times each drug is used

Final_Visits %>% 
  mutate(Drug = factor(Drug, levels = 1:3, labels = paste("Drug", 1:3))) %>% 
  count(Drug) %>% 
  mutate(Proportion = n/nrow(Final_Visits)) %>% 
  kbl(digits = 3) %>%
  kable_styling()
```

@fig-total_bands_rescaled_by_drugs shows the boxplot of 
**Total bands (x100/ul)**, categorized by drugs, while @fig-total_bands_by_drugs 
shows the boxplot of **Total bands (x1000/ul)**, categorized by drugs.

```{r}
#| label: fig-total_bands_rescaled_by_drugs
#| fig-cap: Boxplot of the total bands (x100/ul), categorized by drugs.

boxplot(Final_Visits$Total_Bands_Rescaled ~ Final_Visits$Drug,
        ylab="Total bands (x100/ul)",
        xlab = "Drug")
```

```{r}
#| label: fig-total_bands_by_drugs
#| fig-cap: Boxplot of the total bands (x1000/ul), categorized by drugs.

boxplot(Final_Visits$Total_Bands ~ Final_Visits$Drug,
        ylab="Total bands (x1000/ul)",
        xlab = "Drug")
```

@tbl-total_bands_by_drugs gives the mean and median value of the **Total bands**, 
categorized by drugs.

```{r}
#| label: tbl-total_bands_by_drugs
#| tbl-cap: Summary of total bands (x100/ul), categorized by drugs.

A=matrix(0,ncol=3,nrow=6)
for(i in 1:3){
  A[,i]=summary(Final_Visits$Total_Bands_Rescaled[Final_Visits$Drug==i])
}
colnames(A)=paste0("Drug ",1:3)
rownames(A)=c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max")
round(A,4)%>%   kbl() %>%   kable_styling()
```

@fig-total_bands_rescaled_by_drugs_log shows the boxplot of 
**log(Total bands) (x100/ul)**, categorized by drugs, where -Inf means total 
bands=0. Meanwhile, @fig-total_bands_by_drugs_log shows the boxplot of 
**log(Total bands) (x1000/ul)**, categorized by drugs, where -Inf means total 
bands=0. 

```{r}
#| label: fig-total_bands_rescaled_by_drugs_log
#| fig-cap: Boxplot of the log(total bands)(x100/ul), categorized by drugs.

boxplot(Final_Visits$log_Total_Bands_Rescaled ~ Final_Visits$Drug,
        xlab = "Drug",
        ylab="log(Total bands) (x100/ul)", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))

axis(2, at = axis_positions, labels = axis_labels)
```

```{r}
#| label: fig-total_bands_by_drugs_log
#| fig-cap: Boxplot of the log(total bands [x1000/ul]), categorized by drugs.

boxplot(Final_Visits$log_Total_Bands ~ Final_Visits$Drug,
        xlab = "Drug",
        ylab="log(Total bands [x1000/ul])", ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

::: {.callout-note}
@fig-total_bands_by_drugs_log above becomes Figure 2 in the manuscript. 
:::

@fig-total_bands_by_drugs_log_part shows the boxplot of 
**log(Total bands) (x100/ul)** with only Total bands \> 0, categorized by drugs.

```{r}
#| label: fig-total_bands_by_drugs_log_part
#| fig-cap: Boxplot of the log(total bands [x100/ul]), categorized by drugs.

boxplot(Final_Visits$log_Total_Bands_Rescaled[Final_Visits$log_Total_Bands_Rescaled>-3] ~ Final_Visits$Drug[Final_Visits$log_Total_Bands_Rescaled>-1],
        ylab="log(Total bands [x100/ul])",
        xlab = "Drug")

```

@tbl-total_bands_by_drugs_zero shows the number and proportion of instances 
where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_drugs_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each drug used.
#| 
Final_Visits %>% 
  mutate(Drug = factor(Drug, levels = 1:3, labels = paste("Drug", 1:3)),
         Zero_Bands = Total_Bands == 0) %>% 
  group_by(Drug) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Drug", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

### Summary of total bands by visit time

#### Total

@tbl-total_bands_by_visit provides the mean and standard deviation of 
**Total bands** for each visit time. From this data, we can observe that the 
mean value of Total bands tends to decrease over time.

```{r}
#| label: tbl-total_bands_by_visit
#| tbl-cap: Summary of total bands (x100/ul) for each visit time.
Total_Bands_Rescaled_counts <- Final_Visits %>%
  group_by(Visit_Number) %>%
   summarise(
    mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
    sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
    min = min(Total_Bands_Rescaled, na.rm = TRUE),
    max= max(Total_Bands_Rescaled, na.rm = TRUE)
  )
Total_Bands_Rescaled_counts%>%kable()%>%kable_styling() 
```

@fig-total_bands_by_visit shows the boxplot of total bands (x100/ul) for each 
visit time.

```{r}
#| label: fig-total_bands_by_visit
#| fig-cap: Boxplot of total bands (x100/ul) for each visit time.
boxplot(Total_Bands_Rescaled~Visit_Number,data=Final_Visits,xlab="Visit number",ylab="Total bands (x100/ul)")
```

@fig-total_bands_rescaled_by_visit_log shows the boxplot of 
**log(total bands x100/ul)** for each visit time, where -Inf means total bands = 
0. @fig-total_bands_by_visit_log shows the boxplot of 
**log(total bands x1000/ul)** for each visit time, where -Inf means total bands = 
0.

```{r}
#| label: fig-total_bands_rescaled_by_visit_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time.
boxplot(log_Total_Bands_Rescaled ~ Visit_Number, data=Final_Visits,
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])",ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

```{r}
#| label: fig-total_bands_by_visit_log
#| fig-cap: Boxplot of log(total bands [x1000/ul]) for each visit time.
boxplot(log_Total_Bands ~ Visit_Number, data=Final_Visits,
        xlab="Visit number",
        ylab="log(Total bands [x1000/ul])",ylim = c(-5, 2), yaxt = "n")
axis_positions <- c(-5, seq(-3, 2, length.out = 6))
axis_labels <- c("-Inf", formatC(seq(-3, 2, length.out = 6), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

::: {.callout-note}
@fig-total_bands_by_visit_log above becomes Figure 3a in the manuscript. 
:::


@fig-total_bands_by_visit_log_part shows the boxplot of log(total bands x100/ul)
for each visit time and only include those with total bands \> 0.

```{r}
#| label: fig-total_bands_by_visit_log_part
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time (only include those with total bands \> 0).
boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled > 0],
        data=Final_Visits,
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_zero shows the number and proportion of instances 
where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time.

Final_Visits %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

#### Drug 1

@tbl-total_bands_by_visit_drug1 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 1 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug1
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 1 at each visit time.

Final_Visits %>% 
  filter(Drug==1) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_total_bands is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_total_bands
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 1"
boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug1_log shows the boxplot of 
log(total bands x100/ul) for individuals using Drug 1 at each visit time, where 
-Inf means total bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1."

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug1_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 1 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug1_log_part
#| fig-cap: "Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 1 (only include those with total bands > 0)."

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=Final_Visits[Final_Visits$Drug==1,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug1_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug1_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 1.

Final_Visits %>% 
  filter(Drug == 1) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

```{r}
#| label: drug
drug1 <- Final_Visits %>% filter(Drug==1)
drug2 <- Final_Visits %>% filter(Drug==2)
drug3 <- Final_Visits %>% filter(Drug==3)
```

@fig-visit_times_of_drug_1 shows the times of using drug 1 at each visit number.

```{r}
#| label: fig-visit_times_of_drug_1
#| fig-cap: Times of using drug 1 at each visit number

drug_counts <- drug1 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 2  

@tbl-total_bands_by_visit_drug2 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 2 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug2
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 2 at each visit time.

Final_Visits %>% 
  filter(Drug==2) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```


@fig-boxplot_drug_2 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_2
#| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 2"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug2_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug2_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 2 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug2_log_part
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 2 (only include those with total bands \> 0).

boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled > 0] ~ 
          Visit_Number[Total_Bands_Rescaled>0],
        data=Final_Visits[Final_Visits$Drug==2,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])")
```

@tbl-total_bands_by_visit_drug2_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug2_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 2.

Final_Visits %>% 
  filter(Drug == 2) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

@fig-visit_times_of_drug_2 shows the times of using drug 2 at each visit number

```{r}
#| label: fig-visit_times_of_drug_2
#| fig-cap: Times of using drug 2 at each visit number

drug_counts <- drug2 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

#### Drug 3


@tbl-total_bands_by_visit_drug3 provides the mean and standard deviation of 
**Total bands** for individuals using Drug 3 at each visit time.

```{r}
#| label: tbl-total_bands_by_visit_drug3
#| tbl-cap: Summary of total bands (x100/ul) for individuals using Drug 3 at each visit time.

Final_Visits %>% 
  filter(Drug==3) %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Bands_Rescaled, na.rm = TRUE),4),
            sd= round(sd(Total_Bands_Rescaled, na.rm = TRUE),4),
            min = min(Total_Bands_Rescaled, na.rm = TRUE),
            max= max(Total_Bands_Rescaled, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

@fig-boxplot_drug_3 is a boxplot that visualize the information.

```{r}
#| label: fig-boxplot_drug_3
#| #| fig-cap: "Boxplot of total bands (x100/ul) for each visit time for drug 3"

boxplot(Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==3,],
        xlab="Visit number",
        ylab="Total bands (x100/ul)")
```

@fig-total_bands_by_visit_drug3_log shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time, where -1 means total 
bands=0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log
#| fig-cap: Boxplot of log(total bands [x100/ul]) for each visit time for those using drug 3.

boxplot(log_Total_Bands_Rescaled~Visit_Number,
        data=Final_Visits[Final_Visits$Drug==3,],
        xlab="Visit number",
        ylab="log(Total bands [x100/ul])", ylim = c(-3, 5), yaxt = "n")
axis_positions <- c(-3, seq(-1, 5, length.out = 7))
axis_labels <- c("-Inf", formatC(seq(-1, 5, length.out = 7), 
                                 format = "f", digits = 0))
axis(2, at = axis_positions, labels = axis_labels)
```

@fig-total_bands_by_visit_drug3_log_part shows the boxplot of log(total bands 
x100/ul) for individuals using Drug 3 at each visit time and only include those 
with total bands\>0.

```{r}
#| label: fig-total_bands_by_visit_drug3_log_part
#| fig-cap: Boxplot of log(total bands x100/ul) for each visit time for those using drug 3.
boxplot(log_Total_Bands_Rescaled[Total_Bands_Rescaled>0]~Visit_Number[Total_Bands_Rescaled>0],data=Final_Visits[Final_Visits$Drug==3,],xlab="Visit number",ylab="log(Total bands) (x100/ul)")
```

@tbl-total_bands_by_visit_drug3_zero shows the number and proportion of 
instances where total bands equal 0 for each drug used.

```{r}
#| label: tbl-total_bands_by_visit_drug3_zero
#| tbl-cap: Number and proportion of instances where total bands equal 0 for each visit time for those using drug 3.

Final_Visits %>% 
  filter(Drug == 3) %>% 
  mutate(Zero_Bands = Total_Bands == 0) %>% 
  group_by(Visit_Number) %>% 
  summarize(n = n(),
            Zero_Bands_N = sum(Zero_Bands),
            Zero_Bands_Proportion = mean(Zero_Bands)) %>% 
  kable(digits = 3, 
        col.names = c("Visit Number", "Total N", "Total Bands = 0, N", 
                      "Total Bands = 0, Proportion")) %>% 
  kable_styling()
```

@fig-visit_times_of_drug_3 shows the times of using drug 3 at each visit number

```{r}
#| label: fig-visit_times_of_drug_3
#| fig-cap: Times of using drug 3 at each visit number

drug_counts <- drug3 %>%
  group_by(Visit_Number) %>%
  summarise(Count = n())
ggplot(drug_counts, aes(x = Visit_Number, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 2.5) +
  labs(x = "Visit number", y = "Times") +
  scale_x_continuous(breaks = c(1:16))+
  theme_minimal()
```

### Additional values needed  
#### Total white blood cells (WBC) 
@tbl-total_WBC_by_visit provides the mean and standard deviation of 
**Total WBC** for each visit time. From this data, we can observe that the mean 
value of Total bands tends to decrease over time

```{r}
#| label: tbl-total_WBC_by_visit
#| tbl-cap: Summary of total white blood cells (x1000/uL) for each visit time.
Final_Visits %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_WBC, na.rm = TRUE),4),
            sd= round(sd(Total_WBC, na.rm = TRUE),4),
            min = min(Total_WBC, na.rm = TRUE),
            max= max(Total_WBC, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```


#### neutrophils (neuts)
@tbl-total_neutrophil_by_visit provides the mean and standard deviation of 
**Total neutrophil** for each visit time. 

```{r}
#| label: tbl-total_neutrophil_by_visit
#| tbl-cap: Summary of total neutrophil for each visit time.

Final_Visits %>%
  group_by(Visit_Number) %>%
  summarise(mean = round(mean(Total_Neutrophil, na.rm = TRUE),4),
            sd= round(sd(Total_Neutrophil, na.rm = TRUE),4),
            min = min(Total_Neutrophil, na.rm = TRUE),
            max= max(Total_Neutrophil, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling() 
```

### Bandemia grade  

```{r}
A=matrix(0,nrow = 6,ncol=4)
colnames(A)=c("Bandemia grade","Value according to MSUVDL reference range (k/uL)","Numbers","Proportion")
A[,1]=c(0:5)
A[,2]=c("0.0-0.10","0.11-0.29","0.30-0.59","0.60-1.00",">1.00","death")
A[,3]=c(sum(Final_Visits$Total_Bands_Rescaled<=1),
        sum(Final_Visits$Total_Bands_Rescaled>1&Final_Visits$Total_Bands_Rescaled<3),
        sum(Final_Visits$Total_Bands_Rescaled>=3&Final_Visits$Total_Bands_Rescaled<6),
        sum(Final_Visits$Total_Bands_Rescaled>=6&Final_Visits$Total_Bands_Rescaled<=10),
        sum(Final_Visits$Total_Bands_Rescaled>10),0)
A=as.data.frame(A)
A$Numbers=as.numeric(A$Numbers)
A[,4]=A[,3]/609
A %>% 
  kable(digits = 3) %>%
  kable_styling()
```

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
