---
title: Generalized Linear Mixed Model Analysis of CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
bibliography: references.bib
csl: apa.csl
params:
  SourceFile: "GLMM_Analysis.qmd"
  LogFile: "GLMM_Analysis.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    output-file: "GLMM_Analysis.html"
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
---

# Purpose
This file is part of a research compendium [@Pierce-RN8714]. It runs a
generalized linear mixed model (GLMM) on data about dogs receiving CHOP
chemotherapy at MSU Veterinary Medicine Clinic (VMC) between 2016 and 2022. The
study aims to assess the incidence of bandemia (i.e., band neutrophil elevation)
in canine lymphoma patients undergoing CHOP chemotherapy and test predictors of
bandemia incidence. Specifically, we examine the effects of two predictors: the
drug administered and the dog's body surface area (BSA). Some results here are
used in a manuscript [@Eliason-RN8603].

# Setup
This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Methods. 

## Define Global Options
Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages
Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(devtools)    # for session_info()
library(here)        # for here(), i_am()
library(tidyverse)   # loads dplyr, forcats, ggplot2, lubridate, purrr, readr, 
                     # stringr, tibble, & tidyr by default. 
                     # for filter(), group_by(), mutate(), rename(), summarise()
library(broom)       # for glance(), tidy()
library(broom.mixed) # for support for mixed models. 
library(DHARMa)      # for plotQQunif(), plotResiduals(), simulateResiduals(), 
                     # testDispersion()
library(glmmTMB)     # for diagnose(), glmmTMB()
library(marginaleffects)  # for predictions(), comparisons(), slopes(), etc. 
library(modelsummary) # for modelsummary()
library(performance) # for check_model(), icc(), model_performance(), r2(), 
                     # test_lrt()
library(knitr)       # for kable()
library(kableExtra)  # for kable_styling 
library(piercer)     # for for brier(), display_num(), file_details(), 
                     # git_report(), R2Dev()
library(quarto)      # for quarto_version()
library(CHOPStudy)   # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Descriptive_Analysis.qmd")
```

## Load Data {#sec-Load-Data}

::: {.callout-caution}
Render the `scripts/Import_Data.qmd` file to create the data file loaded below. 
:::

@tbl-imported-data-file shows meta-data about the data file we will load.

```{r}
#| label: tbl-imported-data-file
#| tbl-cap: "Meta-Data About the Imported Data File Loaded"

DataFile <- here("data/Imported_CHOP_Data.RData")

file_details(DataFile) %>% 
  kable(, format = "html", 
        col.names = c("File Name", "Size", "Last Modified")) %>% 
  kable_styling() 
```

```{r}
#| label: load-data

load(file=here("data/Imported_CHOP_Data.RData"))
```

# Methods {#sec-Methods}
Each dog in the sample was undergoing CHOP chemotherapy, which involves
administering different drugs according to a multi-phase protocol that stretches
over multiple clinic visits. This analysis uses data from visits 1-10 during
phase 1 of the protocol. The protocol specifies which drug should be
administered at each visit, but there are deviations from the protocol based on
clinical circumstances and professional judgement. At each visit, a CBC sample 
was collected and analyzed to measure the outcome. 

We have a multilevel data structure with `r nrow(Final_Visits)` visits and the 
associated CBC samples (level 1 [L1] units) nested within 
`r nrow(Final_Patients)` dogs (level 2 [L2] units). The number of visits per dog 
varying from `r min(Final_Patients$N_Visits)` to 
`r max(Final_Patients$N_Visits)` (mean = 
`r round(mean(Final_Patients$N_Visits), digits = 1)`, SD = 
`r round(sd(Final_Patients$N_Visits), digits = 1)`). 

Due to the high correlation between weight and body surface area (BSA), we have 
included only BSA in the model. 

## Measures
There are no missing data in the measures used in this analysis (see descriptive
analysis output file). We centered the predictors to facilitate obtaining 
unconflated estimates of the effects of variables at each level of analysis
[@Rights-RN8274; @Rights-RN8755; @Schielzeth-RN1814; @Yaremych-RN8330]. 

### Bandemia (`Bandemia`)
Our outcome for the logistic models is a binary variable measuring whether the 
CBC sample taken at a given visit exhibited bandemia, which was defined as 
having total band neutrophils $\gt$ 0.10 k/ul and coded (0 = No, 1 = Yes). This 
means that we can interpret estimated marginal means on the response scale as 
the incidence rate of bandemia. 

### Body Surface Area (`BSA_m2_C`)
Body surface area at the start of CHOP (`Surface_Area_m2`) was measured in 
square meters ($m^2$) at the beginning of the CHOP cycle. This is a
patient-level (L2) measure that was invariant across visits. Prior to modeling,
we grand-mean centered this variable to improve interpretatibility of model
coefficients [@Schielzeth-RN1814], yielding `BSA_m2_C`.

### Prior Drug Administered (`DrugC_C`, `DrugD_C`, `M_DrugC_C`, `M_DrugD_C`)
The drug administered prior to CBC sample collection (`Drug`) was treated as a 
nominal, categorical variable with three levels: 1, vincristine (the reference 
levels); 2, cyclophosphamide; and 3, doxorubicin. This was a visit-level (L1) 
variable. It was dummy coded into two variables: `DrugC` codes for whether the 
drug was cyclophosphamide (0 = no, 1 = yes) and `DrugD` codes for whether the
drug was doxorubicin (0 = no, 1 = yes). 

The dummy codes were aggregated to create L2 (patient-specific) means (`M_DrugC` 
and `M_DrugD`). These means were added to the GLMM as L2 predictors so we can 
properly control for drug when testing the effect of BSA at L2 
[@Rights-RN8274; @Rights-RN8755; @Yaremych-RN8330]. 

While the practice is unfamiliar to many researchers, centering L1 dummy-coded
categorical predictors around their group means is necessary for obtaining 
unconflated estimates of their effects in multilevel models [@Yaremych-RN8330].
Accordingly, we computed L1 group-mean centered versions of the L1 predictors, 
yielding `DrugC_C = DrugC - M_DrugC` and `DrugD_C = DrugD - M_DrugD`. The 
centered variables were used in the models. 

## Statistical Analysis
Generalized linear mixed models (GLMMs) with a logit link function were employed
to assess the effects of each drug and BSA on presence of bandemia. This model
was chosen to capture the random effects pertaining to repeated observations
from each dog while dealing with the binary outcome. We set the Type I error
rate at $\alpha = 0.05$. We used the glmmTMB package for R to fit the models
[@Brooks-RN8667; @McGillycuddy-RN8874].

The basic modeling strategy involved fitting a set of three nested models. We
use likelihood ratio tests to compare models in sequence.

* Model 0 (`m0`): The null model includes only a fixed intercept and a random 
  intercept. The ICC for this model shows us what proportion of variability in 
  bandemia is attributable to individual differences between dogs rather than 
  to differences between visits for the same dog. 
* Model 1 (`m1`): This model expands on Model 0 by adding fixed effect slopes 
  for the L1 predictors (centered dummy codes for the drug administered, named 
  `DrugC_C` for cyclophosphamide and `DrugD_C` for doxorubicin). It still 
  includes both fixed and random intercepts. 
* Model 2 (`m2`): This model expands on Model 1 by adding the fixed effect 
  slopes for the L2 predictors. The aggregated means of the L1 drug dummy codes 
  (`M_DrugC` and `M_DrugD`) allow us to control for the fact that there was 
  variability in the proportion visits at which each dog was exposed to each 
  drug. We want to control for that when testing the effect of centered BSA. The 
  model retains L1 effects for the centered drug dummy codes plus the fixed and 
  random intercept terms. 

:::{.callout-important}
Our manuscript [@Eliason-RN8603] will only report Model 2, but Models 0 and 1
are useful as part of a model building process. They help us better understand
the data.
:::

### Model Fit Indices
We used the performance package [@Lüdecke-RN8788] to obtain a variety of common 
model fit indices for GLMMs, including AIC, BIC, ICC, and RMSE. In addition, it 
also reports the conditional and marginal $R^2$ [@Nakagawa-RN3697].

### Effect Size Estimates
Interpretation was facilitated by using the `marginaleffects` package 
[@Arel-Bundock-RN8812] to obtain estimated marginal means for bandemia incidence
associated with each drug along with relevant 95% confidence intervals, 
risk differences (RD) for bandemia when comparing pairs of drugs, and graphing 
average predicted bandemia incidence as a function of each continuous predictor. 
We also estimated the odds ratio (OR) for each predictor along with 95% 
confidence intervals.

### Collinearity
The `performance` package [@Lüdecke-RN8788] provides a function for examining 
the generalized variance inflation factor (GVIF) [@Fox-RN8789] for terms in 
regression models. We checked for multicollinearity problems by examining the
GVIF and a transformation of it that adjusts for categorical terms that use
multiple degrees of freedom [@Fox-RN3484]. In general, one can compare 
$GVIF^{(1/(2*df))}$ to cutoffs defined for $\sqrt{VIF}$. Some suggested cutoffs 
indicating potential problems are $\sqrt{VIF} \ge 2$ or $\sqrt{VIF} \ge 3$. We 
applied the latter cutoff, so hope to see $GVIF^{(1/(2*df))} < 3$ in our models.

### Model Diagnostics 
We used the `performance` package to do several model diagnostics. First, we 
used a posterior predictive check comparing model predictions to the observed 
data. Second, we examined binned residuals by plotting the average residual 
within a bin against the estimated probability of bandemia. Ideally, all points 
on that plot should fall within the model's error bounds. 

The `DHARMa` package in R uses simulation-based methods to compute scaled
residuals that should follow a uniform distribution if the data are consistent
with the actual distribution used to model the outcome, making it easier to
check for deviations from model assumptions [@Hartig-RN8668]. Third, we examined  
quantile-quantile plots of those residuals to check uniformity. As usual with a 
Q-Q plot, data points should lay close to the reference line for the theoretical 
distribution. 

Finally, we examined a Q-Q plot of the random effects to see whether they 
followed a normal distribution. 

We applied all of these diagnostic tests to the models below, summarizing the
results in figures. 

# Results {#sec-Results}

## Model 0 (No Predictors)
@tbl-m0  shows the parameter estimates for Model 0. 

```{r}
#| label: tbl-m0 
#| tbl-cap: Model 0 Parameters
#| results: asis

# Null model.
m0 <- glmmTMB(Bandemia ~ 1 + (1 | Patient), 
              data = Final_Visits, family = binomial(link = "logit"), 
              ziformula = ~ 0, dispformula = ~ 1, REML = FALSE)

FN = paste0("Log Likelihood = ", round(logLik(m0), digits = 3),
            ". Deviance = ", round(-2*logLik(m0), digits = 3),
            ". Model df = ", length(m0$fit$par), 
            ". Residual df = ", nobs(m0) - length(m0$fit$par),
            ". 95% CIs reported.")

modelsummary(list("Model 0" = m0), output = "kableExtra", 
             align = "llrrrrr", escape = FALSE,
             col.names = c("Component", "Term", "Est.", "CI.LL", "CI.UL", "z", "p"),
             shape = component + term ~ model + statistic, 
             statistic = c("conf.int", "statistic", "p.value"), 
             gof_omit = "R2 Marg.|R2 Cond.|ICC",
             notes = FN) %>% 
  cat()
```

Before we do much to interpret the model, we do some quick diagnostic checks. 
The chunk below should return text describing the result of some checks. The 
warning it provides is a result of the low incidence of bandemia in this sample. 

```{r}
#| label: diag-m0

diagnose(m0)
```

@fig-check-m0 displays some additional graphs used to check model assumptions. 
While the binned residuals do indicate minor issues with the model especially 
for situations where the probability of bandemia is very low or very high, the 
other panels don't show any cause for major concern. 

```{r}
#| label: fig-check-m0
#| fig-cap: Diagnostic Checks on Model 0
#| fig-height: 7
#| fig-width: 8

check_model(m0, type = "discrete_both", residual_type = "simulated",
            check = c("pp_check", "binned_residuals", "qq", "reqq"),
            base_size = 9)
```

@tbl-m0-icc shows the intraclass correlation (ICC). This is actually the main 
reason we ran Model 0. Focus on the adjusted ICC. It shows the proportion of 
variance in the bandemia that is attributable to individual differences between 
dogs. 

```{r}
#| label: tbl-m0-icc
#| tbl-cap: Model 0 Intraclass Correlations

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(26485) 

# Bootstrapped 95% CIs via lme4::bootMer()
icc(m0, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(ICC = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```
Because Model 0 does not use any substantive predictors, we do not expect its 
fixed effects to explain any variance. Thus, an marginal R2 of 0 is expected. 
The conditional R2 includes variance explained by both fixed and random effects,
so we expect it to be higher than the marginal R2 (usually by an amount 
equal to the adjusted ICC). 

```{r}
#| label: tbl-m0-r2
#| tbl-cap: Model 0 R2 Measures

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(44512) 

# Bootstrapped 95% CIs via lme4::bootMer()
r2(m0, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(R2 = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```

Another thing to examine is the population marginal mean for bandemia incidence 
for a typical dog (random effect of 0), which is shown in @tbl-m0-emm-intercept.

```{r}
#| label: tbl-m0-emm-intercept
#| tbl-cap: Model 0 Population-Average Estimated Marginal Mean (EMM) for Grand Mean

avg_predictions(m0, newdata = Final_Visits, conf_level = 0.95, re.form = NA) %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(estimate, std.error, conf.low, conf.high, statistic, p.value, s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("EMM", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2, "Wald Test" = 2, " "))
```

## Model 1 (L1 Predictors)
@tbl-m1 shows the parameters for Model 1 and @tbl-m1-OR shows the corresponding 
odds-ratios. 

```{r}
#| label: tbl-m1 
#| tbl-cap: Model 1 Parameters
#| results: asis

# Add L1 predictors.
m1 <- glmmTMB(Bandemia ~ DrugC_C + DrugD_C + (1 | Patient),
              data = Final_Visits, family = binomial(link = "logit"),
              ziformula = ~ 0, dispformula = ~ 1, REML = FALSE)

# m1 <- glmmTMB(Bandemia ~ Name + (1 | Patient), 
#               data = Final_Visits, family = binomial(link = "logit"), 
#               ziformula = ~ 0, dispformula = ~ 1, REML = FALSE)


FN = paste0("Log Likelihood = ", round(logLik(m1), digits = 3),
            ". Deviance = ", round(-2*logLik(m1), digits = 3),
            ". Model df = ", length(m1$fit$par), 
            ". Residual df = ", nobs(m1) - length(m1$fit$par),
            ". Drug reference level is Vincristine",
            ". 95% CIs reported.")

modelsummary(list("Model 1" = m1), output = "kableExtra", 
             align = "llrrrrr", escape = FALSE,
             col.names = c("Component", "Term", "Est.", "CI.LL", "CI.UL", "z", "p"),
             shape = component + term ~ model + statistic, 
             statistic = c("conf.int", "statistic", "p.value"), 
             gof_omit = "R2 Marg.|R2 Cond.|ICC",
             notes = FN) %>% 
  cat()
```

```{r}
#| label: tbl-m1-OR 
#| tbl-cap: Model 1 Odds-Ratios

FN = paste0("95% CIs reported.")

modelsummary(list("Model 1" = m1), output = "kableExtra", 
             align = "llrrr", escape = FALSE, exponentiate = TRUE,
             col.names = c("Component", "Term", "OR", "OR.LL", "OR.UL"),
             shape = component + term ~ model + statistic, 
             statistic = c("conf.int"), 
             gof_map = "none",
             notes = list(FN)) 
```

Before we do much to interpret the model, we do some quick diagnostic checks. 
The chunk below should return text describing the result of some checks. The 
warning it provides is a result of the low incidence of bandemia in this sample. 
Because we only added predictors that are dummy codes for drug (a categorical
variable), there is no need to check for multicollinearity.

```{r}
#| label: diag-m1

diagnose(m1)
#check_collinearity(m1)
```

@fig-check-m1 displays some additional graphs used to check model assumptions. 
While the binned residuals do indicate minor issues with the model especially 
for situations where the probability of bandemia is very low or very high, the 
other panels don't show any cause for major concern. 

```{r}
#| label: fig-check-m1
#| fig-cap: Diagnostic Checks on Model 1
#| fig-height: 7
#| fig-width: 8

check_model(m1, type = "discrete_both", residual_type = "simulated",
            check = c("pp_check", "binned_residuals", "qq", "reqq"),
            base_size = 9)
```

Next, we want to see whether Model 1 fit the data better than Model 0. The
resulting LRT in @tbl-compare-m0-m1 shows that accounting for which drug was
used prior to the visit improves the model.

```{r}
#| label: tbl-compare-m0-m1
#| tbl-cap: Likelihood Ratio Test Comparing Models 0 vs 1.

#compare_performance(m1, m2)

test_lrt(m0, m1) %>% 
  mutate(Model = case_when(Name == "m0" ~ "No Predictors",
                           Name == "m1" ~ "L1 Predictors")) %>% 
  kable(digits = 3, row.names = FALSE) %>% 
  kable_styling() 
```

@tbl-m1-icc shows that we still have a similar level of unexplained variability 
between dogs. That is expected because the visit-level drug predictors were 
group-mean centered. They can explain variance between visits for the same dog 
but cannot explain any between-dog variance. 

```{r}
#| label: tbl-m1-icc
#| tbl-cap: Model 1 Intraclass Correlations

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(99487) 

# Bootstrapped 95% CIs via lme4::bootMer()
icc(m1, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(ICC = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```

@tbl-m1-r2 Shows the R2 measures. The marginal R2 tells us how much of the 
overall variance in bandemia is explained by which drug the dog received prior 
to the visit's CBC sample being taken. 

```{r}
#| label: tbl-m1-r2
#| tbl-cap: Model 1 R2 Measures

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(32156) 

# Bootstrapped 95% CIs via lme4::bootMer()
r2(m1, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(R2 = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```

@tbl-m1-emm-intercept shows that the population marginal mean for bandemia 
incidence for a typical dog (random effect of 0) remains similar to that 
observed in Model 0. That's what we would expect. 

```{r}
#| label: tbl-m1-emm-intercept
#| tbl-cap: Model 1 Population-Average Estimated Marginal Mean (EMM) for Grand Mean

avg_predictions(m1, newdata = Final_Visits, conf_level = 0.95, re.form = NA) %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(estimate, std.error, conf.low, conf.high, statistic, p.value, s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("EMM", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2, "Wald Test" = 2, " "))
```

@tbl-m1-emm-name shows the population average estimated marginal means for 
bandemia incidence associated with each drug. The estimates assume a typical dog 
(i.e., one with a random intercept of 0). 

```{r}
#| label: tbl-m1-emm-name
#| tbl-cap: Model 1 Population-Average Estimated Marginal Mean (EMM) by Drug 
#|          Main Effect at Level 1

avg_predictions(m1, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                by = c("Name"))  %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(Name, estimate, std.error, conf.low, conf.high, statistic, p.value, 
         s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("Name","EMM", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 3, "95% CI" = 2, "Wald Test" = 2, " "))
```



## Model 2 (L1 & L2 Predictors)
@tbl-m2 shows the parameters for Model 2, which includes all the variables 
we intend to test and @tbl-m2-OR shows the corresponding odds-ratios. 

```{r}
#| label: tbl-m2 
#| tbl-cap: Model 2 Parameters
#| results: asis

# Add L2 predictors
m2 <- glmmTMB(Bandemia ~ DrugC_C + DrugD_C + M_DrugC + M_DrugD + BSA_m2_C + 
                (1 | Patient), 
              data = Final_Visits, family = binomial(link = "logit"), 
              ziformula = ~ 0, dispformula = ~ 1, REML = FALSE)

FN = paste0("Log Likelihood = ", round(logLik(m2), digits = 3),
            ". Deviance = ", round(-2*logLik(m2), digits = 3),
            ". Model df = ", length(m2$fit$par), 
            ". Residual df = ", nobs(m2) - length(m2$fit$par),
            ". Drug reference level is Vincristine",
            ". 95% CIs reported.")

modelsummary(list("Model 2" = m2), output = "kableExtra", 
             align = "llrrrrr", escape = FALSE,
             col.names = c("Component", "Term", "Est.", "CI.LL", "CI.UL", "z", "p"),
             shape = component + term ~ model + statistic, 
             statistic = c("conf.int", "statistic", "p.value"), 
             gof_omit = "R2 Marg.|R2 Cond.|ICC",
             notes = list(FN, paste("95% CIs reported.")),
             footnote_as_chunk = TRUE) %>% 
  cat()
```

```{r}
#| label: tbl-m2-OR 
#| tbl-cap: Model 2 Odds-Ratios

FN = paste0("95% CIs reported.")

modelsummary(list("Model 2" = m2), output = "kableExtra", 
             align = "llrrr", escape = FALSE, exponentiate = TRUE,
             col.names = c("Component", "Term", "OR", "OR.LL", "OR.UL"),
             shape = component + term ~ model + statistic, 
             statistic = c("conf.int"), 
             gof_map = "none",
             notes = FN) 
```

Before we do much to interpret the model, we do some quick diagnostic checks. 
The chunk below should return text describing the result of some checks. The 
warning it provides is a result of the low incidence of bandemia in this sample. 
We also check for multicollinearity, but find it is not a problem in this model. 

```{r}
#| label: diag-m2

diagnose(m2)
check_collinearity(m2)
```

@fig-check-m1 displays some additional graphs used to check model assumptions. 
While the binned residuals do indicate minor issues with the model especially 
for situations where the probability of bandemia is very low or very high, the 
other panels don't show any cause for major concern. 

```{r}
#| label: fig-check-m2
#| fig-cap: Diagnostic Checks on Model 2
#| fig-height: 7
#| fig-width: 8

check_model(m2, type = "discrete_both", residual_type = "simulated",
            check = c("pp_check", "binned_residuals", "qq", "reqq"),
            base_size = 9)
```

@tbl-compare-m0-m1-m2 shows comparisons between Models 0, 1, and 2. It is clear 
that Model 1 is an improvement over Model 0, and that Model2 is an improvement 
over model 1. 

```{r}
#| label: tbl-compare-m0-m1-m2
#| tbl-cap: Likelihood Ratio Tests Comparing Models 0, 1, and 2.

bind_rows(test_lrt(m0, m2), 
          test_lrt(m1, m2)) %>% 
  mutate(Model = case_when(Name == "m0" ~ "No Predictors",
                           Name == "m1" ~ "L1 Predictors",
                           Name == "m2" ~ "L1 & L2 Predictors")) %>% 
  kable(digits = 3, row.names = FALSE) %>% 
  kable_styling() %>% 
  group_rows(., group_label = "Model 0 vs Model 2", start_row = 1, end_row = 2) %>% 
  group_rows(., group_label = "Model 1 vs Model 2", start_row = 3, end_row = 4) 
```

:::{.callout-important}
We should interpret Model 2 instead of Model 1. 
:::

@tbl-m2-icc shows that we still have a similar level of unexplained variability 
between dogs. This is unexpected given that BSA appears to be a meaningful 
predictor of bandemia in @tbl-m2. 

```{r}
#| label: tbl-m2-icc
#| tbl-cap: Model 2 Intraclass Correlations

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(74156) 

# Bootstrapped 95% CIs via lme4::bootMer()
icc(m2, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(ICC = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```

@tbl-m2-r2 shows the marginal R2 for Model 2. 

```{r}
#| label: tbl-m2-r2
#| tbl-cap: Model 2 R2 Measures

# Set random number seed to ensure reproducibility of bootstrap result. 
set.seed(84361) 

# Bootstrapped 95% CIs via lme4::bootMer()
r2(m2, ci = 0.95, ci_method = NULL, robust = TRUE, iterations = 100) %>% 
  as.data.frame() %>% 
  t() %>% 
  data.frame() %>% 
  rename(R2 = X1, LL = X2, UL = X3) %>% 
  head(n = 2) %>% 
  kable(digits = 2, format.args = list(nsmall = 2)) %>% 
  kable_styling() %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2))
```

@tbl-m2-emm-intercept shows that the population marginal mean for bandemia 
incidence for a typical dog (random effect of 0) remains similar to that 
observed in Model 0. That's what we would expect. 

```{r}
#| label: tbl-m2-emm-intercept
#| tbl-cap: Model 2 Population-Average Estimated Marginal Mean (EMM) for Grand Mean

avg_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA) %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(estimate, std.error, conf.low, conf.high, statistic, p.value, s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("EMM", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 2, "95% CI" = 2, "Wald Test" = 2, " "))
```

Dogs were exposed to different drugs prior to each visit, so one key objective 
was to examine the bandemia incidence rates associated with each drug. 
@tbl-m1-emm-name shows the population average estimated marginal means for 
bandemia incidence associated with each drug. The estimates assume a typical dog 
(i.e., one with a random intercept of 0). @fig-m2-emm-name displays those means. 

```{r}
#| label: tbl-m2-emm-name
#| tbl-cap: Model 2 Population-Average Estimated Marginal Mean (EMM) for 
#|          Bandemia by Drug (Main Effect at Level 1)

avg_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                by = c("Name"))  %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(Name, estimate, std.error, conf.low, conf.high, statistic, p.value, 
         s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("Name","EMM", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 3, "95% CI" = 2, "Wald Test" = 2, " "))
```

```{r}
#| label: fig-m2-emm-name
#| tbl-cap: Model 2 Population-Average Estimated Marginal Means (EMMs) for 
#|          Bandemia by Drug (Main Effect at Level 1)
#| fig-width: 4
#| fig-height: 2

Fig3 <- plot_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                by = c("Name"), vcov = TRUE) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 0.25)) +
  labs(x = "Drug name", y = "Bandemia") 

Fig3

# Export figure for use in journal submission
ggsave(filename = here(paste0("scripts/output/Figure_3.png")), plot = Fig3, 
       device = "png", width = 4, height = 2, units = "in", dpi = 600)
```

::: {.callout-important}
@fig-m2-emm-name will be Figure 3 in the manuscript we are preparing
[@Eliason-RN8603].
:::

The simplest way to compare the bandemia rates is by examining the risk 
difference between each pair of drugs, as shown in @tbl-m2-RD-name. The risk
difference (RD) is simply subtracting one rate from the other. 

```{r}
#| label: tbl-m2-RD-name
#| tbl-cap: Model 2 Population-Average Bandemia Risk Differences Between 
#|          Different Drugs 

# Omit warning message about specifying hypothesis tests using positional 
# indices like b1, b2, etc. I've verified that this is correct. 
options(marginaleffects_safe = FALSE)

avg_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                hypothesis = c("Vincristine - Cyclophosphamide" =  "b1 - b2 = 0",
                               "Vincristine - Doxorubicin" =  "b1 - b3 = 0",
                               "Doxorubicin - Cyclophosphamide" =  "b3 - b2 = 0"), 
                by = "Name")  %>% 
  mutate(p.value = display_num(p.value)) %>% 
  select(hypothesis, estimate, std.error, conf.low, conf.high, statistic, p.value, 
         s.value) %>% 
  kable(digits = 2, format.args = list(nsmall = 2), align = "r",
        col.names = c("Hypothesis","RD", "SE", "LL", "UL", "z", "p", "s")) %>% 
  kable_styling(full_width = FALSE) %>% 
  add_header_above(header = c(" " = 3, "95% CI" = 2, "Wald Test" = 2, " "))
```

:::{.callout-caution}
In @tbl-m2-RD-name the p-value for the difference in bandemia rates between 
doxorubicin and cyclophoshamide is not significant but the lower CI bound is 
just barely below 0. We can't conclude there is a difference between those 
bandemia rates even though the point estimate for the risk difference 
matches that of the risk difference between vincristine and cyclophoshamide. 
That's because we have less data about doxorubicin than vincristine, so there is
more uncertainty about its marginal mean and differences between it and other
those of drugs. You can see that in the wider confidence interval for
doxorubicin in @fig-m2-emm-name. 
:::

In this study BSA is a patient-level variable rather than visit-level variable 
because only BSA at the start of CHOP was recorded and used in Model 2. 
@fig-m2-BSA shows the marginal effect of BSA on bandemia incidence. The 
incidence decreases considerably as BSA increases. The predicted value for the
smallest dog in the sample is around 19%, while for dogs of average size
(centered BSA values of 0) the rate is just under 10%, and for the largest dog
it is around 2.5%.

```{r}
#| label: fig-m2-BSA
#| tbl-cap: Model 2 Population-Average Predicted Bandemia by Body Surface Area 
#|          Main Effect at Level 2
#| fig-width: 4
#| fig-height: 3

Fig2 <- plot_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                condition = c("BSA_m2_C"), vcov = TRUE) +
  scale_x_continuous(name = expression("Mean-centered body surface area (" *  m^2 * ")"),
                     sec.axis = sec_axis(~ . + mean(Final_Patients$Surface_Area_m2), 
                                         name = expression("Uncentered body surface area (" * m^2 * ")"))) +
  geom_rug(aes(x = Surface_Area_m2 - mean(Surface_Area_m2)), 
               data = Final_Patients) +
  theme_bw() +
  labs(y = "Bandemia")

Fig2

# Export figure for use in journal submission
ggsave(filename = here(paste0("scripts/output/Figure_2.png")), plot = Fig2, 
       device = "png", width = 4, height = 3, units = "in", dpi = 600)
```

::: {.callout-important}
@fig-m2-BSA will be Figure 2 in the manuscript we are preparing 
[@Eliason-RN8603].
:::

@fig-m2-M-DrugC and @fig-m2-M-DrugD respectively show the marginal effects of 
mean exposure to cyclophosphamide and doxorubicin. Here, mean exposure refers to 
the proportion of visits at which the indicated drug was administered. These 
plots are consistent with @tbl-m2, which shows that neither of those variables 
had noteworthy effects on bandemia incidence. 

```{r}
#| label: fig-m2-M-DrugC
#| tbl-cap: Model 2 Population-Average Predicted Bandemia by Mean Exposure to 
#|          Clyclophosphamide Main Effect at Level 2
#| fig-width: 4
#| fig-height: 3

plot_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                condition = c("M_DrugC"), vcov = TRUE) +
  geom_rug(aes(x = M_DrugC), data = Final_Patients) +
  theme_bw() +
  coord_cartesian(ylim = c(0.00, 0.80)) +
  labs(x = "Mean Exposure to Cyclophosphamide (Proportion)",
       y = "Bandemia")
```

```{r}
#| label: fig-m2-M-DrugD
#| tbl-cap: Model 2 Population-Average Predicted Bandemia by Mean Exposure to 
#|          Doxorubicin Main Effect at Level 2
#| fig-width: 4
#| fig-height: 3

plot_predictions(m2, newdata = Final_Visits, conf_level = 0.95, re.form = NA, 
                condition = c("M_DrugD"), vcov = TRUE) +
  geom_rug(aes(x = M_DrugD), data = Final_Patients) +
  theme_bw() +
  coord_cartesian(ylim = c(0.00, 0.30)) +
  labs(x = "Mean Exposure to Doxorubicin (Proportion)",
       y = "Bandemia")
```

# References
::: {#refs}
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
