---
title: Generalized Linear Mixed Model Analysis of CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
bibliography: references.bib
csl: apa.csl
params:
  SourceFile: "GLMM_Analysis.qmd"
  LogFile: "GLMM_Analysis.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    output-file: "GLMM_Analysis.html"
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
---

# Purpose
This file is part of a research compendium [@Pierce-RN8714]. It runs a
generalized linear mixed model (GLMM) on data about dogs receiving CHOP
chemotherapy at MSU Veterinary Medicine Clinic (VMC) between 2016 and 2022. The
study aims to assess the incidence of bandemia (i.e., band neutrophil elevation)
in canine lymphoma patients undergoing CHOP chemotherapy and test predictors of
bandemia incidence. Specifically, we examine the effects of two predictors: the
drug administered and the dog's body surface area (BSA).

# Setup
This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Methods. 

## Define Global Options
Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages
Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(devtools)    # for session_info()
library(here)        # for here(), i_am()
library(tidyverse)   # loads dplyr, forcats, ggplot2, lubridate, purrr, readr, 
                     # stringr, tibble, & tidyr by default. 
                     # for filter(), group_by(), mutate(), rename(), summarise()
library(broom)       # for glance(), tidy()
library(broom.mixed) # for support for mixed models. 
library(car)         # for vif()
library(DHARMa)      # for plotQQunif(), plotResiduals(), simulateResiduals(), 
                     # testDispersion()
library(emmeans)     # for contrast(), emmeans(), pairs()
library(glmmTMB)     # for diagnose(), glmmTMB()
library(marginaleffects)  # for predictions(), comparisons(), slopes(), etc. 
library(modelsummary) # for modelsummary()
library(performance) # for model_performance(), r2()
library(knitr)       # for kable()
library(kableExtra)  # for kable_styling 
library(piercer)     # for for brier(), display_num(), file_details(), 
                     # git_report(), R2Dev()
library(quarto)      # for quarto_version()
library(CHOPStudy)   # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Descriptive_Analysis.qmd")
```

## Load Data {#sec-Load-Data}

::: {.callout-caution}
Render the `scripts/Import_Data.qmd` file to create the data file loaded below. 
:::

@tbl-imported-data-file shows meta-data about the data file we will load.

```{r}
#| label: tbl-imported-data-file
#| tbl-cap: "Meta-Data About the Imported Data File Loaded"

DataFile <- here("data/Imported_CHOP_Data.RData")

file_details(DataFile) %>% 
  kable(, format = "html", 
        col.names = c("File Name", "Size", "Last Modified")) %>% 
  kable_styling() 
```

```{r}
#| label: load-data

load(file=here("data/Imported_CHOP_Data.RData"))
```

# Methods {#sec-Methods}
Each dog in the sample was undergoing CHOP chemotherapy, which involves
administering different drugs according to a multi-phase protocol that stretches
over multiple clinic visits. This analysis uses data from visits 1-10 during
phase 1 of the protocol. The protocol specifies which drug should be
administered at each visit, but there are deviations from the protocol based on
clinical circumstances and professional judgement. At each visit, a CBC sample 
was collected and analyzed to measure the outcome. 

The dataset includes `r nrow(Final_Patients)` dogs and `r nrow(Final_Visits)`, 
with the number of visits per dog varying from `r min(Final_Patients$N_Visits)` 
to `r max(Final_Patients$N_Visits)` (
mean = `r round(mean(Final_Patients$N_Visits), digits = 1)`, 
SD = `r round(sd(Final_Patients$N_Visits), digits = 1)`) . Thus we have a 
multilevel data structure: with visits (and the associated CBC samples) at level
1 (L1) nested within dogs at level 2 (L2).

Due to the high correlation between weight and body surface area (BSA), we have 
included only BSA in the model. 

## Measures
There are no missing data in the measures used in this analysis (see descriptive
analysis output file). We centered the predictors to facilitate obtaining 
unconflated estiamtes of the effects of variables at each level of analysis
[@Rights-RN8274; @Rights-RN8755; @Schielzeth-RN1814; @Yaremych-RN8330]. 

### Bandemia (`Bandemia`)
Our outcome for the logistic models is a binary variable measuring whether the 
CBC sample taken at a given visit exhibited bandemia, which was defined as 
having total band neutrophils $\ge$ 0.20 k/ul and coded (0 = No, 1 = Yes). This 
means that we can interpret estimated marginal means on the response scale as 
the incidence rate of bandemia. 

### Body Surface Area (`BSA_m2_C`)
Body surface area at the start of CHOP (`Surface_Area_m2`) was measured in 
square meters ($m^2$). This is a patient-level (L2) measure that was invariant 
across visits. Prior to modeling, we grand-mean centered this variable to 
improve interpretatibility of model coefficients [@Schielzeth-RN1814], yielding 
`BSA_m2_C`.

### Prior Drug Administered (`DrugC_C`, `DrugD_C`, `M_DrugC_C`, `M_DrugD_C`)
The drug administered prior to CBC sample collection (`Drug`) was treated as a 
nominal, categorical variable with three levels: 1, vincristine (the reference 
levels); 2, cyclophosphamide; and 3, doxorubicin. This was a visit-level (L1) 
variable. It was dummy coded into two variables: `DrugC` codes for whether the 
drug was cyclophosphamide (0 = no, 1 = yes) and `DrugD` codes for whether the
drug was doxorubicin (0 = no, 1 = yes). 

The dummy codes were aggregated to create L2 (patient-specific) means (`M_DrugC` 
and `M_DrugD`). These means were added to the GLMM as L2 predictors so we can 
properly control for drug when testing the effect of BSA at L2 
[@Rights-RN8274; @Rights-RN8755; @Yaremych-RN8330]. 

While the practice is unfamiliar to many researchers, centering L1 dummy-coded
categorical predictors around their group means is necessary for obtaining 
unconflated estimates of their effects in multilevel models [@Yaremych-RN8330].
Accordingly, we computed L1 group-mean centered versions of the L1 predictors, 
yielding `DrugC_C = DrugC - M_DrugC` and `DrugD_C = DrugD - M_DrugD`. 

## Statistical Analysis
Generalized linear mixed models (GLMMs) with a logit link function were employed
to assess the effects of each drug, BSA, and visit number on presence of
bandemia. This model was chosen to capture the random effects pertaining to
repeated observations from each dog while dealing with the binary outcome. We
set the Type I error rate at $\alpha = 0.05$.

The basic modeling strategy involved fitting a set of three nested models. 

* Model 0 (`m0`): The null model includes only a fixed intercept and a random 
  intercept. The ICC for this model shows us what proportion of variability in 
  bandemia is attributable to individual differences between dogs rather than 
  to differences between visits within dogs. 
* Model 1 (`m1`): This model expands on Model 0 by adding fixed effect slopes 
  for the L1 predictors (centered dummy codes for the drug administered). It 
  still includes both fixed and random intercepts. 
* Model 2 (`m2`): This model expands on Model 1 by adding the fixed effect 
  slopes for the L2 predictors. The aggregated means of the L1 drug dummy codes 
  allow us to control for drug effects while testing the effect of centrerd BSA. 
  The model retails L1 effects for the centered drug dummy codes plus the fixed 
  and random intercept terms. 

We use likelihood ratio tests to compare models in sequence. 

### Model Fit Indices
We used the performance package [@Lüdecke-RN8788] report a variety of common 
model fit indices for GLMMs, including AIC, BIC, ICC, and RMSE. In addition, we 
it also reports the conditional and marginal $R^2$ [@Nakagawa-RN3697].

### Effect Size Estimates
Interpretation was facilitated by using the emmeans package [@Lenth-RN8188] to 
obtain estimated marginal means for bandemia incidence associated with each drug 
along with relevant 95% confidence intervals, and by examining the odds ratios 
for pairwise comparisons of the drugs and for the BSA effects along with 95% 
confidence intervals. 

### Collinearity
The `performance` package [@Lüdecke-RN8788] provides a function for examining 
the generalized variance inflation factor (GVIF) [@Fox-RN8789] for terms in 
regression models. We checked for multicollinearity problems by examining the
GVIF and a transformation of it that adjusts for categorical terms that use
multiple degrees of freedom [@Fox-RN3484]. In general, one can compare 
$GVIF^{(1/(2*df))}$ to cutoffs defined for $\sqrt{VIF}$. Some suggested cutoffs 
indicating potential problems are $\sqrt{VIF} \ge 2$ or $\sqrt{VIF} \ge 3$. We 
applied the latter cutoff, so hope to see $GVIF^{(1/(2*df))} < 3$ in our models.

### Residual Distributions 
The `DHARMa` package in R uses simulation-based methods to compute scaled
residuals that should follow a uniform distribution if the data are consistent
with the actual distribution used to model the outcome, making it easier to
check for deviations from model assumptions [@Hartig-RN8668]. We used those 
residuals to run several tests of model assumptions. 

First, we conducted a Kolmogorov-Smirnov (K-S) test of overall uniformity of the 
residual distribution, pairing it with a quantile-quantile (Q-Q) plot of the 
observed versus expected residual values. As usual with a Q-Q plot, data points 
should lay close to the reference line for the theoretical distribution. 

Second, we split the data by threshold and used both a K-S test of within group
deviation from uniformity and a Levene test for homogeneity of variance, pairing
the result with a a set of box plots showing the residual distribution within a
specific level of threshold. That plot should show the upper and lower edges of
the boxes laying close to the dashed lines for the 25th and 75th quantiles, and
the median bar of the box laying at the 50th quantile. Heterogeneity of variance
would be evident if the boxes associated with different thresholds vary
substantially in height.

Third, we conducted two-sided binomial tests for outliers on either end of the
residual distribution. The accompanying plot is a histogram of the residual 
distribution, annotated with red lines for any outliers detected. 

Finally, we ran a simulation-based a test for over- or under-dispersion in the
the scaled residuals. This is paired with a histogram of the distribution of 
simulated dispersion values, with a red line marking the position of the 
observed dispersion parameter value. 

We applied all of these `DHARMa` tests to the models below, summarizing the
results in figures. For all of the `DHARMa` tests, good models should produce
large $p$-values (sometimes displayed on plots as just `n.s.`) because small
values (e.g., $p < 0.05$) indicate violation of model assumptions.

# Results {#sec-Results}

```{r}
#|label: fit-models

# Null model.
m0 <- glmmTMB(Bandemia ~ 1 + (1 | Patient), 
              data = Final_Visits, family = binomial(link = "logit"), 
              ziformula = ~ 0, dispformula = ~ 1)

# Add L1 predictors. 
m1 <- glmmTMB(Bandemia ~ DrugC_C + DrugD_C + (1 | Patient), 
              data = Final_Visits, family = binomial(link = "logit"), 
              ziformula = ~ 0, dispformula = ~ 1)

# Add L2 predictors
m2 <- glmmTMB(Bandemia ~ DrugC_C + DrugD_C + M_DrugC + M_DrugD + BSA_m2_C + 
                (1 | Patient), 
              data = Final_Visits, family = binomial(link = "logit"), 
              ziformula = ~ 0, dispformula = ~ 1)
```


```{r}
#|label: model-perf

compare_performance(m0, m1, m2)
test_lrt(m0, m1, m2)

diagnose(m0)
check_model(m0)

diagnose(m1)
check_model(m1)
check_collinearity(m1)

diagnose(m2)
check_model(m2)
check_collinearity(m2)
```


#### Results

```{r}
#|label: summary_model_logistic
summary(m1)
```

In the model, Drug 1 is set as the baseline. Drug 2 is found to significantly
decrease the probability of causing Bandemia compared to the baseline (Drug 1).

@fig-EMMs_with_CI_logit shows the estimated means of logit(Bandemia) for each
drug. And it is calculated assuming the Surface_area equals the mean of
Surface_area ($0.9470$), and the visit number is 1.

```{r}
#| label: fig-EMMs_with_CI_logit
#| fig-cap: Estimated Marginal Means of logit(Bandemia) for each drug

m1.emm <- emmeans(m1, ~ Drug)

emm_df <- as.data.frame(m1.emm)

emm_df %>% 
  ggplot(., aes(x = Drug, y = emmean)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Means:logit(Bandemia)") +
  ggtitle("EMMs with Confidence Intervals")
```

@tbl-logit_bandemia lists the values used to generate the figure above.

```{r}
#| label: tbl-logit_bandemia
#| tbl-cap: Estimated Means of logit(bandemia) for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)

emm_df%>%kable()%>%kable_styling()
```

@fig-Estimated_Marginal_Proportions shows the estimated means of proportion of
Bandemia for each drug. And it is calculated assuming the Surface_area equals
the mean of Surface_area ($0.9470$), and the visit number is 1.

```{r}
#| label: fig-Estimated_Marginal_Proportions
#| fig-cap: Estimated Marginal Proportions of Bandemia
emm <- emmeans(model, ~ Drug,type="response")
emm_df <- as.data.frame(emm)
ggplot(emm_df, aes(x = Drug, y = prob)) +
  geom_point() +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.1) +
  xlab("Drug") +
  ylab("Estimated Marginal Proportions: Bandemia") +
  ggtitle("Estimated values with Confidence Intervals")
```

@tbl-bandemia lists the values used to generate the figure above.

```{r}
#| label: tbl-bandemia
#| tbl-cap: Estimated Proportions of Bandemia for each Drug

emm_df[,c(2,3,5,6)]=round(emm_df[,c(2,3,5,6)],4)
emm_df%>%kable()%>%kable_styling()
```

@tbl-odds_ratio presents the odds ratios comparing between the drugs. The odds
ratio of 3.2610 indicates that Drug 1 is significant (pvalue=0.003<0.5)
associated with a higher likelihood of the outcome compared to Drug 2, while
Drug 2 has an odds ratio of 0.2987, suggesting a significant (pvalue=0.018<0.5)
lower likelihood of the outcome as compared to Drug 3.

```{r}
#| label: tbl-odds_ratio
#| tbl-cap: odds ratios comparing between the drugs
emm <- emmeans(model, ~ Drug)

pairwise <- pairs(emm)
contrast <- contrast(emm, method = "revpairwise")

rr <- summary(pairwise, infer = c(TRUE, TRUE), type = "response")
rr=rr[,-4]
rr=rr[,-6]
colnames(rr)=c("contrast","rate ratio","SE","lower CI","upper CI","z.ratio","p.value")
rr[,c(2:7)]=round(rr[,c(2:7)],4)

rr%>%kable()%>%kable_styling()
```

We get similar results from these two models. In summary, the likelihood of
Bendemia occurrence with drug 2 is significantly lower compared to its
occurrence with drugs 1 and 3. Additionally, a negative association was found
between surface area and the total bands value, with this association being
statistically significant.

# References
::: {#refs}
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
