---
title: Import and Manage CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
bibliography: references.bib
csl: apa.csl
params:
  SourceFile: "Import_Data.qmd"
  LogFile: "Import_Data.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    output-file: "Import_Data_Draft.html"
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true
---

# Purpose
This file is part of a research compendium [@Pierce-RN8714]. It imports raw 
data provided by Alison Masyr and Caitlin Eliason that contains observations of
dogs receiving CHOP chemotherapy at MSU Veterinary Medicine Clinic (VMC) between
2016 and 2022. This file does some data cleaning and management, then saves out
revised data files that will be used for the analyses in other scripts. These 
data will be used in a manuscript [@Eliason-RN8603]. 

# Setup
This section documents some setup tasks that are useful to the statistician on 
the team. Most readers of this document will probably want skip directly to 
@sec-Data-Mgt.

## Define Global Options
Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages
Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(devtools)     # for session_info()
library(dplyr)      # for filter(), group_by(), mutate(), rename(), summarise()
library(here)       # for here(), i_am()
library(knitr)      # for kable()
library(kableExtra) # for kable_styling 
library(labelled)   # for set_variable_labels()
library(readxl)     # for read_xlsx()
library(piercer)    # for file_details(), git_report()
library(quarto)     # for quarto_version()
library(CHOPStudy)  # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Import_Data.qmd")
```

## Read Data {#sec-Read-Data}
@tbl-raw-data-file shows meta-data about the raw data file we will import.

```{r}
#| label: tbl-raw-data-file
#| tbl-cap: "Meta-Data About Raw Data File"

# Store path to raw data file. 
RawFile <- here("scripts/extdata/Vinc_Cases_2025-12-05.xlsx")

file_details(RawFile) %>% 
  kable(, format = "html", 
        col.names = c("File Name", "Size", "Last Modified")) %>% 
  kable_styling()
```

Below we read in the data file and then list the names of all variables in that 
file. 

```{r}
#| label: import-data

Raw_Visits <- read_xlsx(RawFile) 
names(Raw_Visits)
```

# Data Management {#sec-Data-Mgt}
The first task is just renaming variables. After that, we also need to convert 
some variables to factors and compute some additional variables. 

The units of measurement for `Total_Bands` are 1000/ul (k/ul) and some are not
integers (they have non-zero digits to the right of the decimal place).
Therefore, we rescaled into a new variable that has units of 100/ul by computing
`Total_Bands_Rescale = Total_Bands*10`. That rescaled variable was used in 
draft count models we ultimately abandoned. We decided to focus on a logistic 
model instead of the count model. The variables and some associated descriptive
analyses remain in the repository, but we never added the count model to the
repository. 

We also created log transformed versions of both `Total_Bands` and 
`Total_Bands_Rescaled` for use in graphing data. Because `log(0) = -Infinity`, 
we recode the `-Infinity` values for `log_Total_Bands` to -5 and the `-Infinity` 
values for `log_Total_Bands_Rescaled` to -3 so that we can make compact graphs 
of the distributions. Do not use those two variables for descriptive statistics 
or models because they don't correctly represent the infinite values properly
for other purposes. These variables and associated descriptive analyses are also 
no longer part of our manuscript plans but we just left the code for them in 
the data management and descriptive analyses. 

The code below does some data management, then lists the names of variables 
again. 

```{r}
#| label: create-All-Visits

All_Visits <- Raw_Visits %>% 
  mutate(Patient = factor(Patient),
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7,
         Total_Bands_Rescaled = Total_Bands*10,
         # Special recoded log variables for use only in graphs! 
         log_Total_Bands = if_else(Total_Bands == 0, 
                                   true = -5, 
                                   false = log(Total_Bands)),
         log_Total_Bands_Rescaled = if_else(Total_Bands_Rescaled == 0, 
                                   true = -3, 
                                   false = log(Total_Bands_Rescaled)),
         # Binary transformations of Total_Bands. 
         Bands_Present = if_else(Total_Bands > 0, true = 1, false = 0),
         Bandemia = if_else(Total_Bands > 0.1, true = 1, false = 0),
         # Categorize Total_Bands. 
         Band_Category = factor(case_when(Total_Bands == 0 ~ 1,
                                          Total_Bands <= 0.1 ~ 2,
                                          Total_Bands > 0.1 ~ 3),
                                levels = 1:3, 
                                labels = c("Absent (0.0 k/ul)", 
                                           "Present (<= 0.1 k/ul)",
                                           "Bandemia (> 0.1 k/ul)")),
         # Recode Toxic_Change
         Toxic_Change = if_else(Toxic_Change == "yes", true = 1, false = 0),
         # Band:Neutrophil ratio transformations. 
         BN_Ratio = Total_Bands/Total_Neutrophil,
         High_BN_Ratio = if_else(BN_Ratio >= 1, true = 1, false = 0),
         # Add drug name
         Name = factor(Drug, levels = 1:3, 
                       labels = c("Vincristine", "Cyclophosphamide", 
                                  "Doxorubicin")),
         # Add neutered variable, then fix Sex variable. 
         Neutered = case_when(Sex %in% c("F","M") ~ "Intact",
                              Sex == "N" ~ "Neutered",
                              Sex == "S" ~ "Spayed",
                              .default = NA),
         Sex = case_when(Sex == "N" ~ "M",
                         Sex == "S" ~ "F",
                         .default = Sex),
         # Fix handling of missing data.
         Stage = if_else(Stage == "N/A", 
                         true = NA, 
                         false = Stage),
         Substage = if_else(Substage == "N/A", 
                            true = NA, 
                            false = Substage),
         Phenotype = if_else(Phenotype == "N/A", 
                             true = NA, 
                             false = Phenotype),
         # Fix variable type and treat "not given" as Dose = 0. 
         Dose_mgpm2 = if_else(Dose_mgpm2 == "not given", 
                              true = 0, 
                              false = as.numeric(Dose_mgpm2))) %>% 
  arrange(Patient, CHOP_Number, Visit_Number, CBC_Date)

# Show the updated variable names to confirm all looks right. 
names(All_Visits)
```

::: {.callout-note}
In previous versions of the output (2024-04-02 and earlier), we had been 
recoding Patient_Number values > 28 by subtracting 1 from them just to make the 
whole set of numbers sequential. We removed that code because it makes no 
functional difference to our analyses and it is easier to cross-reference back 
to source data file if we don't do that. 
:::

## Check Patient and MRN Variables
I assumed that `Patient` and `MRN` variables were created such that these two
variables have a one-to-one correspondence. Each `Patient` value should always
be paired with the same `MRN` value and vice versa. If that is true, then the
number of unique combinations will be the same as the number of unique values
for each variable separately and then it doesn't matter which of those two
variables we use to identify a given dog.

```{r}
#| label: check-Patient-MRN

tibble(Patients_N = All_Visits %>% select(Patient) %>% n_distinct(), 
       MRNs_N = All_Visits %>% select(MRN) %>% n_distinct(),
       Combinations_N = All_Visits %>% select(Patient, MRN) %>% n_distinct(),
       Match = if_else(Patients_N == MRNs_N & Patients_N == Combinations_N, 
                       true = TRUE, false = FALSE))
```

::: {.callout-tip}
If `Match = TRUE` in the result immediately above, that confirms `Patient` and 
`MRN` variables are both unique identifiers for a dog and have a one-to-one 
correspondence. 
:::

## Check Weeks Since CHOP Started
Here, we examine when a CBC lab sample was collected via the `Weeks` variable, 
which measures how many weeks had elapsed between the start of the CHOP protocol 
and collection of the CBC sample. This should always be a positive number, so 
any negative values are implausible and should be corrected if possible. Any 
extreme positive values would also be a problem.

@tbl-Visit-Timing1 shows descriptive statistics for the visit timing broken down
by visit number. We're looking for aggregate statistics that don't make sense
and could be caused by data entry errors in the `CBC_Date` variable.

```{r}
#| label: tbl-Visit-Timing1 
#| tbl-cap: Descriptive Statistics for Visit Timing (Weeks Since CHOP Started)

All_Visits %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

:::{.callout-tip}
Note: The mean, SD, minimum and maximum number of weeks since CHOP started are
all reasonable and make sense given the expected timing of events under the CHOP
protocol.
:::

@fig-visualize-weeks is a boxplot that visualizes the distribution of `Weeks` 
separately for each `Visit_Number`. This illustrates that there are some 
outliers where the visit was held long after it would normally be scheduled 
relative to the start of the CHOP cycle. We will later subset the data to deal 
with that. 

```{r}
#| label: fig-visualize-weeks
#| fig-cap: Weeks Since CHOP Started by Visit (Before Case Selection)

boxplot(Weeks ~ Visit_Number, data = All_Visits, xlab="Visit Number")
```

## Check Uniqueness of Visit Records
Next we verify that each row in the visit-level data (`All_Visits`) is a unique 
combination of a patient (`Patient` and `MRN`), a CHOP cycle (`CHOP_Number`), 
and a visit number (`Visit_Number`). 

```{r}
#| label: check-visit-uniqueness

# We expect a value of TRUE for Matching_N.
tibble(All_Visits_N = nrow(All_Visits),
       Unique_Visits_N = All_Visits %>% 
         select(MRN, CHOP_Number, Visit_Number) %>% 
         n_distinct(),
       Matching_N = All_Visits_N == Unique_Visits_N)
```

::: {.callout-tip}
If `Matching_N = TRUE` in the result immediately above, that confirms the 
uniqueness of each visit record. 
:::

## Create Patient-Level Dataset
The `All_Visits` dataset has multiple rows per patient because the CHOP protocol 
involves multiple visits, each of which is recorded as a separate visit number. 
We need a patient-level data file to correctly get descriptive statistics on 
patient weight and body surface area. 

```{r}
#| label: create-All-Patients

All_Patients <- All_Visits %>% 
  arrange(Patient, CHOP_Number, Visit_Number) %>% 
  select(Patient, MRN, In_Final_Patients, Age_Yrs, Sex, Neutered, Breed, 
         Diagnosis, Stage, Substage, Phenotype, Weight_kg, Surface_Area_m2) %>% 
  group_by(Patient, MRN) %>% 
  slice_head(n = 1) %>% 
  ungroup()
```

Next, we verify that the number of rows in `All_Patients` is equal to the number 
of unique MRNs in `All_Visits`. If that is true, then all rows in `All_Visits` 
for a given `MRN` must have the exact same values for both weight and surface 
area. 

```{r}
#| label: check-patient-N

# We expect a value of TRUE for Matching_N.
tibble(All_MRNs_N = All_Visits %>% select(MRN) %>% n_distinct(),
       All_Patients_N = nrow(All_Patients),
       Matching_N = All_MRNs_N == All_Patients_N)
```

::: {.callout-tip}
If `Matching_N = TRUE` in the result immediately above, that confirms number of 
rows in `All_Patients` is equal to the number of unique MRNs in `All_Visits`. 
:::

## Create More Variables
After reading peer review feedback on a manuscript that got rejected, the team 
decided on 2024-08-29 to narrow the focus of the analyses. We will now focus 
only on data for patients who had their first visit within 10 days of 
their CHOP start date. This section creates variables we can use to implement 
that decision.

First we create a subset of the data containing only records for visit 1 within 
the respective CHOP cycles, then add a logical variable marking which visit 1 
observations had `CBC_Dates` indicating that the lab sample was collected within 
10 days of the start of that CHOP cycle. 

``` {r}
#| label: create-Visit1-CHOP

Visit1_CHOP <- All_Visits %>% 
  filter(Visit_Number == 1) %>% 
  mutate(Started_in_10d = Days_Since_Start <= 10) %>% 
  select(Patient, MRN, CHOP_Number, Started_in_10d)
```

Next we merge that new `Started_in_10d` variable back onto the `All_Visits` 
dataset. We can then use it later to select cases. 

``` {r}
#| label: merge-All-Visits-Visit1-CHOP

All_Visits <- left_join(x = All_Visits, y = Visit1_CHOP, 
                      by = join_by(Patient, MRN, CHOP_Number)) 
```

## Subset the Visit-Level  Data 
In this section, we begin a sequential filtering process to get to the subset of 
records we want to analyze. Each step adds another filter that drops records. 
Before we start filtering, we store information on the amount of raw data before 
any filtering. 

``` {r}
#| label: create-Step0

Step0 <- All_Visits %>% 
  count(Patient) %>% 
  summarise(Step = 0,
            Event = "Keep all records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

### By CHOP Cycle Number
In this step, we retain only records for the first CHOP cycle for each patient. 

``` {r}
#| label: create-CHOP1

CHOP1 <- All_Visits %>% 
  filter(CHOP_Number == 1)

Step1 <- CHOP1 %>% 
  count(Patient) %>% 
  summarise(Step = 1,
            Event = "Keep only records for CHOP cycle 1", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step1-drops
#| tbl-cap: Records Dropped in Step 1 Subsetting

All_Visits %>% 
  select(Patient, MRN, CHOP_Number) %>% 
  filter(CHOP_Number != 1) %>% 
  group_by(Patient, MRN, CHOP_Number) %>% 
  summarize(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Whether Visit 1 Occurred Within 10 Days of CHOP Cycle Starting
In this step, we retain only records for patients whose first visit during 
their first cycle started within 10 days of the CHOP cycle start date. 

``` {r}
#| label: create-CHOP1-Start10d

CHOP1_Start10d <- CHOP1 %>% 
  filter(Started_in_10d == TRUE)

Step2 <- CHOP1_Start10d %>% 
  count(Patient) %>% 
  summarise(Step = 2,
            Event = "Keep only if patient's visit 1 occurred within 10 days of CHOP start", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step2-drops
#| tbl-cap: Records Dropped in Step 2 Subsetting

CHOP1 %>% 
  filter(Started_in_10d == FALSE) %>% 
  group_by(Patient, MRN) %>% 
  summarise(Visits_Dropped = n(),
            CHOP1_Start_Date = min(CHOP_Start_Date),
            Visit1_CBC_Date = min(CBC_Date), 
            Visit1_Weeks = min(Weeks)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Phase of Treatment
In this step, we retain only records where the visit occurred during the first 
phase of treatment 

``` {r}
#| label: create-CHOP1-Start10d-Phase1

CHOP1_Start10d_Phase1 <- CHOP1_Start10d %>% 
  filter(Phase_of_Trt == 1)

Step3 <- CHOP1_Start10d_Phase1 %>% 
  count(Patient) %>% 
  summarise(Step = 3,
            Event = "Keep only Phase 1 records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step3-drops
#| tbl-cap: Records Dropped in Step 3 Subsetting

CHOP1_Start10d %>% 
  filter(Phase_of_Trt != 1) %>% 
  group_by(Patient, MRN, Phase_of_Trt) %>% 
  summarise(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Drug
Here, we retain only the records where the drug associated with the visit was  
drug 1 (vincristine), drug 2 (cyclophosphamide), or drug 3 (doxorubicin). The 
drug selection is motivated by the fact that we don't have enough data about 
other drugs.

``` {r}
#| label: create-CHOP1-Start10d-Phase1-Drug123

CHOP1_Start10d_Phase1_Drug123 <- CHOP1_Start10d_Phase1 %>% 
  filter(Drug %in% 1:3)

Step4 <- CHOP1_Start10d_Phase1_Drug123 %>% 
  count(Patient) %>% 
  summarise(Step = 4,
            Event = "Keep only Drug 1, 2, & 3 records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step4-drops
#| tbl-cap: Records Dropped in Step 3 Subsetting

CHOP1_Start10d_Phase1 %>% 
  filter(!Drug %in% 1:3) %>% 
  group_by(Patient, MRN, Drug) %>% 
  summarise(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### Rename Data Frame
Here we rename the data object to keep the object name shorter. 

```{r}
#| label: create-Final-Visits

Final_Visits <- CHOP1_Start10d_Phase1_Drug123
```

There are `r nrow(Final_Visits)` rows in `Final_Visits`. 

## Subset the Patient-Level Data
We aggregate data from `Final_Visits` to count how many visits each dog has so  
we can merge that into the final patient-level data.  

```{r}
#| label: create-Visit-Counts

Visit_Counts <- Final_Visits %>% 
  count(Patient) %>% 
  rename(N_Visits = n)
```

There are `r nrow(Visit_Counts)` rows in `Visit_Counts`. 

Next, we need to merge in a patient-level version of `Leukemic` because that 
should be time-invariant within a CHOP cycle.

```{r}
#| label: create-Leukemic

Leukemic <- Final_Visits %>% 
  select(Patient, Leukemic) %>% 
  unique() 
```

There are `r nrow(Leukemic)` rows in `Leuklemic`. 

Now we trim the `All_Patients` data down to just the patients included in the 
`Final_Visits` data frame and merge it with `Visit_Counts` and `Leuklemic`.

```{r}
#| label: create-Final-Patients

Final_Patients <- All_Patients %>% 
  filter(Patient %in% Final_Visits$Patient) %>% 
  left_join(x = ., y = Visit_Counts, by = join_by(Patient)) %>% 
  left_join(x = ., y = Leukemic, by = join_by(Patient))
```

There are `r nrow(Final_Patients)` rows in `Final_Patients`. 

# Summarize Amount of Data Remaining
Now @tbl-amt-data summarizes how much data there was at each step of record 
selection. 

```{r}
#| label: tbl-amt-data
#| tbl-cap: Amount of Data Available By Step in Case Selection

bind_rows(Step0, Step1, Step2, Step3, Step4) %>% 
  mutate(Mean_Obs_Per_Dog = N_Observations/N_Dogs) %>% 
  kable(digits = 1) %>% 
  kable_styling()
```

```{r}
#| label: fig-visualize-weeks-again
#| fig-cap: Weeks Since CHOP Started By Visit Number (After Case Selection)

boxplot(Weeks ~ Visit_Number, data = CHOP1_Start10d_Phase1_Drug123, 
        xlab="Visit Number")
```

```{r}
#| label: tbl-visit-drug
#| tbl-cap: Number of Records by Visit and Drug (After Case Selection)

xtabs(~Visit_Number + Drug, data = CHOP1_Start10d_Phase1_Drug123) %>% 
  addmargins() %>% 
  kable(col.names = c("Visit", "1 Vincristine", "2 Cyclophosphamide", 
                      "3 Doxorubicin", "Sum")) %>% 
  kable_styling() %>% 
  add_header_above(., header = c(" ", "Drug" = 3, " "))
```

# Update Visit-Level Data
Here we finally add some derived variables we will need for the planned GLMM 
analyses. 

```{r}
#| label: update-Final-Visits

Final_Visits <- Final_Visits %>% 
  # Dummy-code Drug (using Drug = 1 (vincristine) as reference level)
  mutate(DrugV = if_else(Drug == 1, true = 1, false = 0), 
         DrugC = if_else(Drug == 2, true = 1, false = 0),
         DrugD = if_else(Drug == 3, true = 1, false = 0)) %>% 
  # Turn on grouping by patient so we can easily get patient-level means.
  group_by(Patient) %>% 
  # Aggregate dummy codes to patient-level means for use in GLMM analyses. 
  mutate(M_DrugV = mean(DrugV), 
         M_DrugC = mean(DrugC),
         M_DrugD = mean(DrugD)) %>% 
  # Turn off grouping by patient (no longer need it)
  ungroup() %>% 
  # Create centered L1 predictors for use in GLMM analyses.
  mutate(DrugV_C = DrugV - M_DrugV,
         DrugC_C = DrugC - M_DrugC,
         DrugD_C = DrugD - M_DrugD) %>% 
  # Center the L2 predictors for use in GLMM analyses.
  mutate(BSA_m2_C = Surface_Area_m2 - mean(Final_Patients$Surface_Area_m2),
         Age_Yrs_C = Age_Yrs - mean(Age_Yrs)) %>% 
  set_variable_labels(DrugV = "Dummy code, Drug = Vincristine (L1, uncentered)",
                      DrugC = "Dummy code, Drug = Cyclophosphamide (L1, uncentered)",
                      DrugD = "Dummy code, Drug = Doxorubicin (L1, uncentered)",
                      M_DrugV = "Mean dummy code, Drug = Vincristine (L2, aggregated)",
                      M_DrugC = "Mean dummy code, Drug = Cyclophosphamide (L2, aggregated)",
                      M_DrugD = "Mean dummy code, Drug = Doxorubicin (L2, aggregated)",
                      DrugV_C = "Dummy code, Drug = Vincristine (L1, group mean centered)",
                      DrugC_C = "Dummy code, Drug = Cyclophosphamide (L1, group mean centered)",
                      DrugD_C = "Dummy code, Drug = Doxorubicin (L1, group mean centered)",
                      BSA_m2_C = "Body surface area in m^2 (L2, grand-mean centered)") 
```

# Update Patient-Level Data
We want the add the means for the dummy-coded drug variables to the patient data 
for use in some graphs later. 

```{r}
#| label: create-Drug-Means

Drug_Means <- Final_Visits %>% 
  select(Patient, M_DrugV, M_DrugC, M_DrugD) %>% 
  unique() %>% 
  set_variable_labels(M_DrugV = "Mean dummy code, Drug = Vincristine (L2, aggregated)",
                      M_DrugC = "Mean dummy code, Drug = Cyclophosphamide (L2, aggregated)",
                      M_DrugD = "Mean dummy code, Drug = Doxorubicin (L2, aggregated)") 
```

```{r}
#| label: update-Final-Patients

Final_Patients <- Final_Patients %>% 
  left_join(x = ., y = Drug_Means, by = join_by(Patient))
```

# Save Data

``` {r}
#| label: create-save-what

# Character vector of object names that will be saved.
save_what <- c("Raw_Visits", "All_Visits", "All_Patients", "Final_Visits", 
               "Final_Patients")
```

``` {r}
#| label: save-data
#| execute: true

DataFile <- here("data/Imported_CHOP_Data.RData")

save(list = save_what, file = DataFile)
```

@tbl-imported-data-file shows meta-data about the data file we just saved.

```{r}
#| label: tbl-imported-data-file
#| tbl-cap: "Meta-Data About the Imported Data File Saved Out"

file_details(DataFile) %>% 
  kable(, format = "html", 
        col.names = c("File Name", "Size", "Last Modified")) %>% 
  kable_styling() 
```

::: {.callout-tip}
The current data file can be loaded into R by copying the following code chunk 
into a script, setting the chunk option `eval: true`, and running the chunk. 
That will overwrite any objects already in memory that have the same names as 
the objects being loaded.

``` {r}
#| label: load-data
#| eval: false

load(file=here("data/Imported_CHOP_Data.RData"))
```
:::

# References
::: {#refs}
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
