---
title: Import and Manage CHOP Study Data
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceFile: "Import_Data.qmd"
  LogFile: "Import_Data.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
    cfsize: "scriptsize"
---

\FloatBarrier

# Purpose

This file provides the analysis results of data file 
`Vinc Cases 1.9.24.xlsx` provided by Alison and Caitlin, which records 
observations from dogs receiving CHOP chemotherapy at MSU VMC between 2016 and 
2022.

::: {.callout-note}
The data file we received was called `Vinc Cases 1.9.23.xlsx`, but we renamed it 
to `Vinc Cases 1.9.24.xlsx` before using it because that was the date we 
received it. 
:::

# Setup

This section of the report documents some setup tasks that are useful to the
statistician on the team. Most readers of this report will probably want skip
directly to @sec-Results.

## Define Global Options

Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages

Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(here)
library(kableExtra)
library(dplyr)
library(knitr)
library(gtsummary)
library(glmmTMB)
library(readxl)
library(quarto)
library(lme4)
library(emmeans)
library(ggplot2)
```

## Declare Path

```{r}
#| label: declare-path

i_am("Masyr_C1358_2024-05-03.qmd")
```

## Read Data {#sec-Read-Data}

```{r}
#| label: load-data

data_all <- read_xlsx(here("../Data/Vinc Cases 1.9.24.xlsx")) 
names(data_all)
```

## Data Management {#sec-Data-Mgt}

The units of measurement for `Total_Bands` are 1000/ul and some are not integers 
(they have non-zero digits to the right of the decimal place). We create the 
`Total_Bands_Rescale` that is rescaled to units of 100/ul by computing 
`Total_Bands_Rescale = Total_Bands*10`. That rescaled variable is intended for 
use in count models because those require non-negative integer values for all 
counts. We also created log transformed versions of both `Total_Bands` and 
`Total_Bands_Rescaled` below for use in graphing data. Because the 
`log(0) = -Infinity`, we recode the `-Infinity` values for `log_Total_Bands` to 
-5 and the `-Infinity` values for `log_Total_Bands_Rescaled` to - 3 so that we 
can make compact graphs of the distributions. 

```{r}
#| label: manage-data

data_all <- data_all %>% 
  rename(Patient_Number = `Patient Number`,
         Visit_Number = `Visit #`,
         CHOP_Start_Date = `CHOP start date`,
         Phase_of_Treatment = `Phase of treatment`,
         CHOP_Number = `CHOP number`,
         CBC_Date = `CBC date`,
         Total_WBC = `Total WBC (x10.e3/uL)`,
         Total_Neutrophil = `Total neutrophil`,
         Total_Bands = `Total bands`, 
         Bandemia_Grade = `Bandemia grade`,
         Weight_kg = kg,
         Surface_Area_m2 = `m^2`,
         Notes = `...15`) %>% 
  mutate(Patient_Number = factor(Patient_Number),
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7,
         Total_Bands_Rescaled = Total_Bands*10,
         log_Total_Bands = if_else(Total_Bands == 0, 
                                   true = -5, 
                                   false = log(Total_Bands)),
         log_Total_Bands_Rescaled = if_else(Total_Bands_Rescaled == 0, 
                                   true = -3, 
                                   false = log(Total_Bands_Rescaled)),
         Bandemia = if_else(Total_Bands >= 0.2, true = 1, false = 0))

# Show the updated variable names to confirm all looks right. 
names(data_all)
```

::: {.callout-note}
In previous versions of the output (04/02/2024 and earlier), Ruxin had been 
recoding Patient_Number values > 28 by subtracting 1 from them. Steve removed 
that code because it makes no functional difference to our analyses and it is 
easier to cross-reference back to source data file if we don't do that. 
:::

### Only include data from phase I of CHOP 1

For the following analysis, we would only include samples from phase I of CHOP 1.

```{r}
#| label: Phase-I-of-CHOP-1

data_chop1 <- data_all %>% filter(CHOP_Number == 1)

data_chop1_phase_1 <- data_chop1 %>% filter(Phase_of_Treatment==1)
```

### Only include data about drug 1, drug 2, drug 3

Given the sample size, we would only include results of drug 1 (vincristine), 
drug 2 (cyclophosphamide) and drug 3 (doxorubicin).

```{r}
#| label: drug123
data_analyse <- data_chop1_phase_1 %>% filter(Drug %in% 1:3)
#data_analyse=as.data.frame(data_analyse)
```

### Data Cleaning: Check for Implausible Visit Timing
Here, we examine the visit timing (in weeks since CHOP started) when a CBC lab 
sample was collected. We break this down by visit number. We're looking for 
aggregate statistics that don't make sense, such as negative numbers, which 
would imply that a lab visit occurred before CHOP started. That could be caused 
by data entry errors in date variables. 

## Visit Timing (Weeks Since CHOP Started) by Visit Number
@tbl-Visit-Timing shows descriptive statistics for the visit timing broken down 
by visit number. 

```{r}
#| label: tbl-Visit-Timing 
# Days from the start date

data_analyse %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable() %>% 
  kable_styling()
```


:::{.callout-caution}
Note: The mean, minimum, and maximum number of weeks since CHOP started are all 
**negative** values at Visit 3. That implies a data problem we need to 
investigate and fix. 
:::

So we find the individuals with weeks less than 0.

```{r}
#| label: individuals_with_weeks_less_than_0
data_analyse %>% 
  filter(Weeks < 0) %>% 
  kable() %>% 
  kable_styling()
```

The Patient_Number is 80. Print all the relevant records of that patient.

```{r}
#| label: patient_80
data_analyse[data_analyse$Patient_Number==80,] %>%
  kable() %>%
  kable_styling()
```

It should be **2020-01-30** based on the other records. So we corrected it.

```{r}
#| label: patient_80_correct
data_analyse$CBC_Date[data_analyse$Weeks < 0] = "2020-01-30"
# Days from the start date
```

Then we find the individuals with weeks larger than 10 at the third visit time.

```{r}
#| label: individuals_with_weeks_larger_than_10 
data_analyse %>% 
  filter(Weeks > 10 & Visit_Number==3) %>% 
  kable() %>% 
  kable_styling()
```

The outlier comes from Patient_Number 62. Print all the relevant records of that patient.

```{r}
#| label: patient_62
data_analyse %>% 
  filter(Patient_Number == 62) %>% 
  kable() %>% 
  kable_styling()
```

It should be **2018-5-25** based on the other records.

Now we summarize the time since first visit in weeks per visit again.

```{r}
#| label: time_since_first_visit_2
#| tbl-cap: Time since first visit in weeks per visit
data_analyse$CBC_Date[data_analyse$Weeks > 100] = "2018-05-25 UTC"

# Days from the start date
data_analyse <- data_analyse %>%
  mutate(Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7)

data_analyse %>%
  group_by(Visit_Number) %>%
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable() %>% 
  kable_styling()
```

@fig-visualize_weeks is a boxplot that visualize the information of weeks.

```{r}
#| label: fig-visualize_weeks
#| fig-cap: Time since first visit in weeks per visit
boxplot(Weeks ~ Visit_Number, data = data_analyse, xlab="Visit Number")
```

For your reference, we also find the individuals with weeks larger than 10 at 
the first visit time.

```{r}
#| label: weeks_larger_than_10_at_the_first
data_analyse %>% 
  filter(Weeks > 10 & Visit_Number == 1) %>%
  kable() %>% 
  kable_styling()
```

And printed the relevant records below.

```{r}
#| label: patient_65
data_analyse %>% 
  filter(Patient_Number == 65) %>%
  kable() %>%
  kable_styling()
```

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
