---
title: Import and Manage CHOP Study Data
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceFile: "Import_Data.qmd"
  LogFile: "Import_Data.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true
---

\FloatBarrier

# Purpose

This file imports raw data provided by Alison Masyr and Caitlin Eliason that 
contains observations of dogs receiving CHOP chemotherapy at MSU Veterinary 
Medicine Clinic (VMC) between 2016 and 2022. This file does some data cleaning 
and management, then saves out revised data files that will be used for the 
analyses in other scripts. 

::: {.callout-note}
The data file we received was called `Vinc Cases 1.9.23.xlsx`, but we renamed it 
to `Vinc Cases 1.9.24.xlsx` before using it because that was the date we 
received it. 
:::

# Setup

This section documents some setup tasks that are useful to the statistician on 
the team. Most readers of this document will probably want skip directly to 
@sec-Data-Mgt.

## Define Global Options

Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages

Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(here)       # for here(), i_am()
library(dplyr)      # for filter(), group_by(), mutate(), rename(), summarise()
library(knitr)      # for kable()
library(kableExtra) # for kable_styling 
library(gtsummary)  # for 
library(readxl)     # for read_xlsx()
library(quarto)     # for quarto_version()
library(CHOPStudy)  # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Import_Data.qmd")
```

## Read Data {#sec-Read-Data}

```{r}
#| label: import-data

Raw_Visits <- read_xlsx(here("../Data/Vinc Cases 1.9.24.xlsx")) 
names(Raw_Visits)
```

# Data Management {#sec-Data-Mgt}
The first task is just renaming variables. After that, we also need to convert 
some to factors and compute some additional variables. 

The units of measurement for `Total_Bands` are 1000/ul and some are not integers 
(they have non-zero digits to the right of the decimal place). Therefore, we 
rescaled into a new variable that has units of 100/ul by computing 
`Total_Bands_Rescale = Total_Bands*10`. That rescaled variable was used in 
count models because those require non-negative integer values for all counts. 
We also created log transformed versions of both `Total_Bands` and 
`Total_Bands_Rescaled` for use in graphing data. Because `log(0) = -Infinity`, 
we recode the `-Infinity` values for `log_Total_Bands` to -5 and the `-Infinity` 
values for `log_Total_Bands_Rescaled` to -3 so that we can make compact graphs 
of the distributions. Do not use those two variables for descriptive statistics 
or models because they don't correctly represent the infinite values properly
for other purposes. 

```{r}
#| label: create-All-Visits

All_Visits <- Raw_Visits %>% 
  rename(Patient = `Patient Number`,
         Visit_Number = `Visit #`,
         CHOP_Start_Date = `CHOP start date`,
         Phase_of_Treatment = `Phase of treatment`,
         CHOP_Number = `CHOP number`,
         CBC_Date = `CBC date`,
         Total_WBC = `Total WBC (x10.e3/uL)`,
         Total_Neutrophil = `Total neutrophil`,
         Total_Bands = `Total bands`, 
         Bandemia_Grade = `Bandemia grade`,
         Weight_kg = kg,
         Surface_Area_m2 = `m^2`,
         Notes = `...15`) %>% 
  mutate(Patient = factor(Patient),
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7,
         Total_Bands_Rescaled = Total_Bands*10,
         # Special recoded log variables for use only in graphs! 
         log_Total_Bands = if_else(Total_Bands == 0, 
                                   true = -5, 
                                   false = log(Total_Bands)),
         log_Total_Bands_Rescaled = if_else(Total_Bands_Rescaled == 0, 
                                   true = -3, 
                                   false = log(Total_Bands_Rescaled)),
         # Binary transformations of Total_Bands. 
         Bands_Present = if_else(Total_Bands > 0, true = 1, false = 0),
         Bandemia = if_else(Total_Bands >= 0.2, true = 1, false = 0),
         # Band:Neutrophil ratio transformations. 
         BN_Ratio = Total_Bands/Total_Neutrophil,
         High_BN_Ratio = if_else(BN_Ratio >= 1, true = 1, false = 0)) %>% 
  arrange(Patient, CHOP_Number, Visit_Number, CBC_Date)

# Show the updated variable names to confirm all looks right. 
names(All_Visits)
```

::: {.callout-note}
In previous versions of the output (2024-04-02 and earlier), we had been 
recoding Patient_Number values > 28 by subtracting 1 from them just to make the 
whole set of numbers sequential. We removed that code because it makes no 
functional difference to our analyses and it is easier to cross-reference back 
to source data file if we don't do that. 
:::

## Check Patient and MRN Variables
I assumed that that `Patient` and `MRN` variables were created such that these 
two variables have a one-to-one correspondence. Each `Patient` value should 
always be paired with the same `MRN` value and vice versa. If that is true, 
then the number of unique combinations will be the same as the number of unique 
values for each variable separately and then it doesn't matter which of those 
two variables we use to identify a given dog. 

```{r}
#| label: check-Patient-MRN

tibble(Patients_N = All_Visits %>% select(Patient) %>% n_distinct(), 
       MRNs_N = All_Visits %>% select(MRN) %>% n_distinct(),
       Combinations_N = All_Visits %>% select(Patient, MRN) %>% n_distinct(),
       Match = if_else(Patients_N == MRNs_N & Patients_N == Combinations_N, 
                       true = TRUE, false = FALSE))
```

## Fix Negative Values for Weeks Since CHOP Started
Here, we examine when a CBC lab sample was collected via the `Weeks` variable, 
which measures how many weeks had elapsed between the start of the CHOP protocol 
and collection of the CBC sample. This should always be a positive number, so 
any negative values are implausible and should be corrected if possible. 

@tbl-negative-weeks shows that there are two patients with `Weeks < 0`. The 
`CBC_Date` values are clearly implausible. Therefore we will examine data for 
those two patients more closely in @tbl-patient55 and @tbl-patient80. 

```{r}
#| label: tbl-negative-weeks
#| tbl-cap: Records with Implausible Visit Timing (Weeks < 0)

All_Visits %>% 
  filter(Weeks < 0) %>% 
  select(Patient, MRN, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

```{r}
#| label: tbl-patient55
#| tbl-cap: Visit Dates for Patient 55 (Before Corrections)

All_Visits %>% 
  filter(Patient == 55) %>%
  select(Patient, CHOP_Number, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable(digits = 2) %>%
  kable_styling()
```

```{r}
#| label: tbl-patient80
#| tbl-cap: Visit Dates for Patient 80 (Before Corrections)

All_Visits %>% 
  filter(Patient == 80) %>%
  select(Patient, CHOP_Number, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable(digits = 2) %>%
  kable_styling()
```

::: {.callout-caution}
Both patients 55 (@tbl-patient55) and 80 (@tbl-patient80) each have one visit 
with a clearly inaccurate date where the year was entered as 1930 instead of 
2020. For both dates, correcting the year while leaving the month and day 
portions of the date alone will correct the error and yield dates that make 
sense given the sequence and temporal spacing of other visits for that patient. 
:::

Below we fix the `CBC_Date` and then re-calculate variables that depended on it. 

```{r}
#| label: fix-patient55_80
All_Visits <- All_Visits %>% 
  mutate(CBC_Date = case_when(Patient == 55 & CBC_Date == as.POSIXct("1930-08-03", tz = "UTC") 
                              ~ as.POSIXct("2020-08-03", tz = "UTC"), 
                              Patient == 80 & CBC_Date == as.POSIXct("1930-01-30", tz = "UTC") 
                              ~ as.POSIXct("2020-01-30", tz = "UTC"),
                              .default = CBC_Date),
         # Re-compute variables that depend on CBC_Date. 
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7)
```

## Fix Extreme Positive Values for Weeks Since CHOP Started
We've fixed negative values for `Weeks`, but still need to look for any other 
outlier values that may be data quality problems. So, we look at a summary by
visit number. We're looking for aggregate statistics that don't make sense and 
could be caused by data entry errors in the `CBC_Date` variable. 
@tbl-Visit-Timing1 shows descriptive statistics for the visit timing broken down 
by visit number. 

```{r}
#| label: tbl-Visit-Timing1 
#| tbl-cap: Descriptive Statistics for Visit Timing (Weeks Since CHOP Started)

All_Visits %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

:::{.callout-caution}
Note: The mean, SD, and maximum number of weeks since CHOP started are all 
unexpectedly large at visit 3 and don't make sense given the expected timing of 
events under the CHOP protocol. That implies a data problem we need to 
investigate and fix. 
:::

@tbl-outlier-weeks shows the records for patient visits where `Weeks > 34`, 
which was just a bit higher than the the maximum number of weeks seen at any 
other visit number. We can see that Patient = 62 is a very clear outlier, with a 
`CBC_Date` in the future. That's clearly a data entry error. So, now we inspect 
data for `Patient = 62` more closely to determine if we can fix the anomalous 
date. @tbl-patient62 shows all the records for that patient. 

```{r}
#| label: tbl-outlier-weeks
#| tbl-cap: Visit 3 Records with Implausible Visit Timing (Weeks > 34)

All_Visits %>% 
  filter(Visit_Number == 3 & Weeks > 34) %>% 
  select(Patient, MRN, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

```{r}
#| label: tbl-patient62
#| tbl-cap: Visit Dates for Patient 62 (Before Corrections)

All_Visits %>% 
  filter(Patient == 62) %>%
  select(Patient, CHOP_Number, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable(digits = 2) %>%
  kable_styling()
```

::: {.callout-caution}
It is clear that for `Patient = 62`, visit 3 should have occurred in the year
2018 instead of 2028. We assume the month and day are accurate because after
correcting the year, the `CBC_Date` for visit 3 would fall between those for
visits 2 and 4. The sequence would then be sensible.
:::

Below we fix the `CBC_Date` and then re-calculate variables that depended on it. 

```{r}
#| label: fix-patient62
All_Visits <- All_Visits %>% 
  mutate(CBC_Date = if_else(Patient == 62 & CBC_Date == as.POSIXct("2028-05-25", 
                                                                   tz = "UTC"), 
                            true =  as.POSIXct("2018-05-25", tz = "UTC"), 
                            false = CBC_Date),
         # Re-compute variables that depend on CBC_Date. 
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7)
```

## Recheck for Outlier Values for Weeks Since CHOP Started
We've fixed negative values for Weeks, but still need to look for any other 
outlier values that may be data quality problems. So, we look at a summary by
visit number. We're looking for aggregate statistics that don't make sense and 
could be caused by data entry errors in the `CBC_Date` variable. 
@tbl-Visit-Timing2 shows descriptive statistics for the visit timing broken down 
by visit number. 

```{r}
#| label: tbl-Visit-Timing2 
#| tbl-cap: Updated Descriptive Statistics for Visit Timing (Weeks Since CHOP Started)

All_Visits %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

:::{.callout-tip}
Note: The descriptive statistics now look much more sensible. 
:::

@fig-visualize-weeks is a boxplot that visualizes the distribution of `Weeks` 
separately for each `Visit_Number`. This illustrates that there are some 
outliers where the visit was held long after it would normally be scheduled 
relative to the start of the CHOP cycle. We will later subset the data to deal 
with that. 

```{r}
#| label: fig-visualize-weeks
#| fig-cap: Weeks Since CHOP Started by Visit (Before Case Selection)

boxplot(Weeks ~ Visit_Number, data = All_Visits, xlab="Visit Number")
```

## Check Uniqueness of Visit Records
Next we verify that each row in the visit-level data (`All_Visits`) is a unique 
combination of a patient (`Patient` and `MRN`), a CHOP cycle (`CHOP_Number`), 
and a visit number (`Visit_Number`). 

```{r}
#| label: check-visit-uniqueness

# We expect a value of TRUE for Matching_N.
tibble(All_Visits_N = nrow(All_Visits),
       Unique_Visits_N = All_Visits %>% 
         select(MRN, CHOP_Number, Visit_Number) %>% 
         n_distinct(),
       Matching_N = All_Visits_N == Unique_Visits_N)
```

## Create Patient-Level Dataset
The `All_Visits` dataset has multiple rows per patient because the CHOP protocol 
involves multiple visits, each of which is recorded as a separate visit number. 
We need a patient-level data file to correctly get descriptive statistics on 
patient weight and body surface area. 

```{r}
#| label: create-All-Patients

All_Patients <- All_Visits %>% 
  select(Patient, MRN, Weight_kg, Surface_Area_m2) %>% 
  unique()
```

Next, we verify that the number of rows in `All_Patients` is equal to the number 
of unique MRNs in `All_Visits`. If that is true, then all rows in `All_Visits` 
for a given `MRN` must have the exact same values for both weight and surface 
area. 

```{r}
#| label: check-patient-N

# We expect a value of TRUE for Matching_N.
tibble(All_MRNs_N = All_Visits %>% select(MRN) %>% n_distinct(),
       All_Patients_N = nrow(All_Patients),
       Matching_N = All_MRNs_N == All_Patients_N)
```

## Create More Variables
After reading peer review feedback on a manuscript that got rejected, the team 
decided on 2024-08-29 to narrow the focus of the analyses. We will now focus 
only on data for patients who had their first visit within 2 weeks of 
their CHOP start date. This section creates variables we can use to implement 
that decision.

First we create a subset of the data containing only records for visit 1 within 
the respective CHOP cycles, then add a logical variable marking which visit 1 
observations had `CBC_Dates` indicating that the lab sample was collected within 
2 weeks of the start of that CHOP cycle. 

``` {r}
#| label: create-Visit1-CHOP

Visit1_CHOP <- All_Visits %>% 
  filter(Visit_Number == 1) %>% 
  mutate(Started_in_2Weeks = Weeks <= 2) %>% 
  select(Patient, MRN, CHOP_Number, Started_in_2Weeks)
```

Next we merge that new `Started_in_2Weeks` variable back onto the `All_Visits` 
dataset. We can then use it later to select cases. 

``` {r}
#| label: merge-All-Visits-Visit1-CHOP

All_Visits <- left_join(x = All_Visits, y = Visit1_CHOP, 
                      by = join_by(Patient, MRN, CHOP_Number)) 
```

## Subset the Visit-Level  Data 
In this section, we begin a sequential filtering process to get to the subset of 
records we want to analyze. Each step adds another filter that drops records. 
Before we start filtering, we store information on the amount of raw data before 
any filtering. 

``` {r}
#| label: create-Step0

Step0 <- All_Visits %>% 
  count(Patient) %>% 
  summarise(Step = 0,
            Event = "Keep all records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

### By CHOP Cycle Number
In this step, we retain only records for the first CHOP cycle for each patient. 

``` {r}
#| label: create-CHOP1

CHOP1 <- All_Visits %>% 
  filter(CHOP_Number == 1)

Step1 <- CHOP1 %>% 
  count(Patient) %>% 
  summarise(Step = 1,
            Event = "Keep only records for CHOP cycle 1", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step1-drops
#| tbl-cap: Records Dropped in Step 1 Subsetting

All_Visits %>% 
  select(Patient, MRN, CHOP_Number) %>% 
  filter(CHOP_Number != 1) %>% 
  group_by(Patient, MRN, CHOP_Number) %>% 
  summarize(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Whether Visit 1 Occurred Within 2 Weeks of CHOP Cycle Starting
In this step, we retain only records for patients whose first visit during 
their first cycle started within 2 weeks of the CHOP cycle start date. 

``` {r}
#| label: create-CHOP1-Start2W

CHOP1_Start2W <- CHOP1 %>% 
  filter(Started_in_2Weeks == TRUE)

Step2 <- CHOP1_Start2W %>% 
  count(Patient) %>% 
  summarise(Step = 2,
            Event = "Keep only if patient's visit 1 occurred within 2 weeks of CHOP start", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step2-drops
#| tbl-cap: Records Dropped in Step 2 Subsetting

CHOP1 %>% 
  filter(Started_in_2Weeks == FALSE) %>% 
  group_by(Patient, MRN) %>% 
  summarise(Visits_Dropped = n(),
            CHOP1_Start_Date = min(CHOP_Start_Date),
            Visit1_CBC_Date = min(CBC_Date), 
            Visit1_Weeks = min(Weeks)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Phase of Treatment
In this step, we retain only records where the visit occurred during the first 
phase of treatment 

``` {r}
#| label: create-CHOP1-Start2W-Phase1

CHOP1_Start2W_Phase1 <- CHOP1_Start2W %>% 
  filter(Phase_of_Treatment == 1)

Step3 <- CHOP1_Start2W_Phase1 %>% 
  count(Patient) %>% 
  summarise(Step = 3,
            Event = "Keep only Phase 1 records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step3-drops
#| tbl-cap: Records Dropped in Step 3 Subsetting

CHOP1_Start2W %>% 
  filter(Phase_of_Treatment != 1) %>% 
  group_by(Patient, MRN, Phase_of_Treatment) %>% 
  summarise(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### By Drug
Here, we retain only the records where the drug associated with the visit was  
drug 1 (vincristine), drug 2 (cyclophosphamide), or drug 3 (doxorubicin). The 
drug selection is motivated by the fact that we don't have enough data about 
other drugs.

``` {r}
#| label: create-CHOP1-Start2W-Phase1-Drug123

CHOP1_Start2W_Phase1_Drug123 <- CHOP1_Start2W_Phase1 %>% 
  filter(Drug %in% 1:3)

Step4 <- CHOP1_Start2W_Phase1_Drug123 %>% 
  count(Patient) %>% 
  summarise(Step = 4,
            Event = "Keep only Drug 1, 2, & 3 records", 
            N_Dogs = n(),
            N_Observations = sum(n))
```

The chunk below identified the list of patients removed by the filtering step 
above. This is just for reference in case the team wants to understand what 
happened there.

```{r}
#| label: tbl-Step4-drops
#| tbl-cap: Records Dropped in Step 3 Subsetting

CHOP1_Start2W_Phase1 %>% 
  filter(!Drug %in% 1:3) %>% 
  group_by(Patient, MRN, Drug) %>% 
  summarise(Visits_Dropped = n()) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

### Rename Data Frame

```{r}
#| label: create-Final-Visits

Final_Visits <- CHOP1_Start2W_Phase1_Drug123
```

There are `r nrow(Final_Visits)` rows in `Final_Visits`. 

## Subset the Patient-Level Data
We aggregate data from `Final_Visits` to count how many visits each dog has so  
we can merge that into the final patient-level data.  

```{r}
#| label: create-Visit-Counts

Visit_Counts <- Final_Visits %>% 
  count(Patient) %>% 
  rename(N_Visits = n)
```

There are `r nrow(Visit_Counts)` rows in `Visit_Counts`. 

Now we trim the `All_Patients` data down to just the patients included in the 
`Final_Visits` data frame and merge it with `Visit_Counts`. 

```{r}
#| label: create-Final-Patients

Final_Patients <- All_Patients %>% 
  filter(Patient %in% Final_Visits$Patient) %>% 
  left_join(x = ., y = Visit_Counts, by = join_by(Patient))
```

There are `r nrow(Final_Patients)` rows in `Final_Patients`. 

## Summarize Amount of Data Remaining
Now @tbl-amt-data summarizes how much data there was at each step of record 
selection. 

```{r}
#| label: tbl-amt-data
#| tbl-cap: Amount of Data Available By Step in Case Selection

bind_rows(Step0, Step1, Step2, Step3, Step4) %>% 
  mutate(Mean_Obs_Per_Dog = N_Observations/N_Dogs) %>% 
  kable(digits = 1) %>% 
  kable_styling()
```

```{r}
#| label: fig-visualize-weeks-again
#| fig-cap: Weeks Since CHOP Started By Visit Number (After Case Selection)

boxplot(Weeks ~ Visit_Number, data = CHOP1_Start2W_Phase1_Drug123, 
        xlab="Visit Number")
```

```{r}
#| label: tbl-visit-drug
#| tbl-cap: Number of Records by Visit and Drug (After Case Selection)

xtabs(~Visit_Number + Drug, data = CHOP1_Start2W_Phase1_Drug123) %>% 
  addmargins() %>% 
  kable(col.names = c("Visit", "1 vincristine", "2 cyclophosphamide", 
                      "3 doxorubicin", "Sum")) %>% 
  kable_styling() %>% 
  add_header_above(., header = c(" ", "Drug" = 3, " "))
```

# Save Data

``` {r}
#| label: create-save-what

# Character vector of object names that will be saved.
save_what <- c("Raw_Visits", "All_Visits", "All_Patients", "Final_Visits", 
               "Final_Patients")
```

``` {r}
#| label: save-data
#| execute: true

save(list = save_what, 
     file=here("data/Imported_CHOP_Data.RData"))
```

::: {.callout-tip}
The current data file can be loaded into R by copying the following code chunk 
into a script, setting the chunk option `eval: true`, and running the chunk. 
That will overwrite any objects already in memory that have the same names as 
the objects being loaded.

``` {r}
#| label: load-data
#| eval: false

load(file=here("data/Imported_CHOP_Data.RData"))
```
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
