---
title: Import and Manage CHOP Study Data
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Ruxin Shi
    orcid: 0009-0001-9483-6444
    email: shiruxin@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params:
  SourceFile: "Import_Data.qmd"
  LogFile: "Import_Data.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true 
    cfsize: "scriptsize"
---

\FloatBarrier

# Purpose

This file imports raw data provided by Alison Masyr and Caitlin Eliason that 
contains observations of dogs receiving CHOP chemotherapy at MSU Veterinary 
Medicine Clinic (VMC) between 2016 and 2022. This file does some data cleaning 
and management, then saves out revised data files that will be used for the 
analyses in other scripts. 

::: {.callout-note}
The data file we received was called `Vinc Cases 1.9.23.xlsx`, but we renamed it 
to `Vinc Cases 1.9.24.xlsx` before using it because that was the date we 
received it. 
:::

# Setup

This section documents some setup tasks that are useful to the statistician on 
the team. Most readers of this document will probably want skip directly to 
@sec-Data-Mgt.

## Define Global Options

Global R chunk options are defined in the YAML header but local chunk options
will over-ride global options. We can temporarily disable an individual chunk by
inserting `#| eval: false` on a line at the top of the chunk.

## Load Packages

Load R packages that we need to get additional functions.

```{r}
#| label: load-packages
library(here)       # for here(), i_am()
library(dplyr)      # for filter(), group_by(), mutate(), rename(), summarise()
library(knitr)      # for kable()
library(kableExtra) # for kable_styling 
library(gtsummary)  # for 
library(readxl)     # for read_xlsx()
library(quarto)     # for quarto_version()
library(CHOPStudy)  # for version info 
```

## Declare Path

```{r}
#| label: declare-path

here::i_am(path = "scripts/Import_Data.qmd")
```

## Read Data {#sec-Read-Data}

```{r}
#| label: import-data

Raw_CHOP <- read_xlsx(here("../Data/Vinc Cases 1.9.24.xlsx")) 
names(Raw_CHOP)
```

# Data Management {#sec-Data-Mgt}
The first task is just renaming variables. After that, we also need to convert 
some to factors and compute some additional variables. 

The units of measurement for `Total_Bands` are 1000/ul and some are not integers 
(they have non-zero digits to the right of the decimal place). Therefore, we 
rescaled into a new variable that has units of 100/ul by computing 
`Total_Bands_Rescale = Total_Bands*10`. That rescaled variable was used in  
count models because those require non-negative integer values for all counts. 
We also created log transformed versions of both `Total_Bands` and 
`Total_Bands_Rescaled` for use in graphing data. Because `log(0) = -Infinity`, 
we recode the `-Infinity` values for `log_Total_Bands` to -5 and the `-Infinity` 
values for `log_Total_Bands_Rescaled` to -3 so that we can make compact graphs 
of the distributions. 

```{r}
#| label: create-All_CHOP

All_CHOP <- Raw_CHOP %>% 
  rename(Patient = `Patient Number`,
         Visit_Number = `Visit #`,
         CHOP_Start_Date = `CHOP start date`,
         Phase_of_Treatment = `Phase of treatment`,
         CHOP_Number = `CHOP number`,
         CBC_Date = `CBC date`,
         Total_WBC = `Total WBC (x10.e3/uL)`,
         Total_Neutrophil = `Total neutrophil`,
         Total_Bands = `Total bands`, 
         Bandemia_Grade = `Bandemia grade`,
         Weight_kg = kg,
         Surface_Area_m2 = `m^2`,
         Notes = `...15`) %>% 
  mutate(Patient = factor(Patient),
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7,
         Total_Bands_Rescaled = Total_Bands*10,
         log_Total_Bands = if_else(Total_Bands == 0, 
                                   true = -5, 
                                   false = log(Total_Bands)),
         log_Total_Bands_Rescaled = if_else(Total_Bands_Rescaled == 0, 
                                   true = -3, 
                                   false = log(Total_Bands_Rescaled)),
         Bandemia = if_else(Total_Bands >= 0.2, true = 1, false = 0))

# Show the updated variable names to confirm all looks right. 
names(All_CHOP)
```

::: {.callout-note}
In previous versions of the output (04/02/2024 and earlier), Ruxin had been 
recoding Patient_Number values > 28 by subtracting 1 from them. Steve removed 
that code because it makes no functional difference to our analyses and it is 
easier to cross-reference back to source data file if we don't do that. 
:::

## Subset the Data 
Here, we select a subset of records that we want to use for the study, namely 
we retain only records for the first phase of the first instance of using the 
CHOP protocol (`CHOP_Number` = 1) for each patient. We retain only data about 
drug 1 (vincristine), drug 2 (cyclophosphamide), and drug 3 (doxorubicin). The drug
selection is motivated by the fact that we don't have enough data about other
drugs.

```{r}
#| label: create-Subset_CHOP

Subset_CHOP <- All_CHOP %>% 
  filter(CHOP_Number == 1) %>% 
  filter(Phase_of_Treatment==1) %>% 
  filter(Drug %in% 1:3)
```

## Check for Negative Weeks
Here, we examine when a CBC lab sample was collected via the `Weeks` variable, 
which measures how many weeks had elapsed between the start of the CHOP protocol 
and collection of the CBC sample. This should always be a positive number, so 
any negative values are implausible and should be corrected if possible. 

@tbl-negative-weeks shows that there is only one patient with `Weeks < 0`. The 
`CBC_Date` is clearly implausible. 

```{r}
#| label: tbl-negative-weeks
#| tbl-cap: Records with Implausible Visit Timing (Weeks < 0)

Subset_CHOP %>% 
  filter(Weeks < 0) %>% 
  select(Patient, MRN, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable() %>% 
  kable_styling()
```

So, now we inspect data for `Patient = 80` more closely to determine if we can 
fix the anomalous date. @tbl-patient80 shows all the records for that patient 
and it is clear that visit 3 should have occurred in the year 2020 instead of 
1930. We assume the month and day are accurate because after correcting the 
year, the `CBC_Date` for visit 3 would fall between those for visits 2 and 4. 
The sequence would then be sensible. 

```{r}
#| label: tbl-patient80

Subset_CHOP %>% 
  filter(Patient == 80) %>%
  select(Patient, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable() %>%
  kable_styling()
```

Below we fix the `CBC_Date` and then re-calculate variables that depended on it. 

```{r}
#| label: fix-patient80
Subset_CHOP <- Subset_CHOP %>% 
  mutate(CBC_Date = if_else(Patient == 80 & CBC_Date == as.POSIXct("1930-01-30", 
                                                                   tz = "UTC"), 
                            true =  as.POSIXct("2020-01-30", tz = "UTC"), 
                            false = CBC_Date),
         # Re-compute variables that depend on CBC_Date. 
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7)
```

## Check for Outlier Weeks
We've fixed negative values for Weeks, but still need to look for any other 
outlier values that may be data quality problems. So, we look at a summary by
visit number. We're looking for aggregate statistics that don't make sense and 
could be caused by data entry errors in the `CBC_Date` variable. 
@tbl-Visit-Timing1 shows descriptive statistics for the visit timing broken down 
by visit number. 

```{r}
#| label: tbl-Visit-Timing1 
# Days from the start date

Subset_CHOP %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

:::{.callout-caution}
Note: The mean, SD, and maximum number of weeks since CHOP started are all 
unexpectedly large at visit 3 and don't make sense given the expected timing of 
events under the CHOP protocol. That implies a data problem we need to 
investigate and fix. 
:::

@tbl-outlier-weeks shows the records for patient visits where `Weeks > 10`, 
which was a rough guess at a potential value that would be unusually high. We 
can see that Patient = 62 is a very clear outlier, with a CBC_Date in the 
future. That's clearly a data entry error. 

```{r}
#| label: tbl-outlier-weeks
#| tbl-cap: Visit 3 Records with Implausible Visit Timing (Weeks > 10)

Subset_CHOP %>% 
  filter(Visit_Number == 3 & Weeks > 10) %>% 
  select(Patient, MRN, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable() %>% 
  kable_styling()
```

So, now we inspect data for `Patient = 62` more closely to determine if we can 
fix the anomalous date. @tbl-patient62 shows all the records for that patient 
and it is clear that visit 3 should have occurred in the year 2018 instead of 
2028 We assume the month and day are accurate because after correcting the 
year, the `CBC_Date` for visit 3 would fall between those for visits 2 and 4. 
The sequence would then be sensible. 

```{r}
#| label: tbl-patient62

Subset_CHOP %>% 
  filter(Patient == 62) %>%
  select(Patient, Visit_Number, CHOP_Start_Date, CBC_Date, Weeks) %>% 
  kable() %>%
  kable_styling()
```

```{r}
#| label: fix-patient62
Subset_CHOP <- Subset_CHOP %>% 
  mutate(CBC_Date = if_else(Patient == 62 & CBC_Date == as.POSIXct("2028-05-25", 
                                                                   tz = "UTC"), 
                            true =  as.POSIXct("2018-05-25", tz = "UTC"), 
                            false = CBC_Date),
         # Re-compute variables that depend on CBC_Date. 
         Days_Since_Start = as.numeric(CBC_Date - CHOP_Start_Date, units = "days"),
         Weeks = Days_Since_Start/7)
```


## ReCheck for Outlier Weeks
We've fixed negative values for Weeks, but still need to look for any other 
outlier values that may be data quality problems. So, we look at a summary by
visit number. We're looking for aggregate statistics that don't make sense and 
could be caused by data entry errors in the `CBC_Date` variable. 
@tbl-Visit-Timing2 shows descriptive statistics for the visit timing broken down 
by visit number. 

```{r}
#| label: tbl-Visit-Timing2 
# Days from the start date

Subset_CHOP %>% 
  group_by(Visit_Number) %>% 
  summarise(mean_weeks = mean(Weeks, na.rm = TRUE),
            sd_weeks = sd(Weeks, na.rm = TRUE),
            min_weeks = min(Weeks, na.rm = TRUE),
            max_weeks = max(Weeks, na.rm = TRUE)) %>% 
  kable(digits = 2) %>% 
  kable_styling()
```

:::{.callout-tip}
Note: The descriptive statistics now look much more sensible. 
:::

@fig-visualize_weeks is a boxplot that visualize the information of weeks.

```{r}
#| label: fig-visualize_weeks
#| fig-cap: Time since first visit in weeks per visit
boxplot(Weeks ~ Visit_Number, data = Subset_CHOP, xlab="Visit Number")
```

## Further Subsetting of the Data
After reading peer review feedback on a manuscript that got rejected, the team 
decided on 2024-08-29 to narrow the focus of the analyses. This means we will 
now focus only on data for patients who had their first visit within 2 weeks of 
their CHOP start date. This section implements that decision. 

``` {r}
#| label: tbl-visit1-weeks
#| tbl-cap: Visit 1 Frequency Distribution for Weeks Since CHOP Start

# CONTINUE HERE
```


# Save Data

``` {r}
#| label: create-save-what

# Character vector of object names that will be saved.
save_what <- c("Raw_CHOP", "All_CHOP", "Subset_CHOP")
```

``` {r}
#| label: save-data

save(list = save_what, 
     file=here("data/Imported_CHOP_Data.RData"))
```

::: {.callout-tip}
The current data file can be loaded into R by copying the following code chunk 
into a script, setting the chunk option `eval: true`, and running the chunk. 
That will overwrite any objects already in memory that have the same names as 
the objects being loaded.

``` {r}
#| label: load-data
#| eval: false

load(file=here("data/Imported_CHOP_Data.RData"))
```
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies:

```{r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
