---
title: "Render CHOPStudy Scripts"
subtitle: | 
          | CSTAT Case: C1358
          | Client: Masyr, Alison
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
bibliography: references.bib
csl: apa.csl
params: 
  SourceDir: "scripts/" 
  SourceFile: "Render_Scripts.qmd" 
  LogFile: "Render_Scripts.html"
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
format: 
  html:
    output-file: "Render_Scripts.html"
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: true
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  message: true
  warning: false
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true
---

# Purpose
This file is part of a research compendium [@Pierce-RN8714]. It renders other 
scripts that import, manage, and analyze data for use a manuscript 
[@Eliason-RN8603]. Specifically, this Quarto script is a convenient way to 
render the other various scripts in a sensible order and with useful parameter 
values. It can be used interactively to render selected scripts, or you can just 
render this entire file to automate rendering all the scripts shown below. 

# Assumptions
All the assumptions built into the various scripts being rendered must be 
satisfied.

# Setup
This section contains chunks that you must run before trying to use the code 
chunks below it to interactively render a specific script. 

Global R chunk options are defined in the YAML header but local chunk options 
will over-ride global options. We can temporarily disable an individual chunk by 
inserting `#| eval: false` on a line at the top of the chunk. 

## Declare Path
This next chunk declares the path to this script relative to the project-level 
root directory. If the file is not in the right location under the project root
you'll get a warning message. This helps ensure relative paths are all working 
as expected. 

``` {r}
#| label: declare-path

# Declare path to this script relative to the project root directory.
here::i_am(path = paste0(params$SourceDir, params$SourceFile))
```

## Load Packages & Set Package Options
Next we load R packages required for this script to work. 

``` {r}
#| label: load-packages

library(devtools)     # for session_info()
library(here)         # for here()
library(rmarkdown)    # for pandoc_version()
library(quarto)       # for quarto_render(), quarto_version()
library(CHOPStudy)    # for package version information
library(piercer)      # for all_classes(), git_report()
# Packages loaded below here were used in other scripts rendered below. Loading 
# them here just makes the Software Information section output from this script 
# more complete. 
library(tidyverse)    # loads dplyr, forcats(), ggplot2, lubridate, stringr, 
                      # tibble(), tidyr all at once.
                      # for %>%, filter(), mutate(), map_dfr(), map_chr(), 
                      # rowid_to_colum(), and more...
library(broom)        # for glance(), tidy()
library(broom.mixed)  # for support for mixed models. 
library(DHARMa)       # for plotQQunif(), plotResiduals(), simulateResiduals(), 
                      # testDispersion()
library(ggpmisc)      # for stat_correlation()
library(glmmTMB)      # for diagnose(), glmmTMB()
library(knitr)        # for kable()
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)   # for kable_styling(), add_header_above(), column_spec(), 
                      # collapse_rows()
library(labelled)     # for set_value_labels(), set_variable_labels(), etc.
library(marginaleffects)  # for predictions(), comparisons(), slopes(), etc. 
library(modelsummary) # for modelsummary()
library(patchwork)    # for plot_layout()
library(performance)  # for check_model(), icc(), model_performance(), r2(), 
                      # test_lrt()
library(readxl)       # for read_xlsx()
library(tinytex)      # for install_tinytex()
```

# Import Data

``` {r}
#| label: Import-Data
#| eval: true

# Memory garbage collection (for more consistent chunk timing).
gc()
# Start chunk timer.
chunk.start <- proc.time()

OutFile <- paste0("Import_Data_", Sys.Date(), ".html")

# Create an object holding arguments for quarto_render(). 
quarto_render(input = here("scripts/Import_Data.qmd"),
              output_format = "html",
              output_file = OutFile, 
              execute_params = list(
                    SourceFile = "Import_Data.qmd", 
                    LogFile = OutFile),
              execute_dir = here::here("scripts/"),
              as_job = "auto",
              quarto_args = c("--output", OutFile))

# End chunk timer
chunk.end <- proc.time()
chunk.end - chunk.start
```

# Analyze Data

``` {r}
#| label: Descriptive-Analysis
#| eval: true

# Memory garbage collection (for more consistent chunk timing).
gc()
# Start chunk timer.
chunk.start <- proc.time()

OutFile <- paste0("Descriptive_Analysis_", Sys.Date(), ".html")

# Create an object holding arguments for quarto_render(). 
quarto_render(input = here("scripts/Descriptive_Analysis.qmd"),
              output_format = "html",
              output_file = OutFile, 
              execute_params = list(
                    SourceFile = "Descriptive_Analysis.qmd", 
                    LogFile = OutFile),
              execute_dir = here::here("scripts/"),
              as_job = "auto",
              quarto_args = c("--output", OutFile))

# End chunk timer
chunk.end <- proc.time()
chunk.end - chunk.start
```

``` {r}
#| label: GLMM-Analysis
#| eval: true

# Memory garbage collection (for more consistent chunk timing).
gc()
# Start chunk timer.
chunk.start <- proc.time()

OutFile <- paste0("GLMM_Analysis_", Sys.Date(), ".html")

# Create an object holding arguments for quarto_render(). 
quarto_render(input = here("scripts/GLMM_Analysis.qmd"),
              output_format = "html",
              output_file = OutFile, 
              execute_params = list(
                    SourceFile = "GLMM_Analysis.qmd", 
                    LogFile = OutFile),
              execute_dir = here::here("scripts/"),
              as_job = "auto",
              quarto_args = c("--output", OutFile))

# End chunk timer
chunk.end <- proc.time()
chunk.end - chunk.start
```

# References
::: {#refs}
:::

# Software Information
This section documents information that is important for reproducibility. Most
users will not need to read it. It is primarily here for use by the statistician
on the team.

We used [R](https://www.r-project.org/) as our main computing environment and
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto.

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.

## Versions
This document was generated using the following computational environment and 
dependencies: 

``` {r}
#| label: show-version
#| echo: true

# Get R and R package version numbers in use.
devtools::session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
