---
title: "Render CHOPStudy Scripts"
author: 
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
params: 
  SourceDir: "scripts/" 
  SourceFile: "Render_Scripts.qmd" 
  LogFile: "Render_Scripts.pdf"
execute-dir: file
format: 
  pdf:
    output-file: "Render_Scripts.pdf"
    output-ext: "pdf"
    documntclass: scrartcl
    papersize: letter
    geometry: 
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    colorlinks: true
    template-partials: 
      - title.tex
    include-in-header: 
      - file: compact-title.tex
      - text: | 
          \usepackage{fancyhdr}
          \usepackage[noblocks]{authblk}
          \renewcommand*{\Authsep}{, }
          \renewcommand*{\Authand}{, }
          \renewcommand*{\Authands}{, }
          \renewcommand\Affilfont{\small}
          \usepackage[yyyymmdd,hhmmss]{datetime}
          \usepackage{lastpage}         %% For pageref command
          \usepackage{fontspec}
          \defaultfontfeatures{Ligatures=TeX}
          \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
          \usepackage{url}
          \usepackage{floatrow}         %% Enable controlling float placement
          \floatsetup[table]{capposition=top} % Caption table top.
          \floatsetup[longtable]{margins=centering} % Longtable tables centered.
          \floatplacement{figure}{!ht}
          \floatplacement{table}{!ht}
          \usepackage{verbatim}         %% For the comment command
          \usepackage{placeins}         %% For FloatBarrier command
          \usepackage{booktabs}         %% Used by kableExtra
          \usepackage{longtable}        %% Used by kableExtra
          \usepackage{array}            %% Used by kableExtra
          \usepackage{multirow}         %% Used by kableExtra
          \usepackage{wrapfig}          %% Used by kableExtra
          \usepackage{float}            %% Used by kableExtra
          \usepackage{colortbl}         %% Used by kableExtra
          \usepackage{pdflscape}        %% Used by kableExtra
          \usepackage{tabu}             %% Used by kableExtra
          \usepackage[normalem]{ulem}   %% Used by kableExtra
          \usepackage{makecell}         %% Used by kableExtra
          \usepackage{xcolor}           %% Used by kableExtra
          \usepackage{threeparttable}   %% Used by kableExtra
          \usepackage{threeparttablex}  %% Used by kableExtra
          \usepackage[labelfont=bf, textfont=bf]{caption}  %% For bolded table/figure captions
          \usepackage{dcolumn}
          \usepackage{titling}
          \pretitle{\begin{center}\LARGE\bfseries}
          \posttitle{\end{center}}
          \pagestyle{fancy}
          \lhead{Render TEACHPilot Scripts}
          \rhead{\today\ \currenttime}
          \cfoot{ }
          \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
          \renewcommand{\headrulewidth}{0.4pt}
          \renewcommand{\footrulewidth}{0.4pt}
          \fancypagestyle{plain}{\pagestyle{fancy}}
execute: 
  eval: true 
  echo: fenced 
  output: true 
  message: true 
  warning: true 
  error: true 
  include: true 
knitr:
  opts_chunk: 
    message: true
    cfsize: "scriptsize"
---

\begin{comment}
# Hidden: Notes for User
This script makes use of the LaTeX `verbatim` package's `\comment` command to
build in hidden headings and text that are here for the team preparing the
report but will be omitted from the knitted PDF output file. This makes it
easier to put important comments directly in the file while still creating fully
formatted output that only shows the desired content. Text inside a comment
block will be omitted from the PDF output.

We can also use code chunks with chunk options such as `#| include: false`, 
`#| echo: false`, `#| message: false`, `#| warning: false` and so on to 
precisely control what R output shows up in the final PDF file. Thus, a script 
file can both serve as process documentation and efficiently generate the final 
version of a report output by applying the reproducible research and literate
programming concepts and tools. 

The `\lfoot` command just below sets the left footer to match the script's 
`LofgFile` parameter value, which we can pass along when rendering the script. 
Ideally it would be in the YAML `include-in-header:` with other pieces, but
I can't get inline code to work if I do it that way. This is the only way I 
could get the filename-based footer to work. 

The LaTeX `placeins` package provides the `\FloatBarrier` command. Using that 
before section headings help keep the figures and tables in their respective 
sections. 
\end{comment}

\lfoot{\texttt{\small \detokenize{`r params$LogFile`}}} 

\FloatBarrier

# Purpose
This Quarto script is a convenient way to render various scripts in a sensible 
order and with useful parameter values. It can be used interactively to render 
selected scripts, or you can just render this entire file to automate rendering 
all the scripts shown below. 

\FloatBarrier

# Assumptions
All the assumptions built into the various scripts being rendered must be 
satisfied.

\FloatBarrier

# Setup
This section contains chunks that you must run before trying to use the code 
chunks below it to interactively render a specific script. 

Global R chunk options are defined in the YAML header but local chunk options 
will over-ride global options. We can temporarily disable an individual chunk by 
inserting `#| eval: false` on a line at the top of the chunk. The method for
creating a `cfsize` option that controls font size in code chunks and their text
output is based on an answer to a question posted on
[stackoverflow.com](https://stackoverflow.com/a/46526740).

``` {r}
#| label: global-options
#| message: true

# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", 
         paste0("\n \\", options$cfsize,"\n\n", x, "\n\n \\normalsize"), 
         x)
  })
```

\FloatBarrier

## Declare Path
This next chunk declares the path to this script relative to the project-level 
root directory. If the file is not in the right location under the project root
you'll get a warning message. This helps ensure relative paths are all working 
as expected. 

``` {r}
#| label: declare-path

# Declare path to this script relative to the project root directory.
here::i_am(path = paste0(params$SourceDir, params$SourceFile))
```

\FloatBarrier

## Load Packages & Set Package Options
Next we load R packages required for this script to work. 

``` {r}
#| label: load-packages

library(devtools)     # for session_info()
library(here)         # for here()
library(rmarkdown)    # for pandoc_version()
library(quarto)       # for quarto_render(), quarto_version()
library(CHOPStudy)    # for package version information
library(piercer)      # for all_classes(), git_report(), which_latex()
# Packages loaded below here were used in other scripts rendered below. Loading 
# them here just makes the Software Information section output from this script 
# more complete. 
library(tidyverse)    # loads dplyr, ggplot2, lubridate, stringr, tidyr all at once.
                      # for %>%, filter(), mutate(), map_dfr(), map_chr(), 
                      # rowid_to_colum()
library(knitr)        # for kable()
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)   # for kable_styling(), add_header_above(), column_spec(),collapse_rows()
library(haven)        # for ymd(), mdy() and %within% operator.
library(labelled)     # for set_value_labels(), set_variable_labels(), etc.
```

\FloatBarrier

# Import Data

``` {r}
#| label: Import-Data
#| eval: true

# Memory garbage collection (for more consistent chunk timing).
gc()
# Start chunk timer.
chunk.start <- proc.time()

OutFile <- paste0("Import_Data_", Sys.Date(), ".html")

# Create an object holding arguments for quarto_render(). 
quarto_render(input = here("scripts/Import_Data.qmd"),
              output_format = "html",
              output_file = OutFile, 
              execute_params = list(
                    SourceDir = params$SourceDir, 
                    SourceFile = "Import_Data.qmd", 
                    LogFile = OutFile),
              execute_dir = here::here("scripts/"),
              as_job = "auto")

# Now copy it to the desired path and delete the original file
move_file(from = OutFile, 
          to = "output/", overwrite = TRUE) 

# End chunk timer
chunk.end <- proc.time()
chunk.end - chunk.start
```

\FloatBarrier

# Analyze Data

``` {r}
#| label: Analyze-Data
#| eval: true

# Memory garbage collection (for more consistent chunk timing).
gc()
# Start chunk timer.
chunk.start <- proc.time()

OutFile <- paste0("Analyze_Data_", Sys.Date(), ".html")

# Create an object holding arguments for quarto_render(). 
quarto_render(input = here("scripts/Analyze_Data.qmd"),
              output_format = "html",
              output_file = OutFile, 
              execute_params = list(
                    SourceDir = params$SourceDir, 
                    SourceFile = "Analyze_Data.qmd", 
                    LogFile = OutFile),
              execute_dir = here::here("scripts/"),
              as_job = "auto")

# Now copy it to the desired path and delete the original file
move_file(from = OutFile, 
          to = "output/", overwrite = TRUE) 

# End chunk timer
chunk.end <- proc.time()
chunk.end - chunk.start
```

# Software Information
We used [R](https://www.r-project.org/) as our main computing environment and 
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used 
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto. 

- Software chains: 
    - **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > html file**.
    - **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > tex file > TinyTeX > PDF file**.
- Source file: `r params$SourceFile`
- Output file: `r params$LogFile` 
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.
- [TinyTeX](https://yihui.org/tinytex/) compiles LaTeX files (`*.tex`) into PDF 
  files. It should be viable to use [MiKTeX](https://miktex.org) or another 
  LaTeX distribution instead. 

\FloatBarrier

## Versions
This document was generated using the following computational environment and 
dependencies: 

``` {r}
#| label: show-version
#| cfsize: scriptsize
#| echo: true

# Check and report whether we used TinyTex or other LaTeX software. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

\FloatBarrier

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| cfsize: scriptsize
#| echo: true

git_report()
```
